### 2014-08-20 inst/doc/article.Rnw
version 1.1

\documentclass[article, nojss]{jss}

%\VignetteIndexEntry{Introduction to TDA}
%% need no \usepackage{Sweave.sty}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{inputenc}

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{verbatim,float}
\usepackage{graphicx,subfigure,url}
\usepackage{dsfont,bm,color,multirow}

%spacing for captions
\usepackage{sidecap}
\usepackage[font=small,skip=0pt]{caption}
\setlength{\captionmargin}{20pt}

\newcommand{\todo}[1]{{\color{red}[[\textbf{XXX TODO: }#1]]}}

\newtheorem{algorithm}{Algorithm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{conjecture}{Conjecture}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\Hrule}{\noindent\rule{\linewidth}{0.1mm}}

\def\E{\mathrm{E}}
\def\P{\mathrm{P}}
\def\Cov{\mathrm{Cov}}
\def\sign{\mathrm{sign}}
\def\tr{\mathrm{tr}}
\def\col{\mathrm{col}}
\def\row{\mathrm{row}}
\def\nul{\mathrm{null}}
\def\rank{\mathrm{rank}}
\def\nuli{\mathrm{nuli}}
\def\half{\frac{1}{2}}
\def\hbeta{\hat\beta}
\def\hu{\hat{u}}
\def\hy{\hat{y}}
\def\tbeta{\tilde{\beta}}
\def\tu{\tilde{u}}
\def\ty{\tilde{y}}
\def\tA{\widetilde{A}}
\def\tD{\widetilde{D}}
\def\tG{\widetilde{G}}
\def\tP{\widetilde{P}}
\def\tQ{\widetilde{Q}}
\def\tR{\widetilde{R}}
\def\lone{1}
\def\ltwo{2}
\def\linf{\infty}
\def\cA{\mathcal{A}}
\def\cB{\mathcal{B}}
\def\cF{\mathcal{F}}
\def\cG{\mathcal{G}}
\def\cL{\mathcal{L}}
\def\cM{\mathcal{M}}
\def\cN{\mathcal{N}}
\def\cS{\mathcal{S}}
\def\cT{\mathcal{T}}
\def\T{^T}
\def\R{\mathbb{R}}
\def\old{\mathrm{old}}
\def\wBox{{\color{white} \Box}}
\def\sw{\setlength{\arraycolsep}{1pt}}
\def\sh{\renewcommand{\arraystretch}{0.75}}

\DefineVerbatimEnvironment{Sinput}{Verbatim} {xleftmargin=2em}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{xleftmargin=2em}
\DefineVerbatimEnvironment{Scode}{Verbatim}{xleftmargin=2em}
\fvset{listparameters={\setlength{\topsep}{0pt}}}
\renewenvironment{Schunk}{\vspace{\topsep}}{\vspace{\topsep}}

\author{Brittany T. Fasy \and Fabrizio Lecci \\ \small{In collaboration with the CMU TopStat Group}}
\Plainauthor{Brittany Terese Fasy, Fabrizio Lecci}

\title{Introduction to the \pkg{TDA} package}
\Plaintitle{Introduction to the TDA package}
\Abstract{
We present a short tutorial and introduction to using the \proglang{R} package 
\pkg{TDA}, which provides some tools for Topological Data Analysis.
In particular, it includes implementations of functions that, given some data, provide 
topological information about the underlying space, such as the distance function, the distance to a measure,
the kNN density estimator, the kernel density estimator, and the kernel distance.
The salient topological features of the sublevel sets (or superlevel sets) of these functions
can be quantified with persistent homology. We provide an \proglang{R} interface
for the efficient algorithms of the \proglang{Dionysus} C++ library, including a function for 
the persistent homology of the Rips filtration, and for the persistent homology of sublevel
sets (or superlevel sets) of arbitrary functions evaluated over a grid of points.
The significance of the features in the resulting persistence diagrams can be 
analyzed with functions that implement the methods discussed in \cite{fasy2014statistical},
\cite{chazal2013stochastic} and \cite{chazal2014robust}.
The \proglang{R} package \pkg{TDA} also includes the implementation of an algorithm for density 
clustering, which  allows us to identify  the spatial organization of the probability mass associated to 
a density function and visualize it by means of a dendrogram, the cluster tree. 
}
\Keywords{Topological Data Analysis, Persistent Homology, Density Clustering}

\Address{
  Brittany T. Fasy\\
  Computer Science Department\\
  Tulane University\\
  Email: \email{brittany.fasy@alumni.duke.edu}\\

  Fabrizio Lecci\\
  Department of Statistics\\
  Carnegie Mellon University\\
  Email: \email{lecci@cmu.edu}\\

  TopStat Group\\
  Carnegie Mellon University\\
  Website: \url{http://www.stat.cmu.edu/topstat/} \\
  Email: \email{topstat@stat.cmu.edu}

}

<<initial_settings, echo=FALSE>>=
options(stringsAsFactors=FALSE)
options(width=65)
options(prompt="> ")
@

\begin{document}

\section{Introduction}
\label{sec:intro}
Topological Data Analysis (TDA) refers to a collection of methods for finding topological structure in data \citep{carlsson2009topology}. Recent advances in computational topology have made it possible to actually compute topological invariants from data.
The input of these procedures typically takes the form of a point cloud, regarded as possibly noisy observations from an unknown lower-dimensional set $S$ whose interesting topological features were lost during sampling. The output is a collection of data summaries that are used to estimate the topological features of $S$.

In Section \ref{sec:distances}, we describe how to compute some widely studied functions that, starting from a point cloud, provide some topological information about the underlying space: the distance function (\code{distFct}), the distance to a measure function (\code{dtm}), the k Nearest Neighbor density estimator (\code{knnDE}), the kernel density estimator (\code{kde}), and the kernel distance (\code{kernelDist}).

One approach to TDA is persistent homology \citep{edelsbrunner2010computational}, a method for studying the homology at multiple scales simultaneously. More precisely, it provides a framework and efficient algorithms to quantify the evolution of the topology of a family of nested topological spaces. Given a real-valued function $f$, such as the ones described in Section~\ref{sec:distances}, persistent homology describes how the topology of the lower level sets $\{x: f(x) \leq t \}$ (or superlevel sets $\{x: f(x) \geq t \}$) changes as $t$ increases from $-\infty$ to $\infty$ (or decreases from $\infty$ to $-\infty$). 
This information is encoded in the persistence diagram, a multiset of points in the plane, each corresponding to the birth and death of a homological feature that existed for some interval of $t$.
In Section \ref{sec:persistent}, we describe how to compute persistence diagrams of sublevel sets (or superlevel sets) of functions evaluated over a grid of points, using \code{gridDiag}, and how to compute the persistence diagram of the Rips filtration built on top of a point cloud, using \code{ripsDiag}. Our implementation relies on \pkg{Dionysus}, a \proglang{C++} library developed by Dmitriy Morozov and freely available at http://mrzv.org/software/dionysus/ .

One of the key challenges in persistent homology is to find a way to isolate the points of the persistence diagram representing the topological noise. In \cite{fasy2014statistical}, \cite{chazal2013stochastic}, and \cite{chazal2014robust}, we propose several statistical methods to construct confidence sets for persistence diagrams and other summary functions that allow us to separate topological signal from topological noise.
Some of these methods are implemented and described in Section \ref{sec:persistent}.

Finally, the \pkg{TDA} package provides the implementation of an algorithm for density clustering. 
This method allows us to identify and visualize the spatial organization of the data, without specific knowledge about the data generating mechanism and in particular without any a priori information about the number of clusters. 
In Section \ref{sec:density}, we describe the function \code{clusterTree}, that, given a density estimator, encodes the hierarchy of the connected components of its superlevel sets into a dendrogram, the cluster tree. \citep{KpoLux11, kent2013}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Distance Functions and Density Estimators} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%
\label{sec:distances}

As a first toy example to using the \pkg{TDA} package, we show how to compute
distance functions and density estimators over a grid of points.
The setting is the typical one in TDA: a set of points $X=\{x_1, \dots, x_n\}$ 
has been sampled from some distribution $P$ supported on a manifold and we 
are interested in recovering the topological features of the underlying manifold
by studying some functions of the data. 
The following code generates a sample of 400 points from the unit circle and constructs
a grid of points over which we will evaluate the functions.

<<eq1, echo=TRUE>>=
library("TDA")
X = circleUnif(400)

Xlim=c(-1.6, 1.6);   Ylim=c(-1.7, 1.7);   by=0.065

Xseq=seq(Xlim[1], Xlim[2], by=by)
Yseq=seq(Ylim[1], Ylim[2], by=by)
Grid=expand.grid(Xseq,Yseq)
@

The \pkg{TDA} package provides implementations of the following functions:
\begin{itemize}
\item The distance function is defined for each $y \in \mathbb{R}^d$ as
$
\Delta(y)= \inf_{x \in X} \| x-y \|_2
$
and is computed for each point of the \code{Grid} with the following code:
<<eq2, echo=TRUE>>=
distance= distFct(X=X, Grid=Grid)
@
\item Given a probability measure $P$, the distance to measure (DTM) is defined for each $y \in \mathbb{R}^d$ as
$$
d_{m_0}(y) =\sqrt{ \frac{1}{m_0}\int_0^{m_0} ( G_y^{-1}(u))^2 du},
$$
where
$G_y(t) = P( \Vert X-y \Vert \leq t)$ and $0<m_0<1$ is a smoothing parameter. For more details
see \cite{chazal2011geometric}.\\
Given $X=\{x_1, \dots, x_n\}$, the empirical version of the DTM is
$$
\hat d_{m_0}(y) = \sqrt{ \frac{1}{k} \sum_{x_i \in N_k(y)} \Vert x_i-y \Vert^2 },
$$
where $k= \lceil m_0 n \rceil$ and $N_k(y)$ is the 
set containing the $k$ nearest neighbors 
of $y$ among $x_1, \ldots, x_n$.
The DTM is computed for each point of the \code{Grid} with the following code:
<<eq3, echo=TRUE>>=
m0=0.1
DTM=dtm(X=X, Grid=Grid, m0=m0)
@
\item The $k$ Nearest Neighbor density estimator, for each $y \in \mathbb{R}^d$, is defined by
$$
\hat\delta_k(y)=\frac{k}{n \; v_d \; r_k^d(y)},
$$
where $v_n$ is the volume of the Euclidean $d$ dimensional unit ball and $r_k^d(x)$ is the Euclidean distance form point $x$ to its $k$th closest neighbor among the points of $X$.
It is computed for each point of the \code{Grid} with the following code:
<<eq4, echo=TRUE>>=
k=60
kNN=knnDE(X=X, Grid=Grid, k=k)
@
\item The Gaussian Kernel Density Estimator (KDE), for each $y \in \mathbb{R}^d$, is defined by
$$
\hat p_h(y)=\frac{1}{n (\sqrt{2 \pi} h )^d} \sum_{i=1}^n \exp \left( \frac{- \Vert y-x_i \Vert_2^2}{2h^2} \right).
$$
where $h$ is a smoothing parameter.
It is computed for each point of the \code{Grid} with the following code:
<<eq5, echo=TRUE>>=
h=0.3
KDE= kde(X=X, Grid=Grid, h=h)
@
\item The Kernel distance estimator, for each $y \in \mathbb{R}^d$, is defined by
$$
\hat \kappa_h(y)=\sqrt{ \frac{1}{n^2} \sum_{i=1}^n\sum_{j=1}^n K_h(x_i, x_j) + K_h(y,y) - 2 \frac{1}{n} \sum_{i=1}^n K_h(y,x_i)  },
$$
where $K_h(x,y)=\exp\left( \frac{- \Vert x-y \Vert_2^2}{2h^2} \right)$ is the Gaussian Kernel with smoothing parameter $h$.
The Kernel distance is computed for each point of the \code{Grid} with the following code:
<<eq6, echo=TRUE>>=
h=0.3
Kdist= kernelDist(X=X, Grid=Grid, h=h)
@
\end{itemize}
For this 2 dimensional example, we can visualize the functions using \code{persp} form the \pkg{graphics} package. For example the following code produces the KDE plot in Figure \ref{fig:eq8}:

<<eq7, echo=TRUE, eval=FALSE, height=6, width=6>>=
persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)),
xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, 
main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)
@

% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=6in, height=4.6in}
\begin{center}
<<eq8, echo=FALSE, eval=TRUE, fig=TRUE, height=5, width=6.5>>=
par(mfrow=c(2,3))
par(mai=c(0.6,0.25,0.3,0.1))
	plot(X, pch=16, cex=0.6, xlab="", ylab="", main="Sample X")

par(mai=c(0.1,0.15,0.1,0.1))
	persp(Xseq,Yseq,matrix(distance,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="Distance Function", d=0.5, scale=FALSE, expand=1, shade=0.9)

	persp(Xseq,Yseq,matrix(DTM,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="DTM", d=0.5, scale=FALSE, expand=1, shade=0.9)

	persp(Xseq,Yseq,matrix(kNN,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="kNN", d=0.5, scale=FALSE, expand=3, shade=0.9)

	persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)

	persp(Xseq,Yseq,matrix(Kdist,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="Kernel Distance", d=0.5, scale=FALSE, expand=5, shade=0.9)
@
\end{center}
\vspace{-1cm}
\caption{distance functions and density estimators evaluated over a grid of points.}
\label{fig:eq8}
\end{figure}

% PLOT END

\subsection{Bootstrap Confidence Bands}
\label{sec:bootstrap}

We can construct a $(1-\alpha)$ confidence band for a function using the bootstrap algorithm, which we briefly describe using the kernel density estimator:
\begin{enumerate}
\item Given a sample $X=\{x_1, \dots, x_n\}$, compute the kernel density estimator $\hat p_h$;
\item Draw $X^*=\{x_1^*, \dots, x_n^*\}$ from $X=\{x_1, \dots, x_n \}$ (with replacement), and compute $\theta^*=\sqrt{n} \Vert \hat p_h^*(x) - \hat p_h(x) \Vert_\infty$, where $\hat p_h^*$ is the density estimator computed using $X^*$;
\item Repeat the previous step $B$ times to obtain $\theta_{1}^*, \dots, \theta_{B}^*$;
\item Compute 
$
q_{\alpha} = \inf \left \{q: \frac{1}{B} \sum_{j=1}^B I(\theta_j^* \geq q) \leq \alpha \right\};
$
\item The $(1-\alpha)$ confidence band for $\mathbb{E}[\hat p_h]$ is
$
\left[\hat p_h - \frac{q_{\alpha}}{\sqrt{n}} \,, \, \hat p_h+ \frac{q_{\alpha}}{\sqrt{n}}\right] .
$
\end{enumerate}

\cite{fasy2014statistical} and \cite{chazal2014robust} prove the validity of the bootstrap algorithm for kernel density estimators, distance to measure, and kernel distance, and use it in the framework of persistent homology.
The bootstrap algorithm is implemented in the function \code{bootstrapBand}, which provides the option of parallelizing the algorithm (\code{parallel=TRUE}) using the package \pkg{parallel}.
The following code computes a 90\% confidence band for $\mathbb{E}[\hat p_h]$, showed in Figure \ref{fig:eq10}.

<<eq9, echo=TRUE, eval=TRUE>>=
band=bootstrapBand(X=X, FUN=kde, Grid=Grid, B=100, 
                   parallel=F, alpha=0.1, h=h)
@

% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=5in, height=2.5in}
\begin{center}
<<eq10, echo=FALSE, eval=TRUE, fig=TRUE, width=8, height=4>>=
posYgrid=which(Grid[,2]>0)
posYseq=which(Yseq>0)

par(mfrow=c(1,2))
par(mai=c(0.5,0.25,0.3,0.1))

persp(Xseq,Yseq,matrix(band$band[,1],ncol=length(Yseq), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = range(Yseq),
xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)

persp(Xseq,Yseq[posYseq],matrix(band$band[posYgrid,1],ncol=length(Yseq[posYseq]), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),
xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)

par(new = TRUE)

persp(Xseq,Yseq[posYseq],matrix(band$fun[posYgrid],ncol=length(Yseq[posYseq]), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),,
xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="red", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)

par(new = TRUE)

persp(Xseq,Yseq[posYseq],matrix(band$band[posYgrid,2],ncol=length(Yseq[posYseq]), nrow=length(Xseq)), zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),
xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, 
main="", d=0.5, scale=FALSE, expand=3, shade=0.9, box=F)
@
\end{center}
\vspace{-1cm}
\caption{the 90\% confidence band for $\mathbb{E}[\hat p_h]$ has the form $[\ell, u]= \left[\hat p_h - q_{\alpha}/\sqrt{n} \,, \, \hat p_h+ q_{\alpha}/\sqrt{n}\right]$. The plot on the right shows a section of the functions: the red surface is the KDE $\hat p_h$; the pink surfaces are $\ell$ and $u$.}
\label{fig:eq10}
\end{figure}
% PLOT END

%%%%%%%%%%%%%%%%%%
\section{Persistent Homology} 
%%%%%%%%%%%%%%%%%%
\label{sec:persistent}

We provide an informal description of the implemented methods of persistent homology. We assume the reader is familiar with the basic concepts and, for a rigorous exposition, we refer to the textbook \cite{edelsbrunner2010computational}.

%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Persistent Homology Over a Grid}
%%%%%%%%%%%%%%%%%%%%%%%
In this section, we describe how to use the \code{gridDiag} function to compute the persistent homology of sublevel (and superlevel) sets of the functions described in Section \ref{sec:distances}. 
The function \code{gridDiag} evaluates a given real valued function over a triangulated grid, constructs a filtration of simplices using the values of the function, and computes the persistent homology of the filtration. \code{gridDiag} currently supports grids in dimension 1,2, and 3. Future releases of the \pkg{TDA} package will extend its functionalities.

The following code computes the persistent homology of the superlevel sets (\code{sublevel=FALSE}) of the kernel density estimator (\code{FUN=kde}, \code{h=0.3}) using the point cloud stored in the matrix~\code{X} from the previous example.
The same code would work for the other functions defined in Section \ref{sec:distances} (it is sufficient to replace \code{kde} and its smoothing parameter \code{h} with another function and its parameters). 
The function \code{gridDiag} returns an object of the class \code{"diagram"}.  
The other inputs are the features of the grid over which the \code{kde} is evaluated (\code{Xlim}, \code{Ylim} and \code{by}), the smoothing parameter \code{h}, and a logical variable that indicates whether a progress bar should be printed (\code{printStatus}).

<<eq11, echo=TRUE, eval=TRUE>>=
Diag = gridDiag( X=X, FUN=kde, h=0.3, Xlim=Xlim, Ylim=Ylim, by=by, 
                 sublevel=F, printStatus=F ) 
@

We plot the data and the the diagram, using the function \code{plot}, implemented as a standard \code{S3} method for objects of the class \code{"diagram"}. The following command produces the third plot in Figure \ref{fig:eq11b}.

<<eq11b, echo=TRUE, eval=FALSE>>=
plot(Diag, band=2*band$width, main="KDE Diagram")
@

% PLOT STARTS
\begin{figure}[h!tb]
\setkeys{Gin}{width=5.5in, height=2in}
\begin{center}
<<eq11b, echo=FALSE, eval=TRUE, fig=TRUE, height=2, width=5.5>>=
par(mfrow=c(1,3))
par(mai=c(0.3,0.25,0.3,0.1))

plot(X, pch=16, cex=0.6, xlab="", ylab="",main="Sample X")

persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)

par(mai=c(0.55,0.6,0.3,0.1))

plot(Diag, band=2*band$width, main="KDE Diagram")
@
\end{center}
\caption{The plot on the right shows the persistence diagram of the superlevel sets of the KDE. Black points represent connected components and red triangles represent loops. 
The features are born at high levels of the density and die at lower levels. The pink 90\% confidence band separates significant features from noise.}
\label{fig:eq11b}
\end{figure}
% PLOT END

The option \code{band=2*band$width} produces a pink confidence band for the persistence diagram, using the confidence band constructed for the corresponding kernel density estimator in the previous section. 
The features above the band can be interpreted as representing significant homological features, while points in the band are not significantly different from noise.
The validity of the bootstrap confidence band for  persistence diagrams of KDE, DTM, and Kernel Distance derive from the \textit{Stability Theorem} \citep{chazal2012structure} and is discussed in detail in \cite{fasy2014statistical} and \cite{chazal2014robust}.\\

The function \code{plot} for the class \code{"diagram"} provide the options of rotating the diagram (\code{rotated=TRUE}), drawing the barcode in place of the diagram (\code{barcode=TRUE}), as well as other standard graphical options:

% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=4.5in, height=2.25in}
\begin{center}
<<eq11c, echo=TRUE, eval=TRUE,  fig=TRUE, width=6, height=3>>=
par(mfrow=c(1,2), mai=c(0.8,0.8,0.3,0.1))
plot(Diag, rotated=T, band=band$width, main="Rotated Diagram")
plot(Diag, barcode=T,  main="Barcode")
@
\end{center}
\caption{Rotated Persistence Diagram and Barcode}
\label{fig:eq11c}
\end{figure}
% PLOT END

%%%%%%%%%%%%%%%%%%
\subsection{Rips Diagrams}
%%%%%%%%%%%%%%%%%%
The {\em Vietoris-Rips~complex} $R(X,\varepsilon)$
consists of simplices with vertices in $X$ and diameter
at most $\varepsilon$.  In other words,
a simplex $\sigma$ is included in the complex if each pair of vertices in
$\sigma$ is at most $ \varepsilon$ apart.
The sequence of Rips complexes obtained by gradually
increasing the radius $\varepsilon$ creates a filtration.

The \code{ripsDiag} function computes the persistence diagram of the Rips filtration built on top of a point cloud.
This function is an R wrapper of the function ``rips-pairwise" of the \pkg{Dionysus} \proglang{C++}  library.

The following code generates 60 points from two circles:

<<eq12, echo=TRUE, eval=TRUE>>=
Circle1 = circleUnif(60)
Circle2 = circleUnif(60, r=2) +3
Circles=rbind(Circle1,Circle2)
@

We specify the limit of the Rips filtration and the max dimension of the homological features we are interested in (0 for components, 1 for loops, 2 for voids...):
<<eq12b, echo=TRUE, eval=TRUE>>=
maxscale=5           # limit of the filtration
maxdimension=1       # components and loops
@

and we generate the persistence diagram:
<<eq12c, echo=TRUE, eval=TRUE>>=
Diag=ripsDiag(X=Circles, maxdimension, maxscale, printStatus=F)
@
Alternatively, using the option \code{dist="arbitrary"} in \code{ripsDiag()}, the input \code{X} can be an $n \times n$ matrix of distances.
This option is useful when the user wants to consider a Rips filtration constructed using an arbitrary distance.\\

Finally we plot the data and the the diagram:\\
\phantom{o}
% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=4.6in, height=2.3in}
\begin{center}
<<eq12d, echo=TRUE, eval=TRUE,  fig=TRUE, height=2.8, width=5.6>>=
par(mfrow=c(1,2), mai=c(0.8,0.8,0.3,0.3))
plot(Circles, pch=16, xlab="",ylab="")
plot(Diag)
@
\end{center}
\caption{Rips persistence diagram. Black points represent connected components and red triangles represent loops.}
\label{fig:eq12d}
\end{figure}
% PLOT END

%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Bottleneck and Wasserstein Distances}
%%%%%%%%%%%%%%%%%%%%%%%%%%

Standard metrics for measuring the distance between two persistence diagrams are the bottleneck distance and the $p$th Wasserstein distance \citep{edelsbrunner2010computational}. 
The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \pkg{Dionysus} \proglang{C++}  library .

We generate two persistence diagrams of the Rips filtrations built on top of the two (separate) circles of the previous example,
<<eq13, echo=TRUE, eval=TRUE>>=
Diag1=ripsDiag(Circle1,maxdimension=1, maxscale=5, printStatus=F)
Diag2=ripsDiag(Circle2,maxdimension=1, maxscale=5, printStatus=F)
@
and we compute the bottleneck distance and the $2$nd Wasserstein distance between the two diagrams.
In the following code, the option \code{dimension=1} specifies that the distances between diagrams are computed using only one dimensional features (loops).
<<eq13b, echo=TRUE, eval=TRUE>>=
print( bottleneck(Diag1, Diag2, dimension=1) )
print( wasserstein(Diag1, Diag2, p=2, dimension=1) )
@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Landscapes and Silhouettes}
%%%%%%%%%%%%%%%%%%%%%%

Persistence landscapes and silhouettes are real-valued functions that further summarize the information contained in a persistence diagram. They have been introduced and studied in \cite{bubenik2012statistical}, \cite{chazal2013stochastic}, and \cite{chazal2014subsampling}.
We briefly introduce the two functions.

{\bf Landscape.}  The persistence landscape
is a collection of continuous, piecewise linear
functions~\mbox{$\lambda \colon  \mathbb{Z}^{+} \times \mathbb{R} \to \mathbb{R}$} that summarizes a
persistence diagram.
To define the landscape, consider the set of functions created by tenting
each each point $p=(x,y)= \left( \frac{b+d}{2}, \frac{d-b}{2}  \right)$
representing a birth-death pair $(b,d)$ in the persistence diagram $D$ as follows:
\begin{equation}\label{eq:triangle}
 \Lambda_p(t) =
 \begin{cases}
  t-x+y & t \in [x-y, x] \\
  x+y-t & t \in (x,  x+y] \\
  0 & \text{otherwise}
 \end{cases}
 =
 \begin{cases}
  t-b & t \in [b, \frac{b+d}{2}] \\
  d-t & t \in (\frac{b+d}{2}, d] \\
  0 & \text{otherwise}.
 \end{cases}
\end{equation}
%Notice that $p$ is itself on the graph of $\Lambda_p(t)$.
We obtain
an arrangement of piecewise linear curves by overlaying the graphs of
the functions~\mbox{$\{ \Lambda_p \}_{p}$}; see Figure \ref{fig:eq14a}.
The persistence landscape of $D$ is a  summary of this arrangement.
Formally, the persistence landscape of $D$ is the collection of functions
\begin{equation}\label{eq:landscape}
 \lambda(k,t) = \underset{p}{k\text{max}} ~ \Lambda_p(t), \quad
 t \in [0,T], k \in \mathbb{N},
\end{equation}
where $k$max is the $k$th largest value in the set; in particular,
$1$max is the usual maximum~function.

{\bf Silhouette.}
Consider a persistence diagram with $N$ off diagonal points.
For every $0 < p < \infty$
we define the power-weighted silhouette
$$
\phi^{( p)}(t) = \frac{\sum_{j=1}^N |d_j - b_j|^p \Lambda_j(t)}{\sum_{j=1}^N |d_j - b_j|^p}.
$$

The value $p$ can be thought of as a trade-off parameter between
uniformly treating all pairs in the persistence diagram 
and considering only the most persistent pairs.  Specifically, when $p$ is small,
$\phi^{( p)}(t)$ is dominated by the effect of low persistence features.
Conversely, when $p$ is large,
$\phi^{( p)}(t)$ is dominated by the most persistent features;
see Figure \ref{fig:eq14a} (right).

% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=6in, height=2in}
\begin{center}

<<eq14a, echo=FALSE, eval=TRUE, fig=TRUE, height=2.5, width=7.5>>=
PlotTriangles = function(left,right){
   n = length(left)
   X = (left+right)/2
   Y = (right-left)/2
   xlimit=c( 0, max(X+Y)*1.2 )
   ylimit=c(0, max(Y)*1.2)
   plot(X,Y,type="n",xlab="",ylab="",xaxt="n",yaxt="n", xlim=xlimit, ylim=ylimit, frame.plot=F, main="Triangles")
	axis(1)
	axis(2)
	mtext("(Birth+Death)/2", side=1, line=2.5, cex=1)
	mtext("(Death-Birth)/2", side=2, line=2.5, cex=1)
	#polygon(c(0,0,ylimit[2]),c(0,ylimit[2],ylimit[2]), col="gray87", border=NA)
   for (i in 1:n){
      a <- c(0, left[i], (left[i]+right[i])/2, right[i], 0)
      b <- c(0, 0, (right[i]-left[i])/2, 0, 0)
      lines(a,b, lwd=2)
	}
     B = max(X+Y)
     grid = seq(0,B,length=1000)
     maxfun = rep(0,length(grid))
     tmp = matrix(0,n,length(grid))
     for(i in 1:n){
          tmp[i,] = pmax(pmin(grid-left[i],right[i]-grid),0)
          }
     land = apply(tmp,2,max)
     #lines(grid,land,lwd=2, col=4)
	 points(X,Y,type="p", cex=1, pch=19, col="pink")
     points(X,Y,type="p")
}

par(mfrow=c(1,3), mai=c(0.7,0.7,0.3,0.3))

left = c(0,2,2,3.5,5)
right = c(2,6,3,5,8)

 PlotTriangles(left,right)

Diag1=cbind(rep(0,length(left)), left, right)
tseq=seq(min(Diag1[,2:3]), max(Diag1[, 2:3]), length=500)
Land1=landscape(Diag1, dimension=0, KK=1, tseq )
Sil1=silhouette(Diag1, p=1, dimension=0, tseq )

   X = (left+right)/2
   Y = (right-left)/2
   xlimit=c( 0, max(X+Y)*1.2 )
   ylimit=c(0, max(Y)*1.2)

plot(tseq, Land1, col=4, type="l", lwd=2, xlim=xlimit, ylim=ylimit, axes=F, main="1st Landscape", xlab="", ylab="")
axis(1)
axis(2)

plot(tseq, Sil1, col=4, type="l", lwd=2, xlim=xlimit, ylim=ylimit, axes=F, main="Silhouette p=1", xlab="", ylab="")
axis(1)
axis(2)
@
\end{center}
\caption{	Left: we use the rotated axes to represent a persistence diagram $D$.
	A feature $(b,d) \in D$ is represented by the point
	$(\frac{b+d}{2},\frac{d-b}{2})$ (pink).
	In words,
	the $x$-coordinate is the average parameter value over which the feature
exists,
	and the $y$-coordinate is the half-life of the feature.
        Middle: the blue curve is the landscape $\lambda(1,\cdot)$.
        Right: the blue curve is the silhouette $\phi^{(1)}(\cdot)$.}
\label{fig:eq14a}
\end{figure}
% PLOT END

\vspace{0.5cm}

The landscape and silhouette functions can be evaluated over a one-dimensional grid of points \code{tseq} using the functions \code{landscape} and \code{silhouette}. In the following code, we use the persistence diagram from Figure \ref{fig:eq12d} to construct the corresponding landscape and silhouette for 
one-dimensional features (\code{dimension=1}). The option \code{KK=1} specifies that we are interested in the 1st landscape function, and \code{p=1} is the power of the weights in the definition of the silhouette function.

<<eq14, echo=TRUE, eval=TRUE>>=
tseq=seq(0,maxscale, length=1000)   #domain
Land= landscape(Diag, dimension=1, KK=1, tseq)  
Sil=silhouette(Diag, p=1,  dimension=1, tseq)  
@

The functions \code{landscape} and \code{silhouette} return real valued vectors, which can be simply plotted with \code{plot(tseq, Land, type="l"); plot(tseq, Sil, type="l")}. See Figure \ref{fig:eq14b}.

% PLOT START
\begin{figure}[h!tb]
\setkeys{Gin}{width=5in, height=1.5in}
\begin{center}
<<eq14b, echo=FALSE, eval=TRUE,  fig=TRUE, height=2.4, width=8>>=
par(mfrow=c(1,2), mai=c(0.5,0.45,0.3,0.3))
plot(tseq, Land, type="l", lwd=3, ylab="", main="1st Landscape, dim=1",asp=1, col=2)
plot(tseq, Sil, type="l", lwd=3, ylab="", main="Silhouette (p=1), dim=1",asp=1, col=2)
@
\end{center}
\caption{Landscape and Silhouette of the one-dimensional features of the diagram of Figure \ref{fig:eq12d}.}
\label{fig:eq14b}
\end{figure}
% PLOT END

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Confidence Bands for Landscapes and Silhouettes}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Recent results in \cite{chazal2013stochastic} and \cite{chazal2014subsampling} show how to construct confidence bands for landscapes 
and silhouettes, using a bootstrap algorithm (multiplier bootstrap).
This strategy is useful in the following scenario.
We have a very large dataset with $N$ points. There is a diagram $D$ and landscape $\lambda$ corresponding to some filtration built on the data. When $N$ is large, computing $D$ is prohibitive. Instead, we draw $n$ subsamples, each of size $m$. We compute a diagram and landscape for each subsample yielding landscapes $\lambda_1, \dots, \lambda_n $. (Assuming $m$ is much smaller than $N$, these subsamples are essentially independent and identically distributed.) Then we compute $\frac{1}{n}\sum_i \lambda_i$, an estimate of $\mathbb{E}(\lambda_i)$, which can be regarded as an approximation of $\lambda$. The function \code{multipBootstrap} uses the landscapes $\lambda_1, \dots, \lambda_n$ to construct a confidence band for $\mathbb{E}(\lambda_i)$. The same strategy is valid for silhouette functions. We illustrate the method with a simple example.\\
First we sample $N$ points from two circles:

<<eq15, echo=TRUE, eval=TRUE>>=
N=4000
XX1 = circleUnif(N/2)
XX2 = circleUnif(N/2, r=2) +3
X=rbind(XX1,XX2)
@

Then we specify the number of subsamples $n$, the subsample size $m$, and we create the objects that
will store the $n$ diagrams and landscapes:

<<eq15b, echo=TRUE, eval=TRUE>>=
m=80        # subsample size
n=10        # we will compute n landscapes using subsamples of size m

tseq=seq(0,maxscale, length=500)          #domain of landscapes

Diags=list()                              #here we store n Rips diags
Lands=matrix(0,nrow=n, ncol=length(tseq)) #here we store n landscapes
@

For $n$ times, we subsample from the large point cloud, compute $n$ Rips diagrams and the corresponding
1st landscape functions (\code{KK=1}), using 1 dimensional features (\code{dimension=1}):

<<eq15c, echo=TRUE, eval=TRUE>>=
for (i in 1:n){
	subX=X[sample(1:N,m),]
	Diags[[i]]=ripsDiag(subX,maxdimension=1, maxscale=5)
	Lands[i,]=landscape(Diags[[i]], dimension=1, KK=1, tseq )
}
@

Finally we use the $n$ landscapes to construct a 95\% confidence band for the mean landscape

<<eq15d, echo=TRUE, eval=TRUE>>=
bootLand=multipBootstrap(Lands,B=100,alpha=0.05, parallel=F)
@

which is plotted by the following code. See Figure \ref{fig:eq15f}.

<<eq15e, echo=TRUE, eval=FALSE>>=
plot(tseq, bootLand$mean, main="Mean Landscape with 95% band")
polygon(c(tseq, rev(tseq)), 
        c(bootLand$band[,1],rev(bootLand$band[,2])), col="pink")
lines(tseq, bootLand$mean, lwd=2, col=2)
@

% PLOT STARTS
\begin{figure}[h!tb]
\setkeys{Gin}{width=5in, height=2.5in}
\begin{center}
<<eq15f, echo=FALSE, eval=TRUE,  fig=TRUE, height=4, width=8>>=
par(mfrow=c(1,2))
par(mai=c(0.5,0.45,0.3,0.3))

plot(X, pch=16, cex=0.5, main="Large Sample from Circles")

plot(tseq, bootLand$mean, type="l", lwd=2, xlab=" ", ylab="" ,main="Mean Landscape with 95% band", ylim=c(0,1.2))
polygon(c(tseq, rev(tseq)), c(bootLand$band[,1],rev(bootLand$band[,2])), col="pink")
lines(tseq, bootLand$mean, lwd=2, col=2)
@
\end{center}
\caption{95\% confidence band for the mean landscape function.}
\label{fig:eq15f}
\end{figure}
% PLOT END

%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Selection of Smoothing Parameters}
%%%%%%%%%%%%%%%%%%%%%%%%%

An unsolved problem in topological inference is how to
choose the smoothing parameters, for example $h$ for KDE and 
$m$ for DTM.

\cite{chazal2014robust} suggest the following method, that we describe here for the kernel density estimator,
but works also for the kernel distance and the distance to measure.

Let $\ell_1(h),\ell_2(h),\ldots, $
be the lifetimes
of the features of a persistence diagram at scale $h$.
Let $q_\alpha(h)/\sqrt{n}$
be the width of the confidence band for the kernel density estimator
at scale $h$, as described in Section \ref{sec:bootstrap}.
We define two quantities that measure
the amount of significant information at level $h$:
\begin{itemize}
\item The number of significant features, $N(h) = \# \left\{i:\ \ell(i) > 2\frac{q_\alpha(h)}{\sqrt{n}}\right\}$;
\item The total significant persistence, $S(h) = \sum_i \left[ \ell_i - 2\frac{q_\alpha(h)}{\sqrt{n}}\right]_+$.
\end{itemize}
These measures are small when $h$ is small since $q_\alpha(h)$ is large.
On the other hand,
they are small when $h$ is large since then all the features of the KDE are smoothed out.
Thus we have a kind of topological bias-variance tradeoff.
We choose $h$ to maximize $N(h)$ or $S(h)$.

The method is implemented in the function \code{maxPersistence}, as shown in the following toy example.
First, we sample 1600 point from two circles (plus some clutter noise) and we specify the limits of the grid over which the KDE will be evaluated:

<<eq16, echo=TRUE, eval=TRUE>>=
XX1 = circleUnif(600)
XX2 = circleUnif(1000, r=1.5) +2.5
noise=cbind(runif(80, -2,5),runif(80, -2,5))
X=rbind(XX1,XX2, noise)

# Grid limits
Xlim=c(-2,5)
Ylim=c(-2,5)
by=0.2
@

Then we specify a sequence of smoothing parameters among which we will select the optimal one, the number of bootstrap iterations and the level of the confidence bands to be computed:

<<eq16b, echo=TRUE, eval=TRUE>>=
parametersKDE=seq(0.1,0.6,by=0.05)

B=50        # number of bootstrap iterations. Should be large.
alpha=0.1   # level of the confidence bands
@

The function \code{maxPersistence} can be parallelized (\code{parallel=TRUE}) and a progress bar can be printed (\code{printProgress=TRUE}):

<<eq16c, echo=TRUE, eval=TRUE>>=
maxKDE=maxPersistence(kde, parametersKDE, X, Xlim, Ylim, Zlim=NA, 
                      by=by, sublevel=F, B=B, alpha=alpha, 
                      parallel=TRUE, printProgress=TRUE)
@

The \code{S3} methods \code{summary} and \code{plot} are implemented for the class \code{"maxPersistence"}.
We can display the values of the parameters that maximize the two criteria:
<<eq16d, echo=TRUE, eval=TRUE>>=
print(summary(maxKDE))
@

and produce the summary plot of Figure \ref{fig:eq16e}:\\

% PLOT STARTS
\begin{figure}[h!]
\setkeys{Gin}{width=5in, height=2.5in}
\begin{center}
<<eq16e, echo=TRUE, eval=TRUE,  fig=TRUE, height=4, width=8>>=
par(mfrow=c(1,2), mai=c(0.8,0.8,0.35,0.3))
plot(X, pch=16, cex=0.5, main="Two Circles")
plot(maxKDE, main="Max Persistence - KDE")
@
\end{center}
\caption{Max Persistence Method for the selection of smoothing parameters. For each value of the smoothing parameter we display the persistence of the corresponding homological features, along with a (pink) confidence band that separates the statistically significant features from the topological noise.}
\label{fig:eq16e}
\end{figure}
% PLOT END

\newpage
%%%%%%%%%%%%%%%%%
\section{Density Clustering}
%%%%%%%%%%%%%%%%%
\label{sec:density}

The last example of this vignette illustrates the use of the function \code{clusterTree}, which is an implementation of Algorithm 1 in \cite{kent2013debacl}.

First, we briefly describe the task of density clustering; we defer the reader to \cite{kent2013} for a more rigorous and complete description.
Let $f$ be the density of the probability distribution~$P$ generating the observed sample $X=\{x_1, \dots, x_n\} \subset \mathbb{R}^D$. For a threshold value $\lambda >0$, the corresponding super level set of $f$ is $L_f (\lambda) := \text{cl} (\{x \in \mathbb{R}^D : f(x) > \lambda\})$, and its $D$-dimensional subsets are called high-density regions. The high-density clusters of $P$ are the maximal connected subsets of $L_f (\lambda)$.
By considering all the level sets simultaneously (from $\lambda=0$ to $\lambda=\infty$), we can record the evolution and the hierarchy of the high-density clusters of $P$.
This naturally leads to the notion of the cluster density tree of $P$ (see, e.g., \cite{hartigan1981consistency}), defined as the collection of sets $T := \{L_f (\lambda), \lambda \geq 0\}$, which satisfies the tree property: $A,B \in T$ implies that $A \subset B$ or $B \subset A$ or $A \cap B = \emptyset$. We will refer to this construction as the $\lambda$-tree. Alternatively, \cite{kent2013debacl} introduced the \textit{$\kappa$-tree}, which facilitates the interpretation of the tree by precisely encoding the probability content of each tree branch rather than the density level. Cluster trees are particularly useful for high dimensional data, whose spatial organization is difficult to represent.

We illustrate the strategy with a simple example.
First we generate a 2D point cloud from three (not so well) separated clusters (see top left plot of Figure \ref{fig:eq18d}):

<<eq18, echo=TRUE, eval=TRUE>>=
X1=cbind(rnorm(300,1,.8),rnorm(300,5,0.8))
X2=cbind(rnorm(300,3.5,.8),rnorm(300,5,0.8))
X3=cbind(rnorm(300,6,1),rnorm(300,1,1))
XX=rbind(X1,X2,X3)
@

Then we use the function \code{clusterTree} to compute cluster trees using the k Nearest Neighbors density estimator (\code{k=100} nearest neighbors) and the Gaussian kernel density estimator, with smoothing parameter \code{h}.

<<eq18b, echo=TRUE, eval=TRUE>>=
Tree=clusterTree(XX, k=100, density="knn", printStatus=F)
TreeKDE=clusterTree(XX, k=100, h=0.3, density="kde", printStatus=F)
@

Note that, even when \code{kde} is used to estimate the density, we have to provide the option \code{k=100}, so that the algorithm can compute the connected components at each level of the density using a k Nearest Neighbors graph.

The \code{"clusterTree"} objects \code{Tree} and \code{TreeKDE} contain information about both the $\lambda$-tree and the $\kappa$-tree. The function \code{plot} for objects of the class \code{"clusterTree"} produces the plots in Figure \ref{fig:eq18d}.

<<eq18c, echo=TRUE, eval=FALSE>>=
plot(Tree, type="lambda", main="lambda Tree (knn)")
plot(Tree, type="kappa", main="kappa Tree (knn)")
plot(TreeKDE, type="lambda", main="lambda Tree (kde)")
plot(TreeKDE, type="kappa", main="kappa Tree (kde)")
@

% PLOT STARTS
\begin{figure}[h!tb]
\setkeys{Gin}{width=6in, height=4in}
\begin{center}
<<eq18d, echo=FALSE, eval=TRUE,  fig=TRUE, height=4, width=6>>=

par(mfrow=c(2,3))
par(mai=c(0.25,0.35,0.3,0.3))

plot(XX, pch=16, cex=0.6, main="Data")

# plot lambda trees
plot(Tree, type="lambda", main="lambda Tree (knn)")
plot(TreeKDE, type="lambda", main="lambda Tree (kde)")

# plot clusters
plot(XX, pch=19, cex=0.6, main="cluster labels (knn)")
for (i in Tree$id){
	points(matrix(XX[Tree$DataPoints[[i]],],ncol=2), col=i, pch=19, cex=0.6)
}

#plot kappa trees
plot(Tree, type="kappa", main="kappa Tree (knn)")
plot(TreeKDE, type="kappa", main="kappa Tree (kde)")
@
\end{center}
\caption{The lambda trees and kappa trees of the k Nearest Neighbor density estimator and the kernel density estimator.}
\label{fig:eq18d}
\end{figure}
% PLOT END

\newpage

\bibliography{biblio}

\end{document}

### 2014-08-29 R/bootstrapDiagram.R
version 1.2

@@ -1,0 +1,57 @@
+bootstrapDiagram <- 
+function(X, FUN, lim, by, sublevel=TRUE, library="Dionysus", B=30, alpha=0.05, distance="bottleneck", dimension=1, p=1, printProgress=FALSE, ...){
+     
+	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be a matrix of coordinates")
+	if (class(FUN)!="function") stop("FUN should be function")
+	if (2*ncol(X)!=length(lim)) stop("dimension of X does not match with lim")
+	if (!is.vector(by) || length(by)!=1) stop("by should be a positive number")
+	if (!is.logical(sublevel)) stop("sublevel should be logical")
+	if (!is.numeric(B) || length(B)!=1 || B<1) stop("B should be a positive number")
+	if (!is.numeric(alpha)) stop("alpha should be a number")
+	if (!is.numeric(dimension) || length(dimension)!=1) stop("dimension should be a positive integer")
+	if (!is.vector(p) || length(p)!=1 || p < 1) stop("p should be a positive integer")	
+     # if (!is.logical(parallel)) stop("parallel should be logical")
+     if (!is.logical(printProgress)) stop("printProgress should be logical")
+
+
+	if (distance!="wasserstein" && distance!="bottleneck") stop("distance should be a string: either 'bottleneck' or 'wasserstein'")
+
+     X=as.matrix(X)
+     n = nrow(X)
+
+	maxdimension=dimension
+     parallel=FALSE
+     
+     Diag=gridDiag(X=X, FUN=FUN, lim=lim, by=by, maxdimension=maxdimension, sublevel=sublevel, library=library, printProgress=FALSE, diagLimit=NULL, ...)$diagram
+
+     if (parallel) {boostLapply=mclapply
+     	} else boostLapply=lapply
+
+     if (printProgress) cat("Bootstrap: ")
+     
+     if (distance=="wasserstein") {
+	     width=boostLapply(1:B, FUN=function(i){
+	          I = sample(1:n,replace=TRUE,size=n)
+	          Y = as.matrix(X[I,])
+	          Diag1 = gridDiag(X=Y, FUN=FUN, lim=lim, by=by, maxdimension=maxdimension, sublevel=sublevel, library=library, printProgress=FALSE, diagLimit=NULL, ...)$diagram
+	          	width1 = wasserstein(Diag,Diag1, p=p,dimension=dimension)
[…]

### 2014-08-29 R/multipBootstrap.R

@@ -1,11 +1,11 @@
 multipBootstrap <- 
-function(Y, B=30, alpha=0.05, parallel=FALSE, printStatus=FALSE){
+function(Y, B=30, alpha=0.05, parallel=FALSE, printProgress=FALSE){
 
      if (!is.numeric(Y) || !is.matrix(Y)) stop("Y should be a numeric matrix")
 	if (!is.vector(B) || length(B)!=1 || B<1) stop("B should be a positive integer")
 	if (!is.vector(alpha) || length(alpha)!=1) stop("alpha should be a number between 0 and 1")
      if (!is.logical(parallel)) stop("parallel should be logical")
-     if (!is.logical(printStatus)) stop("printStatus should be logical")
+     if (!is.logical(printProgress)) stop("printProgress should be logical")
 
 
 	n = nrow(Y)
@@ -16,17 +16,17 @@ function(Y, B=30, alpha=0.05, parallel=FALSE, printStatus=FALSE){
      if (parallel) {boostLapply=mclapply
      	} else boostLapply=lapply
 
-    if (printStatus) cat("Bootstrap: ")
+    if (printProgress) cat("Bootstrap: ")
 	
 	width=boostLapply(1:B, FUN=function(i){ 
 			xi=rnorm(n)
 			BootLand=xi* (Y- matrix(MeanLand, nrow=n, ncol=Nseq, byrow=T))
 			Gstar=max(abs( apply(BootLand, 2, sum)/sqrt(n)  )) 
-	        if (printStatus) cat(i," ")
+	        if (printProgress) cat(i," ")
 			return(Gstar)
 		}) 
 	
-    if (printStatus) cat("\n")
+    if (printProgress) cat("\n")
 	width=unlist(width)
 	width = quantile(width,1-alpha) /sqrt(n)
 	

### 2014-08-29 R/plot.diagram.R

@@ -1,5 +1,5 @@
 plot.diagram <-
-function(x, diagLim=NULL, dimension=NULL, col=NULL, rotated=FALSE, barcode=FALSE, band=NULL, add=FALSE,...){
+function(x, diagLim=NULL, dimension=NULL, col=NULL, rotated=FALSE, barcode=FALSE, band=NULL, colorBand="pink", colorBorder=NA, add=FALSE,...){
 
     if (class(x)!="diagram" && class(x)!="matrix" && !is.data.frame(x) )  
     	stop("x should be a diagram, or a P by 3 matrix")
@@ -95,13 +95,13 @@ function(x, diagLim=NULL, dimension=NULL, col=NULL, rotated=FALSE, barcode=FALSE
 		if (rotated==TRUE){
 
 			if (add==FALSE) plot(0, 0,type="n", axes=F, xlim=diagLim, ylim=diagLim, xlab=" ", ylab=" ", ...)
-			if (!is.null(band)) 	polygon(c(0,diagLim[2]+1, diagLim[2]+1,0),c(0,0,band,band),col="pink", lwd=0.01, border="white")
+			if (!is.null(band)) 	polygon(c(0,diagLim[2]+1, diagLim[2]+1,0),c(0,0,band,band),col=colorBand, lwd=1.5, border=colorBorder)
 
 			points((x[,2]+x[,3])/2, (x[,3]-x[,2])/2 ,col=col,pch=symb,lwd=2,cex=1)
 		} else{
 
 			if (add==FALSE) plot(0, 0,type="n", axes=F, xlim=diagLim, ylim=diagLim, xlab=" ", ylab=" ", ...)
-			if (!is.null(band)) 	polygon(c(0,diagLim[2]+1,diagLim[2]+1,0),c(0,diagLim[2]+1, diagLim[2]+1+band,band),col="pink", lwd=0.01, border="white")
+			if (!is.null(band)) 	polygon(c(diagLim[1]-1,diagLim[2]+1,diagLim[2]+1,diagLim[1]-1),c(diagLim[1]-1,diagLim[2]+1, diagLim[2]+1+band,diagLim[1]-1+band),col=colorBand, lwd=1.5, border=colorBorder)
 
 
 			points(x[,2],x[,3],pch=symb,lwd=2, cex=1, col=col)

### 2014-08-29 R/ripsArbitDiag.R

@@ -1,15 +1,15 @@
 ripsArbitDiag <-
-function(distX,maxdimension, maxscale, printStatus=FALSE){
+function(distX,maxdimension, maxscale, printProgress=FALSE){
 	
 	if (!is.numeric(distX) && !is.matrix(distX) && !is.data.frame(distX)) stop("distX should be a matrix of distances")
 	if (!is.vector(maxdimension) || length(maxdimension)!=1) stop("maxdimension should be an integer")
 	if (!is.vector(maxscale) || length(maxscale)!=1) stop("maxscale should be a number")
-	if (!is.logical(printStatus)) stop("printStatus should be logical")
+	if (!is.logical(printProgress)) stop("printProgress should be logical")
 
 		
-	write.table(distX,"inputDionysus.txt", row.names=F, col.names=F, sep=" " )	
-	out1=.C("ripsArbit", as.integer(maxdimension+1), as.double(maxscale), as.integer(printStatus), dup=FALSE, package="persistence")
-	Diag=as.matrix(read.table("outputDionysus.txt", sep=""))
+	write.table(distX,"inputTDA.txt", row.names=F, col.names=F, sep=" " )	
+	out1=.C("ripsArbit", as.integer(maxdimension+1), as.double(maxscale), as.integer(printProgress), dup=FALSE, package="TDA")
+	Diag=as.matrix(read.table("outputTDA.txt", sep=""))
 	
 	N=dim(Diag)[1]
 	remove=NULL  # we remove points with lifetime=0
@@ -28,5 +28,5 @@ function(distX,maxdimension, maxscale, printStatus=FALSE){
 	attributes(Diag)$call=match.call()
 	Diag[1,3]=maxscale
 
-	return(Diag)
+	return(list("diagram"=Diag))
 }

### 2014-08-29 src/gudhi/Simplex_tree/Simplex_tree_node_explicit_storage.h

@@ -1,0 +1,100 @@
+ /*    This file is part of the Gudhi Library. The Gudhi library 
+  *    (Geometric Understanding in Higher Dimensions) is a generic C++ 
+  *    library for computational topology.
+  *
+  *    Author(s):       Clément Maria
+  *
+  *    Copyright (C) 2014  INRIA Sophia Antipolis-Méditerranée (France)
+  *
+  *    This program is free software: you can redistribute it and/or modify
+  *    it under the terms of the GNU General Public License as published by
+  *    the Free Software Foundation, either version 3 of the License, or
+  *    (at your option) any later version.
+  *
+  *    This program is distributed in the hope that it will be useful,
+  *    but WITHOUT ANY WARRANTY; without even the implied warranty of
+  *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+  *    GNU General Public License for more details.
+  *
+  *    You should have received a copy of the GNU General Public License
+  *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+  */
+
+#ifndef GUDHI_SIMPLEX_TREE_NODE_EXPLICIT_STORAGE_H
+#define GUDHI_SIMPLEX_TREE_NODE_EXPLICIT_STORAGE_H
+
+#include <vector>
+#include <iostream>
+
+/* \addtogroup simplex_tree
+  * Represents a node of a Simplex_tree.
+  * @{
+  */
+
+/*
+ * \brief Node of a simplex tree with filtration value
+ * and simplex key.
+ *
+ * It stores explicitely its own filtration value and its own Simplex_key.
+ */
+template < class SimplexTree > 
+class Simplex_tree_node_explicit_storage {
+  public:
+//  friend SimplexTree;
+
+  typedef typename SimplexTree::Siblings         Siblings;
+  typedef typename SimplexTree::Filtration_value Filtration_value;
+  typedef typename SimplexTree::Simplex_key      Simplex_key;
+
+  //private:
+  //friend class Simplex_tree; 
+  // Default constructor.
+  Simplex_tree_node_explicit_storage() :
+  children_(NULL),
+  simplex_key_(-1),
+  filtration_(0) {}
+
+  Simplex_tree_node_explicit_storage(Siblings * sib,
+                                     Filtration_value filtration) :
+  children_(sib),
+  simplex_key_(-1),
+  filtration_(filtration) {}
+
+
+  void assign_key(Simplex_key key) { simplex_key_ = key; }
+
+  /*
[…]

### 2014-08-29 src/tdautils/kernelUtils.h

@@ -1,6 +1,3 @@
-// Fabrizio Lecci
-
-#include <iostream>
 #include <cmath>
 
 

### 2014-08-29 src/tdautils/phatUtils.h

@@ -1,0 +1,76 @@
+// wrapper algorithm that computes the persistence pairs of a given boundary matrix using a specified algorithm
+#include <phat/compute_persistence_pairs.h>
+
+// main data structure (choice affects performance)
+#include <phat/representations/vector_vector.h>
+
+// algorithm (choice affects performance)
+#include <phat/algorithms/standard_reduction.h>
+#include <phat/algorithms/chunk_reduction.h>
+#include <phat/algorithms/row_reduction.h>
+#include <phat/algorithms/twist_reduction.h>
+
+#include <phat/boundary_matrix.h>
+
+#include <limits>
+
+template<typename Flt>
+std::vector< std::vector< std::vector< double > > > computePersistentPairsPhat(Flt f, int maxDimension)
+{
+
+	// If phat is used, convert from Dionysus to phat
+	std::vector<Smplx> simplex_map_inv;
+	phat::boundary_matrix< phat::vector_vector > boundary_matrix;
+
+	std::map<Smplx,phat::index,Smplx::VertexComparison> simplex_map;
+	phat::index size_of_simplex_map=0;
+	simplex_map_inv.resize( f.size() );
+	boundary_matrix.set_num_cols( f.size() );
+	for(Fltr::Index it=f.begin();it!=f.end();it++)
+	{
+		phat::column boundary_indices;
+		const Smplx& c = f.simplex(it);
+		for(Smplx::BoundaryIterator bit = c.boundary_begin(); bit != c.boundary_end(); bit++)
+		{
+			boundary_indices.push_back( simplex_map[*bit] );
+		}
+		std::sort(boundary_indices.begin(),boundary_indices.end());
+		boundary_matrix.set_col( size_of_simplex_map, boundary_indices );
+		phat::dimension dim_of_column = boundary_indices.size()==0 ? 0 : boundary_indices.size()-1;
+		boundary_matrix.set_dim( size_of_simplex_map, dim_of_column );
+		simplex_map_inv[size_of_simplex_map] = c;
+		simplex_map.insert(std::map<Smplx,phat::index>::value_type(c, size_of_simplex_map++));
+	}
+
+	// Compute persistent homology from sorted simplicial complex
+	phat::persistence_pairs pairs;
+
+	// choose an algorithm (choice affects performance) and compute the persistence pair
+	// (modifies boundary_matrix)
[…]

### 2014-08-29 src/tdautils/ripsL2.h

@@ -1,0 +1,24 @@
+#include <topology/rips.h>
+#include <topology/filtration.h>
+#include <topology/static-persistence.h>
+#include <topology/dynamic-persistence.h>
+#include <topology/persistence-diagram.h>
+
+#include <geometry/l2distance.h>
+#include <geometry/distances.h>
+#include <utilities/containers.h>           // for BackInsertFunctor
+#include <utilities/timer.h>
+
+#include <vector>
+
+typedef         PairwiseDistances<PointContainer, L2Distance>           PairDistances;
+typedef         PairDistances::DistanceType                             DistanceType;
+typedef         PairDistances::IndexType                                VertexR;
+typedef         Rips<PairDistances>                                     Generator;
+typedef         Generator::Simplex                                      SmplxR;
+typedef         Filtration<SmplxR>                                       FltrR;
+typedef         StaticPersistence<>                                     PersistenceR;
+//typedef         DynamicPersistenceChains<>                              PersistenceR;
+typedef         PersistenceDiagram<>                                    PDgmR;
+
+

### 2014-08-29 src/topology/persistence-diagram.h

@@ -170,7 +170,7 @@ RealType                bottleneck_distance(const Diagram1& dgm1, const Diagram2
 { return bottleneck_distance(dgm1, dgm2, Linfty<typename Diagram1::Point, typename Diagram2::Point>()); }
 
 template<class Diagram>
-RealType                wasserstein_distance(const Diagram& dgm1, const Diagram& dgm2, unsigned p);
+RealType                wasserstein_distance(const Diagram& dgm1, const Diagram& dgm2, int p);
 
 
 #include "persistence-diagram.hpp"

### 2014-12-22 R/clusterTree.R
version 1.3

@@ -1,7 +1,7 @@
 clusterTree <-
 function(X, k, h=NULL, density="knn", dist="euclidean", d=NULL, Nlambda=100, printProgress=FALSE){
 	
-	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be a matrix of coordinates")
+	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be an n by d matrix of coordinates")
 	if (!is.vector(k) || length(k)!=1) stop("k should be an number")
 	if (!is.null(h) && (!is.vector(h) || length(h)!=1)) stop("h should be a real value")
 	if (!is.null(Nlambda) && (!is.vector(Nlambda) || length(Nlambda)!=1)) stop("Nlambda should be a number")
@@ -23,7 +23,7 @@ function(X, k, h=NULL, density="knn", dist="euclidean", d=NULL, Nlambda=100, pri
 	
 	## start adjacency matrix
 	adjMat=diag(n)
-
+	
 	## Compute density estimator: knn or kde
 	if (density=="knn" && dist=="euclidean"){
 		knnInfo=get.knn(X, k=k, algorithm="kd_tree")
@@ -203,20 +203,57 @@ function(X, k, h=NULL, density="knn", dist="euclidean", d=NULL, Nlambda=100, pri
 			}
 		}
 	}	
-	
+
+
+	#info for alpha Tree
+	ID=1:bb
+	alphaBottom=numeric(bb)
+	alphaTop=numeric(bb)
+	for (i in 1:bb)
+	{
+		alphaBottom[i]=sum(hat.f>bottom[i])/n
+		alphaTop[i]=sum(hat.f>top[i])/n
+	}
+
 	## info for the kappa tree
 	kTree=findKtree(bb, parent, sons, compBranch, n)
-	Ktop=kTree$Ktop
-	Kbottom=kTree$Kbottom
+	kappaTop=kTree$kappaTop
+	kappaBottom=kTree$kappaBottom
+
 	
-	out=list("n"=n, "id"=1:bb, "Xbase"=base, "Ybottom"=bottom, "Ytop"=top, "Kbottom"=Kbottom, "Ktop"=Ktop, "silo"=silo, "sons"=sons, "parent"=parent, "DataPoints"=compBranch, "density"=hat.f)
+	## r Tree
+	if (density!="kde")
+	{
+		rTop=(k/(n*v.d*top))^(1/d)
[…]

### 2014-12-22 src/tdautils/dionysusUtils.h

@@ -1,0 +1,44 @@
+#ifndef __DIONYSUSUTILS_H__
+#define __DIONYSUSUTILS_H__
+
+#include <topology/simplex.h>
+#include <string>
+#include <sstream>
+#include <cstdlib>
+
+std::vector<unsigned int> getVertices(const Simplex<unsigned, double> &smp)
+{
+	std::stringstream sstr;
+	std::vector<unsigned int> vertices(smp.dimension()+1);
+	sstr << smp;
+	std::string vtxStr;
+	std::getline(sstr,vtxStr,'<');
+	unsigned int vtxIdx;
+	for (vtxIdx=0; vtxIdx<smp.dimension();++vtxIdx)
+	{
+		std::getline(sstr,vtxStr,',');
+		vertices[vtxIdx] = (unsigned)std::atoi(vtxStr.c_str());
+	}
+	std::getline(sstr,vtxStr,'>');
+	vertices[vtxIdx] = (unsigned)std::atoi(vtxStr.c_str());
+	return vertices;
+}
+
+unsigned int getLocation(const Simplex<unsigned, double> &smp, const double * const FUNvaluesInput)
+{
+	std::vector< unsigned > vertices;
+	std::vector< unsigned >::const_iterator vertexItr;
+	unsigned int vertex;
+	vertices = getVertices( smp );
+	vertex = *(vertices.begin());
+	for (vertexItr = vertices.begin(); vertexItr != vertices.end(); ++vertexItr)
+	{
+		if (FUNvaluesInput[ *vertexItr ] > FUNvaluesInput[ vertex ])
+		{
+			vertex = *vertexItr;
+		}
+	}
+	return vertex+1;
+}
+
+# endif // __DIONYSUSUTILS_H__

### 2015-09-15 inst/doc/article.Rnw
version 1.4

@@
 
-\author{Brittany T. Fasy \and Fabrizio Lecci \\ \small{In collaboration with the CMU TopStat Group}}
-\Plainauthor{Brittany Terese Fasy, Fabrizio Lecci}
+\author{Brittany T. Fasy, Jisu Kim, Fabrizio Lecci \and Cl\'ement Maria \\ \small{In collaboration with the CMU TopStat Group}}
+\Plainauthor{Brittany Terese Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria}
 
-\title{Introduction to the \pkg{TDA} package}
-\Plaintitle{Introduction to the TDA package}
+\title{Introduction to the \proglang{R} package \pkg{TDA}}
+\Plaintitle{Introduction to the R package TDA}
 \Abstract{
@@
 can be quantified with persistent homology. We provide an \proglang{R} interface
-for the efficient algorithms of the \proglang{Dionysus} C++ library, including a function for 
-the persistent homology of the Rips filtration, and for the persistent homology of sublevel
+for the efficient algorithms of the \proglang{C++} libraries \pkg{GUDHI}, \pkg{Dionysus} and \pkg{PHAT}, 
+including a function for the persistent homology of the Rips filtration, 
+and one for the persistent homology of sublevel
 sets (or superlevel sets) of arbitrary functions evaluated over a grid of points.
@@
 analyzed with functions that implement the methods discussed in \cite{fasy2014statistical},
-\cite{chazal2013stochastic} and \cite{chazal2014robust}.
+\cite{chazal2014stochastic} and \cite{chazal2014robust}.
 The \proglang{R} package \pkg{TDA} also includes the implementation of an algorithm for density 
@@
   Email: \email{brittany.fasy@alumni.duke.edu}\\
+
+  Jisu Kim\\
+  Department of Statistics\\
+  Carnegie Mellon University\\
+  Email: \email{jisuk1@andrew.cmu.edu}\\
+
+  Cl\'ement Maria\\
+  Geometrica Group\\
+  INRIA Sophia Antipolis-M\'editerran\'ee\\
+  Email: \email{clement.maria@inria.fr}\\
 
@@
 <<initial_settings, echo=FALSE>>=
-options(stringsAsFactors=FALSE)
-options(width=65)
-options(prompt="> ")
+options(stringsAsFactors = FALSE)
+options(width = 65)
+options(prompt = "> ")
 @
@@
 
-In Section \ref{sec:distances}, we describe how to compute some widely studied functions that, starting from a point cloud, provide some topological information about the underlying space: the distance function (\code{distFct}), the distance to a measure function (\code{dtm}), the k Nearest Neighbor density estimator (\code{knnDE}), the kernel density estimator (\code{kde}), and the kernel distance (\code{kernelDist}).
-
 One approach to TDA is persistent homology \citep{edelsbrunner2010computational}, a method for studying the homology at multiple scales simultaneously.
@@
 This information is encoded in the persistence diagram, a multiset of points in the plane, each corresponding to the birth and death of a homological feature that existed for some interval of $t$.
-In Section \ref{sec:persistent}, we describe how to compute persistence diagrams of sublevel sets (or superlevel sets) of functions evaluated over a grid of points, using \code{gridDiag}, and how to compute the persistence diagram of the Rips filtration built on top of a point cloud, using \code{ripsDiag}.
-Our implementation relies on \pkg{Dionysus}, a \proglang{C++} library developed by Dmitriy Morozov and freely available at http://mrzv.org/software/dionysus/ .
+
+%There is a growing list of software available to compute persistent homology.
+%\textsf{\textbf{javaPlex}}\footnote{https://code.google.com/p/javaplex/} is a
+%Java software package developed by the Computational Topology workgroup at
+%Stanford University;
+%\textsf{\textbf{phom}}\footnote{
+%http://cran.r-project.org/web/packages/phom/index.html} is an R package written
+%by Andrew Tausz;
+%\textsf{\textbf{PHAT}}\footnote{https://code.google.com/p/phat/} is a C++
+%library developed at the Institute of Science and Technology Austria;
+%\textsf{\textbf{Dionysus}}\footnote{http://www.mrzv.org/software/dionysus/} is
+%a
+%C++ library written and maintained by Dmitriy Morozov;
+%finally, \textsf{\textbf{GUDHI}}\footnote{https://project.inria.fr/gudhi/} is
+%new born project hosted by INRIA and whose goal is the development and
+%implementation of new algorithms for geometric understanding in high
+%dimensions.
+%Preliminary results show that \textsf{\textbf{GUDHI}} outperforms its major
+%competitors \textsf{\textbf{PHAT}} and \textsf{\textbf{Dionysus}}; see
+%\cite{boissonnat2013compressed}.
+
+This paper is devoted to the presentation of the \proglang{R} package \pkg{TDA},
+which provides a user-friendly interface for the efficient algorithms of the 
+\proglang{C++} libraries \pkg{GUDHI} \citep{gudhi_stpcoh}, \pkg{Dionysus} \citep{dionysus_C}, and \pkg{PHAT} \citep{phat_C}.
+
+In Section \ref{sec:distances}, we describe how to compute some widely studied functions that, starting from a point cloud, provide some topological information about the underlying space: the distance function (\code{distFct}), the distance to a measure function (\code{dtm}), the k Nearest Neighbor density estimator (\code{knnDE}), the kernel density estimator (\code{kde}), and the kernel distance (\code{kernelDist}).
+Section \ref{sec:persistent} is devoted to the computation of persistence diagrams: the function \code{gridDiag} can be used to compute persistent homology of sublevel sets (or superlevel sets) of functions evaluated over a grid of points; the function \code{ripsDiag} returns the persistence diagram of the Rips filtration built on top of a point cloud.
 
 One of the key challenges in persistent homology is to find a way to isolate the points of the persistence diagram representing the topological noise.
-In \cite{fasy2014statistical}, \cite{chazal2013stochastic}, and \cite{chazal2014robust}, we propose several statistical methods to construct confidence sets for persistence diagrams and other summary functions that allow us to separate topological signal from topological noise.
-Some of these methods are implemented and described in Section \ref{sec:persistent}.
+Statistical methods for persistent homology provide an alternative to its exact
+computation. Knowing with high confidence that an approximated persistence
+diagrams is close to the true--computationally infeasible--diagram is often
+enough for practical purposes. 
+\cite{fasy2014statistical}, \cite{chazal2014stochastic}, and \cite{chazal2014robust} propose several statistical methods to construct confidence sets for persistence diagrams and other summary functions that allow us to separate topological signal from topological noise.
+The methods are implemented in the \pkg{TDA} package and described in Section \ref{sec:persistent}.
 
@@
 This method allows us to identify and visualize the spatial organization of the data, without specific knowledge about the data generating mechanism and in particular without any a priori information about the number of clusters.
-In Section \ref{sec:density}, we describe the function \code{clusterTree}, that, given a density estimator, encodes the hierarchy of the connected components of its superlevel sets into a dendrogram, the cluster tree.
-\citep{KpoLux11, kent2013}.
+In Section \ref{sec:density}, we describe the function \code{clusterTree}, that, given a density estimator, encodes the hierarchy of the connected components of its superlevel sets into a dendrogram, the cluster tree \citep{KpoLux11, kent2013}.
 
@@
 distance functions and density estimators over a grid of points.
-The setting is the typical one in TDA: a set of points $X=\{x_1, \dots, x_n\}$ 
-has been sampled from some distribution $P$ supported on a manifold and we 
-are interested in recovering the topological features of the underlying manifold
+The setting is the typical one in TDA: a set of points $X=\{x_1, \dots, x_n\} \subset \mathbb{R}^d$ 
+has been sampled from some distribution $P$ and we 
+are interested in recovering the topological features of the underlying space
 by studying some functions of the data. 
@@
 library("TDA")
-X = circleUnif(400)
+X <- circleUnif(400)
 
-Xlim=c(-1.6, 1.6);   Ylim=c(-1.7, 1.7);   by=0.065
+Xlim <- c(-1.6, 1.6);  Ylim <- c(-1.7, 1.7);  by <- 0.065
 
-Xseq=seq(Xlim[1], Xlim[2], by=by)
-Yseq=seq(Ylim[1], Ylim[2], by=by)
-Grid=expand.grid(Xseq,Yseq)
+Xseq <- seq(Xlim[1], Xlim[2], by = by)
+Yseq <- seq(Ylim[1], Ylim[2], by = by)
+Grid <- expand.grid(Xseq, Yseq)
 @
@@
 The \pkg{TDA} package provides implementations of the following functions:
+
 \begin{itemize}
@@
 and is computed for each point of the \code{Grid} with the following code:
+
 <<eq2, echo=TRUE>>=
-distance= distFct(X=X, Grid=Grid)
+distance <- distFct(X = X, Grid = Grid)
 @
@@
 where
-$G_y(t) = P( \Vert X-y \Vert \leq t)$ and $0<m_0<1$ is a smoothing parameter. For more details
+$G_y(t) = P( \Vert X-y \Vert \leq t)$ and $0<m_0<1$ is a smoothing parameter. 
+The DTM can be seen as a smoothed version of the distance function.
+For more details
 see \cite{chazal2011geometric}.\\
@@
 The DTM is computed for each point of the \code{Grid} with the following code:
+
 <<eq3, echo=TRUE>>=
-m0=0.1
-DTM=dtm(X=X, Grid=Grid, m0=m0)
+m0 <- 0.1
+DTM <- dtm(X = X, Grid = Grid, m0 = m0)
 @
-\item The $k$ Nearest Neighbor density estimator, for each $y \in \mathbb{R}^d$, is defined by
+\item The $k$ Nearest Neighbor density estimator, for each $y \in \mathbb{R}^d$, is defined as
 $$
@@
 It is computed for each point of the \code{Grid} with the following code:
+
 <<eq4, echo=TRUE>>=
-k=60
-kNN=knnDE(X=X, Grid=Grid, k=k)
+k <- 60
+kNN <- knnDE(X = X, Grid = Grid, k = k)
 @
-\item The Gaussian Kernel Density Estimator (KDE), for each $y \in \mathbb{R}^d$, is defined by
+\item The Gaussian Kernel Density Estimator (KDE), for each $y \in \mathbb{R}^d$, is defined as
 $$
@@
 It is computed for each point of the \code{Grid} with the following code:
+
 <<eq5, echo=TRUE>>=
-h=0.3
-KDE= kde(X=X, Grid=Grid, h=h)
+h <- 0.3
+KDE <- kde(X = X, Grid = Grid, h = h)
 @
-\item The Kernel distance estimator, for each $y \in \mathbb{R}^d$, is defined by
+\item The Kernel distance estimator, for each $y \in \mathbb{R}^d$, is defined as
 $$
@@
 The Kernel distance is computed for each point of the \code{Grid} with the following code:
+
 <<eq6, echo=TRUE>>=
-h=0.3
-Kdist= kernelDist(X=X, Grid=Grid, h=h)
+h <- 0.3
+Kdist <- kernelDist(X = X, Grid = Grid, h = h)
 @
@@
 <<eq7, echo=TRUE, eval=FALSE, height=6, width=6>>=
-persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)),
-xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, 
-main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)
+persp(Xseq, Yseq,
+      matrix(KDE, ncol = length(Yseq), nrow = length(Xseq)), xlab = "",
+      ylab = "", zlab = "", theta = -20, phi = 35, ltheta = 50,
+      col = 2, border = NA, main = "KDE", d = 0.5, scale = FALSE,
+      expand = 3, shade = 0.9)
 @
@@
 <<eq8, echo=FALSE, eval=TRUE, fig=TRUE, height=5, width=6.5>>=
-par(mfrow=c(2,3))
-par(mai=c(0.6,0.25,0.3,0.1))
-	plot(X, pch=16, cex=0.6, xlab="", ylab="", main="Sample X")
+par(mfrow = c(2, 3))
+par(mai = c(0.6, 0.25, 0.3, 0.1))
+plot(X, pch = 16, cex = 0.6, xlab = "", ylab = "", main = "Sample X")
 
-par(mai=c(0.1,0.15,0.1,0.1))
-	persp(Xseq,Yseq,matrix(distance,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="Distance Function", d=0.5, scale=FALSE, expand=1, shade=0.9)
-
-	persp(Xseq,Yseq,matrix(DTM,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="DTM", d=0.5, scale=FALSE, expand=1, shade=0.9)
-
-	persp(Xseq,Yseq,matrix(kNN,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="kNN", d=0.5, scale=FALSE, expand=3, shade=0.9)
-
-	persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)
-
-	persp(Xseq,Yseq,matrix(Kdist,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="Kernel Distance", d=0.5, scale=FALSE, expand=5, shade=0.9)
+par(mai = c(0.1, 0.15, 0.1, 0.1))
+persp(Xseq, Yseq,
+      matrix(distance, ncol = length(Yseq), nrow = length(Xseq)),
+      xlab = "", ylab = "", zlab = "", theta = -20, phi = 35,
+      ltheta = 50, col = 2, border = NA, main = "Distance Function",
+      d = 0.5, scale = FALSE, expand = 1, shade = 0.9)
+persp(Xseq, Yseq,
+      matrix(DTM, ncol = length(Yseq), nrow = length(Xseq)), xlab = "",
+      ylab = "", zlab = "", theta = -20, phi = 35, ltheta = 50,
+      col = 2, border = NA, main = "DTM", d = 0.5, scale = FALSE,
+      expand = 1, shade = 0.9)
+persp(Xseq, Yseq,
+      matrix(kNN, ncol = length(Yseq), nrow = length(Xseq)), xlab = "",
+      ylab = "", zlab = "", theta = -20, phi = 35, ltheta = 50,
+      col = 2, border = NA, main = "kNN", d = 0.5, scale = FALSE,
+      expand = 3, shade = 0.9)
+persp(Xseq, Yseq,
+      matrix(KDE, ncol = length(Yseq), nrow = length(Xseq)), xlab = "",
+      ylab = "", zlab = "", theta = -20, phi = 35, ltheta = 50,
+      col = 2, border = NA, main = "KDE", d = 0.5, scale = FALSE,
+      expand = 3, shade = 0.9)
+persp(Xseq, Yseq,
+      matrix(Kdist, ncol = length(Yseq), nrow = length(Xseq)),
+      xlab = "", ylab = "", zlab = "", theta = -20, phi = 35,
+      ltheta = 50, col = 2, border = NA, main = "Kernel Distance",
+      d = 0.5, scale = FALSE, expand = 5, shade = 0.9)
 @
@@
 
+\newpage
+
 \subsection{Bootstrap Confidence Bands}
@@
 We can construct a $(1-\alpha)$ confidence band for a function using the bootstrap algorithm, which we briefly describe using the kernel density estimator:
+
+\vspace{0.2cm}
+
 \begin{enumerate}
@@
 
+\vspace{0.2cm}
+
 \cite{fasy2014statistical} and \cite{chazal2014robust} prove the validity of the bootstrap algorithm for kernel density estimators, distance to measure, and kernel distance, and use it in the framework of persistent homology.
@@
 <<eq9, echo=TRUE, eval=TRUE>>=
-band=bootstrapBand(X=X, FUN=kde, Grid=Grid, B=100, 
-                   parallel=F, alpha=0.1, h=h)
+band <- bootstrapBand(X = X, FUN = kde, Grid = Grid, B = 100,
+                      parallel = FALSE, alpha = 0.1, h = h)
 @
@@
 <<eq10, echo=FALSE, eval=TRUE, fig=TRUE, width=8, height=4>>=
-posYgrid=which(Grid[,2]>0)
-posYseq=which(Yseq>0)
+posYgrid <- which(Grid[, 2] > 0)
+posYseq <- which(Yseq > 0)
 
-par(mfrow=c(1,2))
-par(mai=c(0.5,0.25,0.3,0.1))
+par(mfrow = c(1, 2))
+par(mai = c(0.5, 0.25, 0.3, 0.1))
 
-persp(Xseq,Yseq,matrix(band$band[,1],ncol=length(Yseq), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = range(Yseq),
-xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)
+persp(Xseq, Yseq, matrix(band[["band"]][, 1], ncol = length(Yseq),
+      nrow = length(Xseq)),
+      zlim = c(0, max(band[["band"]][posYgrid, 2])),
+      ylim = range(Yseq), xlab = "", ylab = "", zlab="",theta = 0,
+      phi = 25, ltheta = 50, col = "pink", border = NA, d = 0.5,
+      scale = FALSE, expand = 3, shade = 0.9, box = FALSE)
 
-persp(Xseq,Yseq[posYseq],matrix(band$band[posYgrid,1],ncol=length(Yseq[posYseq]), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),
-xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)
+persp(Xseq, Yseq[posYseq], matrix(band[["band"]][posYgrid, 1],
+      ncol = length(Yseq[posYseq]), nrow = length(Xseq)),
+      zlim = c(0, max(band[["band"]][posYgrid, 2])),
+      ylim = c(-0.5, 1.7), xlab = "", ylab = "", zlab = "", theta = 0,
+      phi = 25, ltheta = 50, col = "pink", border = NA, d = 0.5,
+      scale = FALSE, expand = 3, shade = 0.9, box = FALSE)
 
@@
 
-persp(Xseq,Yseq[posYseq],matrix(band$fun[posYgrid],ncol=length(Yseq[posYseq]), nrow=length(Xseq)),zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),,
-xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="red", border=NA, d=0.5, scale=FALSE, expand=3, shade=0.9,box=F)
+persp(Xseq, Yseq[posYseq], matrix(band[["fun"]][posYgrid],
+      ncol = length(Yseq[posYseq]), nrow = length(Xseq)),
+      zlim = c(0, max(band[["band"]][posYgrid, 2])),
+      ylim = c(-0.5, 1.7), xlab = "", ylab = "", zlab = "", theta = 0,
+      phi = 25, ltheta = 50, col = "red", border = NA, d = 0.5,
+      scale = FALSE, expand = 3, shade = 0.9, box = FALSE)
 
@@
 
-persp(Xseq,Yseq[posYseq],matrix(band$band[posYgrid,2],ncol=length(Yseq[posYseq]), nrow=length(Xseq)), zlim=c(0,max(band$band[posYgrid,2])), ylim = c(-0.5,1.7),
-xlab="",ylab="",zlab="",theta=0,phi=25, ltheta=50, col="pink", border=NA, 
-main="", d=0.5, scale=FALSE, expand=3, shade=0.9, box=F)
+persp(Xseq, Yseq[posYseq], matrix(band[["band"]][posYgrid, 2],
+      ncol = length(Yseq[posYseq]), nrow = length(Xseq)),
+      zlim = c(0, max(band[["band"]][posYgrid, 2])),
+      ylim = c(-0.5, 1.7), xlab = "", ylab = "", zlab = "", theta = 0,
+      phi = 25, ltheta = 50, col = "pink", border = NA, main = "",
+      d = 0.5, scale = FALSE, expand = 3, shade = 0.9, box = FALSE)
 @
@@
 % PLOT END
+
+\newpage
 
@@
 The function \code{gridDiag} evaluates a given real valued function over a triangulated grid, constructs a filtration of simplices using the values of the function, and computes the persistent homology of the filtration.
-\code{gridDiag} currently supports grids in dimension 1,2, and 3.
-Future releases of the \pkg{TDA} package will extend its functionalities.
+From version 1.2, \code{gridDiag} works in arbitrary dimension.
+The core of the function is written in \proglang{C++} and the user can choose to compute persistence diagrams using either the \pkg{Dionysus} library or the \pkg{PHAT} library.
 
-The following code computes the persistent homology of the superlevel sets (\code{sublevel=FALSE}) of the kernel density estimator (\code{FUN=kde}, \code{h=0.3}) using the point cloud stored in the matrix~\code{X} from the previous example.
-The same code would work for the other functions defined in Section \ref{sec:distances} (it is sufficient to replace \code{kde} and its smoothing parameter \code{h} with another function and its parameters).
+The following code computes the persistent homology of the superlevel sets \linebreak (\code{sublevel=FALSE}) of the kernel density estimator (\code{FUN=kde}, \code{h=0.3}) using the point cloud stored in the matrix~\code{X} from the previous example.
+The same code would work for the other functions defined in Section \ref{sec:distances} (it is sufficient to replace \code{kde} and its smoothing parameter \code{h} with another function and the corresponding parameter).
 The function \code{gridDiag} returns an object of the class \code{"diagram"}.  
-The other inputs are the features of the grid over which the \code{kde} is evaluated (\code{Xlim}, \code{Ylim} and \code{by}), the smoothing parameter \code{h}, and a logical variable that indicates whether a progress bar should be printed (\code{printStatus}).
+The other inputs are the features of the grid over which the \code{kde} is evaluated (\code{lim} and \code{by}), the smoothing parameter \code{h}, and a logical variable that indicates whether a progress bar should be printed (\code{printProgress}).
 
 <<eq11, echo=TRUE, eval=TRUE>>=
-Diag = gridDiag( X=X, FUN=kde, h=0.3, Xlim=Xlim, Ylim=Ylim, by=by, 
-                 sublevel=F, printStatus=F ) 
+Diag <- gridDiag(X = X, FUN = kde, h = 0.3, lim = cbind(Xlim, Ylim),
+                 by = by, sublevel = FALSE, library = "Dionysus",
+                 printProgress = FALSE)
 @
 
-We plot the data and the the diagram, using the function \code{plot}, implemented as a standard \code{S3} method for objects of the class \code{"diagram"}.
+We plot the data and the diagram, using the function \code{plot}, implemented as a standard \code{S3} method for objects of the class \code{"diagram"}.
 The following command produces the third plot in Figure \ref{fig:eq11b}.
@@
 <<eq11b, echo=TRUE, eval=FALSE>>=
-plot(Diag, band=2*band$width, main="KDE Diagram")
+plot(Diag[["diagram"]], band = 2 * band[["width"]],
+     main = "KDE Diagram")
 @
@@
 <<eq11b, echo=FALSE, eval=TRUE, fig=TRUE, height=2, width=5.5>>=
-par(mfrow=c(1,3))
-par(mai=c(0.3,0.25,0.3,0.1))
-
-plot(X, pch=16, cex=0.6, xlab="", ylab="",main="Sample X")
-
-persp(Xseq,Yseq,matrix(KDE,ncol=length(Yseq), nrow=length(Xseq)), xlab="",ylab="",zlab="",theta=-20,phi=35, ltheta=50, col=2, border=NA, main="KDE", d=0.5, scale=FALSE, expand=3, shade=0.9)
-
-par(mai=c(0.55,0.6,0.3,0.1))
-
-plot(Diag, band=2*band$width, main="KDE Diagram")
+par(mfrow = c(1, 3))
+par(mai = c(0.3, 0.25, 0.3, 0.1))
+plot(X, pch = 16, cex = 0.6, xlab = "", ylab = "", main = "Sample X")
+persp(Xseq, Yseq,
+      matrix(KDE, ncol = length(Yseq), nrow = length(Xseq)),
+      xlab = "", ylab = "", zlab = "", theta = -20, phi = 35,
+      ltheta = 50, col = 2, border = NA, main = "KDE", d = 0.5,
+      scale = FALSE, expand = 3, shade = 0.9)
+par(mai = c(0.55, 0.6, 0.3, 0.1))
+plot(Diag[["diagram"]], band = 2 * band[["width"]],
+     main = "KDE Diagram")
 @
@@
 The features above the band can be interpreted as representing significant homological features, while points in the band are not significantly different from noise.
-The validity of the bootstrap confidence band for  persistence diagrams of KDE, DTM, and Kernel Distance derive from the \textit{Stability Theorem} \citep{chazal2012structure} and is discussed in detail in \cite{fasy2014statistical} and \cite{chazal2014robust}.\\
+The validity of the bootstrap confidence band for  persistence diagrams of KDE, DTM, and Kernel Distance derive from the \textit{Stability Theorem} \citep{chazal2012structure} and is discussed in detail in \cite{fasy2014statistical} and \cite{chazal2014robust}.
 
-The function \code{plot} for the class \code{"diagram"} provide the options of rotating the diagram (\code{rotated=TRUE}), drawing the barcode in place of the diagram (\code{barcode=TRUE}), as well as other standard graphical options:
+The function \code{plot} for the class \code{"diagram"} provide the options of rotating the diagram (\code{rotated=TRUE}), drawing the barcode in place of the diagram (\code{barcode=TRUE}), as well as other standard graphical options.
+See Figure \ref{fig:eq11c}.
 
@@
 <<eq11c, echo=TRUE, eval=TRUE,  fig=TRUE, width=6, height=3>>=
-par(mfrow=c(1,2), mai=c(0.8,0.8,0.3,0.1))
-plot(Diag, rotated=T, band=band$width, main="Rotated Diagram")
-plot(Diag, barcode=T,  main="Barcode")
+par(mfrow = c(1, 2), mai = c(0.8, 0.8, 0.3, 0.1))
+plot(Diag[["diagram"]], rotated = TRUE, band = band[["width"]],
+     main = "Rotated Diagram")
+plot(Diag[["diagram"]], barcode = TRUE,  main = "Barcode")
 @
@@
 The {\em Vietoris-Rips~complex} $R(X,\varepsilon)$
-consists of simplices with vertices in $X$ and diameter
+consists of simplices with vertices in \linebreak $X=\{x_1, \dots, x_n\} \subset \mathbb{R}^d$ and diameter
 at most $\varepsilon$.  In other words,
@@
 The \code{ripsDiag} function computes the persistence diagram of the Rips filtration built on top of a point cloud.
-This function is an R wrapper of the function ``rips-pairwise" of the \pkg{Dionysus} \proglang{C++}  library.
-
-The following code generates 60 points from two circles:
+The user can choose to compute the Rips persistence diagram using either the \proglang{C++}  library \pkg{GUDHI}, or \pkg{Dionysus}.
+\\The following code generates 60 points from two circles:
 
 <<eq12, echo=TRUE, eval=TRUE>>=
-Circle1 = circleUnif(60)
-Circle2 = circleUnif(60, r=2) +3
-Circles=rbind(Circle1,Circle2)
+Circle1 <- circleUnif(60)
+Circle2 <- circleUnif(60, r = 2) + 3
+Circles <- rbind(Circle1, Circle2)
 @
 
-We specify the limit of the Rips filtration and the max dimension of the homological features we are interested in (0 for components, 1 for loops, 2 for voids...):
+\noindent We specify the limit of the Rips filtration and the max dimension of the homological features we are interested in (0 for components, 1 for loops, 2 for voids, etc.):
+
 <<eq12b, echo=TRUE, eval=TRUE>>=
-maxscale=5           # limit of the filtration
-maxdimension=1       # components and loops
+maxscale <- 5        # limit of the filtration
+maxdimension <- 1    # components and loops
 @
 
-and we generate the persistence diagram:
+\noindent and we generate the persistence diagram:
+
 <<eq12c, echo=TRUE, eval=TRUE>>=
-Diag=ripsDiag(X=Circles, maxdimension, maxscale, printStatus=F)
+Diag <- ripsDiag(X = Circles, maxdimension, maxscale,
+                 library = "GUDHI", printProgress = FALSE)
 @
 Alternatively, using the option \code{dist="arbitrary"} in \code{ripsDiag()}, the input \code{X} can be an $n \times n$ matrix of distances.
-This option is useful when the user wants to consider a Rips filtration constructed using an arbitrary distance.\\
+This option is useful when the user wants to consider a Rips filtration constructed using an arbitrary distance and is currently only available for the option \code{library="Dionysus"}.\\
 
@@
 <<eq12d, echo=TRUE, eval=TRUE,  fig=TRUE, height=2.8, width=5.6>>=
-par(mfrow=c(1,2), mai=c(0.8,0.8,0.3,0.3))
-plot(Circles, pch=16, xlab="",ylab="")
-plot(Diag)
+par(mfrow = c(1, 2), mai=c(0.8, 0.8, 0.3, 0.3))
+plot(Circles, pch = 16, xlab = "",ylab = "")
+plot(Diag[["diagram"]])
 @
@@
 Standard metrics for measuring the distance between two persistence diagrams are the bottleneck distance and the $p$th Wasserstein distance \citep{edelsbrunner2010computational}. 
-The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \pkg{Dionysus} \proglang{C++}  library .
+The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \proglang{C++}  library \pkg{Dionysus}.
 
 We generate two persistence diagrams of the Rips filtrations built on top of the two (separate) circles of the previous example,
+
 <<eq13, echo=TRUE, eval=TRUE>>=
-Diag1=ripsDiag(Circle1,maxdimension=1, maxscale=5, printStatus=F)
-Diag2=ripsDiag(Circle2,maxdimension=1, maxscale=5, printStatus=F)
+Diag1 <- ripsDiag(Circle1, maxdimension = 1, maxscale = 5)
+Diag2 <- ripsDiag(Circle2, maxdimension = 1, maxscale = 5)
 @
@@
 In the following code, the option \code{dimension=1} specifies that the distances between diagrams are computed using only one dimensional features (loops).
+
 <<eq13b, echo=TRUE, eval=TRUE>>=
-print( bottleneck(Diag1, Diag2, dimension=1) )
-print( wasserstein(Diag1, Diag2, p=2, dimension=1) )
+print(bottleneck(Diag1[["diagram"]], Diag2[["diagram"]],
+                 dimension = 1))
+print(wasserstein(Diag1[["diagram"]], Diag2[["diagram"]], p = 2,
+                  dimension = 1))
 @
@@
 Persistence landscapes and silhouettes are real-valued functions that further summarize the information contained in a persistence diagram.
-They have been introduced and studied in \cite{bubenik2012statistical}, \cite{chazal2013stochastic}, and \cite{chazal2014subsampling}.
+They have been introduced and studied in \cite{bubenik2012statistical}, \cite{chazal2014stochastic}, and \cite{chazal2014subsampling}.
 We briefly introduce the two functions.
@@
 an arrangement of piecewise linear curves by overlaying the graphs of
-the functions~\mbox{$\{ \Lambda_p \}_{p}$}; see Figure \ref{fig:eq14a}.
+the functions~\mbox{$\{ \Lambda_p \}_{p}$}; see Figure \ref{fig:eq14a} (left).
 The persistence landscape of $D$ is a  summary of this arrangement.
@@
 where $k$max is the $k$th largest value in the set; in particular,
-$1$max is the usual maximum~function.
+$1$max is the usual maximum~function. see Figure \ref{fig:eq14a} (middle).
 
 {\bf Silhouette.}
-Consider a persistence diagram with $N$ off diagonal points.
+Consider a persistence diagram with $N$ off diagonal points $\{(b_j, d_j) \}_{j=1}^N$.
 For every $0 < p < \infty$
@@
 <<eq14a, echo=FALSE, eval=TRUE, fig=TRUE, height=2.5, width=7.5>>=
-PlotTriangles = function(left,right){
-   n = length(left)
-   X = (left+right)/2
-   Y = (right-left)/2
-   xlimit=c( 0, max(X+Y)*1.2 )
-   ylimit=c(0, max(Y)*1.2)
-   plot(X,Y,type="n",xlab="",ylab="",xaxt="n",yaxt="n", xlim=xlimit, ylim=ylimit, frame.plot=F, main="Triangles")
+PlotTriangles <- function(left, right) {
+  n <- length(left)
+  X <- (left + right) / 2
+  Y <- (right - left) / 2
+  xlimit <- c(0, max(X + Y) * 1.2)
+  ylimit <- c(0, max(Y) * 1.2)
+  plot(X, Y, type = "n", xlab = "", ylab = "", xaxt = "n", yaxt = "n",
+       xlim = xlimit, ylim = ylimit, frame.plot = FALSE,
+       main = "Triangles")
   axis(1)
   axis(2)
-	mtext("(Birth+Death)/2", side=1, line=2.5, cex=1)
-	mtext("(Death-Birth)/2", side=2, line=2.5, cex=1)
-	#polygon(c(0,0,ylimit[2]),c(0,ylimit[2],ylimit[2]), col="gray87", border=NA)
-   for (i in 1:n){
-      a <- c(0, left[i], (left[i]+right[i])/2, right[i], 0)
-      b <- c(0, 0, (right[i]-left[i])/2, 0, 0)
-      lines(a,b, lwd=2)
+	mtext("(Birth + Death) / 2", side = 1, line = 2.5, cex = 1)
+	mtext("(Death - Birth) / 2", side = 2, line = 2.5, cex = 1)
+	#polygon(c(0, 0, ylimit[2]), c(0, ylimit[2], ylimit[2]),
+  #        col = "gray87", border = NA)
+  for (i in seq_len(n)) {
+    a <- c(0, left[i], (left[i] + right[i]) / 2, right[i], 0)
+    b <- c(0, 0, (right[i] - left[i]) / 2, 0, 0)
+    lines(a, b, lwd = 2)
   }
-     B = max(X+Y)
-     grid = seq(0,B,length=1000)
-     maxfun = rep(0,length(grid))
-     tmp = matrix(0,n,length(grid))
-     for(i in 1:n){
-          tmp[i,] = pmax(pmin(grid-left[i],right[i]-grid),0)
+  B <- max(X + Y)
+  grid <- seq(0, B, length = 1000)
+  maxfun <- rep(0, length(grid))
+  tmp <- matrix(0, n, length(grid))
+  for(i in seq_len(n)) {
+    tmp[i, ] <- pmax(pmin(grid - left[i], right[i] - grid), 0)
   }
-     land = apply(tmp,2,max)
-     #lines(grid,land,lwd=2, col=4)
-	 points(X,Y,type="p", cex=1, pch=19, col="pink")
-     points(X,Y,type="p")
+  land <- apply(tmp, 2, max)
+  #lines(grid, land, lwd = 2, col = 4)
+	points(X, Y, type = "p", cex = 1, pch = 19, col = "pink")
+  points(X, Y, type = "p")
 }
 
-par(mfrow=c(1,3), mai=c(0.7,0.7,0.3,0.3))
+par(mfrow = c(1, 3), mai = c(0.7, 0.7, 0.3, 0.3))
 
-left = c(0,2,2,3.5,5)
-right = c(2,6,3,5,8)
+left <- c(0, 2, 2, 3.5, 5)
+right <- c(2, 6, 3, 5, 8)
 
- PlotTriangles(left,right)
+PlotTriangles(left, right)
 
-Diag1=cbind(rep(0,length(left)), left, right)
-tseq=seq(min(Diag1[,2:3]), max(Diag1[, 2:3]), length=500)
-Land1=landscape(Diag1, dimension=0, KK=1, tseq )
-Sil1=silhouette(Diag1, p=1, dimension=0, tseq )
+Diag1 <- cbind(rep(0, length(left)), left, right)
+tseq <- seq(min(Diag1[, 2:3]), max(Diag1[, 2:3]), length = 500)
+Land1 <- landscape(Diag1, dimension = 0, KK = 1, tseq)
+Sil1 <- silhouette(Diag1, p = 1, dimension = 0, tseq)
 
-   X = (left+right)/2
-   Y = (right-left)/2
-   xlimit=c( 0, max(X+Y)*1.2 )
-   ylimit=c(0, max(Y)*1.2)
+X <- (left + right) / 2
+Y <- (right - left) / 2
+xlimit <- c(0, max(X + Y) * 1.2)
+ylimit <- c(0, max(Y) * 1.2)
 
-plot(tseq, Land1, col=4, type="l", lwd=2, xlim=xlimit, ylim=ylimit, axes=F, main="1st Landscape", xlab="", ylab="")
+plot(tseq, Land1, col = 4, type = "l", lwd = 2, xlim = xlimit,
+     ylim = ylimit, axes = FALSE, main = "1st Landscape", xlab = "",
+     ylab = "")
 axis(1)
@@
 
-plot(tseq, Sil1, col=4, type="l", lwd=2, xlim=xlimit, ylim=ylimit, axes=F, main="Silhouette p=1", xlab="", ylab="")
+plot(tseq, Sil1, col = 4, type = "l", lwd = 2, xlim = xlimit,
+     ylim = ylimit, axes = FALSE, main = "Silhouette p = 1", xlab = "",
+     ylab = "")
 axis(1)
@@
 <<eq14, echo=TRUE, eval=TRUE>>=
-tseq=seq(0,maxscale, length=1000)   #domain
-Land= landscape(Diag, dimension=1, KK=1, tseq)  
-Sil=silhouette(Diag, p=1,  dimension=1, tseq)  
+tseq <- seq(0, maxscale, length = 1000)   #domain
+Land <- landscape(Diag[["diagram"]], dimension = 1, KK = 1, tseq)
+Sil <- silhouette(Diag[["diagram"]], p = 1, dimension = 1, tseq)
 @
@@
 <<eq14b, echo=FALSE, eval=TRUE,  fig=TRUE, height=2.4, width=8>>=
-par(mfrow=c(1,2), mai=c(0.5,0.45,0.3,0.3))
-plot(tseq, Land, type="l", lwd=3, ylab="", main="1st Landscape, dim=1",asp=1, col=2)
-plot(tseq, Sil, type="l", lwd=3, ylab="", main="Silhouette (p=1), dim=1",asp=1, col=2)
+par(mfrow = c(1, 2), mai = c(0.5, 0.45, 0.3, 0.3))
+plot(tseq, Land, type = "l", lwd = 3, ylab = "",
+     main = "1st Landscape, dim = 1", asp = 1, col = 2)
+plot(tseq, Sil, type = "l", lwd = 3, ylab = "",
+     main = "Silhouette(p = 1), dim = 1", asp = 1, col = 2)
 @
@@
 
-Recent results in \cite{chazal2013stochastic} and \cite{chazal2014subsampling} show how to construct confidence bands for landscapes 
+Recent results in \cite{chazal2014stochastic} and \cite{chazal2014subsampling} show how to construct confidence bands for landscapes 
 and silhouettes, using a bootstrap algorithm (multiplier bootstrap).
@@
 Instead, we draw $n$ subsamples, each of size $m$.
-We compute a diagram and landscape for each subsample yielding landscapes $\lambda_1, \dots, \lambda_n $.
+We compute a diagram and a landscape for each subsample yielding landscapes $\lambda_1, \dots, \lambda_n $.
 (Assuming $m$ is much smaller than $N$, these subsamples are essentially independent and identically distributed.) Then we compute $\frac{1}{n}\sum_i \lambda_i$, an estimate of $\mathbb{E}(\lambda_i)$, which can be regarded as an approximation of $\lambda$.
@@
 <<eq15, echo=TRUE, eval=TRUE>>=
-N=4000
-XX1 = circleUnif(N/2)
-XX2 = circleUnif(N/2, r=2) +3
-X=rbind(XX1,XX2)
+N <- 4000
+XX1 <- circleUnif(N / 2)
+XX2 <- circleUnif(N / 2, r = 2) + 3
+X <- rbind(XX1, XX2)
 @
@@
 <<eq15b, echo=TRUE, eval=TRUE>>=
-m=80        # subsample size
-n=10        # we will compute n landscapes using subsamples of size m
+m <- 80     # subsample size
+n <- 10     # we will compute n landscapes using subsamples of size m
+tseq <- seq(0, maxscale, length = 500)          #domain of landscapes
 
-tseq=seq(0,maxscale, length=500)          #domain of landscapes
-
-Diags=list()                              #here we store n Rips diags
-Lands=matrix(0,nrow=n, ncol=length(tseq)) #here we store n landscapes
+#here we store n Rips diags
+Diags <- list()
+#here we store n landscapes
+Lands <- matrix(0, nrow = n, ncol = length(tseq))
 @
@@
 <<eq15c, echo=TRUE, eval=TRUE>>=
-for (i in 1:n){
-	subX=X[sample(1:N,m),]
-	Diags[[i]]=ripsDiag(subX,maxdimension=1, maxscale=5)
-	Lands[i,]=landscape(Diags[[i]], dimension=1, KK=1, tseq )
+for (i in seq_len(n)) {
+  subX <- X[sample(seq_len(N), m), ]
+  Diags[[i]] <- ripsDiag(subX, maxdimension = 1, maxscale = 5)
+  Lands[i, ] <- landscape(Diags[[i]][["diagram"]], dimension = 1,
+                          KK = 1, tseq)
 }
@@
 <<eq15d, echo=TRUE, eval=TRUE>>=
-bootLand=multipBootstrap(Lands,B=100,alpha=0.05, parallel=F)
+bootLand <- multipBootstrap(Lands, B = 100, alpha = 0.05,
+                            parallel = FALSE)
 @
-
 which is plotted by the following code. See Figure \ref{fig:eq15f}.
@@
 <<eq15e, echo=TRUE, eval=FALSE>>=
-plot(tseq, bootLand$mean, main="Mean Landscape with 95% band")
+plot(tseq, bootLand[["mean"]], main = "Mean Landscape with 95% band")
 polygon(c(tseq, rev(tseq)),
-        c(bootLand$band[,1],rev(bootLand$band[,2])), col="pink")
-lines(tseq, bootLand$mean, lwd=2, col=2)
+        c(bootLand[["band"]][, 1], rev(bootLand[["band"]][, 2])),
+        col = "pink")
+lines(tseq, bootLand[["mean"]], lwd = 2, col = 2)
 @
@@
 <<eq15f, echo=FALSE, eval=TRUE,  fig=TRUE, height=4, width=8>>=
-par(mfrow=c(1,2))
-par(mai=c(0.5,0.45,0.3,0.3))
-
-plot(X, pch=16, cex=0.5, main="Large Sample from Circles")
-
-plot(tseq, bootLand$mean, type="l", lwd=2, xlab=" ", ylab="" ,main="Mean Landscape with 95% band", ylim=c(0,1.2))
-polygon(c(tseq, rev(tseq)), c(bootLand$band[,1],rev(bootLand$band[,2])), col="pink")
-lines(tseq, bootLand$mean, lwd=2, col=2)
+par(mfrow = c(1, 2))
+par(mai = c(0.5, 0.45, 0.3, 0.3))
+plot(X, pch = 16, cex = 0.5, main = "Large Sample from Circles")
+plot(tseq, bootLand[["mean"]], type = "l", lwd = 2, xlab = "",
+     ylab = "", main = "Mean Landscape with 95% band",
+     ylim = c(0, 1.2))
+polygon(c(tseq, rev(tseq)),
+        c(bootLand[["band"]][, 1], rev(bootLand[["band"]][, 2])),
+        col = "pink")
+lines(tseq, bootLand[["mean"]], lwd = 2, col = 2)
 @
@@
 choose the smoothing parameters, for example $h$ for KDE and 
-$m$ for DTM.
+$m_0$ for DTM.
 
@@
 The method is implemented in the function \code{maxPersistence}, as shown in the following toy example.
-First, we sample 1600 point from two circles (plus some clutter noise) and we specify the limits of the grid over which the KDE will be evaluated:
+First, we sample 1600 point from two circles (plus some clutter noise) and we specify the limits of the grid over which the KDE is evaluated:
 
 <<eq16, echo=TRUE, eval=TRUE>>=
-XX1 = circleUnif(600)
-XX2 = circleUnif(1000, r=1.5) +2.5
-noise=cbind(runif(80, -2,5),runif(80, -2,5))
-X=rbind(XX1,XX2, noise)
+XX1 <- circleUnif(600)
+XX2 <- circleUnif(1000, r = 1.5) + 2.5
+noise <- cbind(runif(80, -2, 5), runif(80, -2, 5))
+X <- rbind(XX1, XX2, noise)
 
 # Grid limits
-Xlim=c(-2,5)
-Ylim=c(-2,5)
-by=0.2
+Xlim <- c(-2, 5)
+Ylim <- c(-2, 5)
+by <- 0.2
 @
@@
 <<eq16b, echo=TRUE, eval=TRUE>>=
-parametersKDE=seq(0.1,0.6,by=0.05)
+parametersKDE <- seq(0.1, 0.6, by = 0.05)
 
-B=50        # number of bootstrap iterations. Should be large.
-alpha=0.1   # level of the confidence bands
+B <- 50       # number of bootstrap iterations. Should be large.
+alpha <- 0.1  # level of the confidence bands
 @
@@
 <<eq16c, echo=TRUE, eval=TRUE>>=
-maxKDE=maxPersistence(kde, parametersKDE, X, Xlim, Ylim, Zlim=NA, 
-                      by=by, sublevel=F, B=B, alpha=alpha, 
-                      parallel=TRUE, printProgress=TRUE)
+maxKDE <- maxPersistence(kde, parametersKDE, X,
+              lim = cbind(Xlim, Ylim), by = by, sublevel = FALSE,
+              B = B, alpha = alpha, parallel = TRUE,
+              printProgress = TRUE, bandFUN = "bootstrapBand")
 @
@@
 We can display the values of the parameters that maximize the two criteria:
+
 <<eq16d, echo=TRUE, eval=TRUE>>=
@@
 @
-
-and produce the summary plot of Figure \ref{fig:eq16e}:\\
+and produce the summary plot of Figure \ref{fig:eq16e}.\\
 
@@
 <<eq16e, echo=TRUE, eval=TRUE,  fig=TRUE, height=4, width=8>>=
-par(mfrow=c(1,2), mai=c(0.8,0.8,0.35,0.3))
-plot(X, pch=16, cex=0.5, main="Two Circles")
-plot(maxKDE, main="Max Persistence - KDE")
+par(mfrow = c(1, 2), mai = c(0.8, 0.8, 0.35, 0.3))
+plot(X, pch = 16, cex = 0.5, main = "Two Circles")
+plot(maxKDE, main = "Max Persistence - KDE")
 @
@@
 
-\newpage
 %%%%%%%%%%%%%%%%%
@@
 First, we briefly describe the task of density clustering; we defer the reader to \cite{kent2013} for a more rigorous and complete description.
-Let $f$ be the density of the probability distribution~$P$ generating the observed sample $X=\{x_1, \dots, x_n\} \subset \mathbb{R}^D$.
-For a threshold value $\lambda >0$, the corresponding super level set of $f$ is $L_f (\lambda) := \text{cl} (\{x \in \mathbb{R}^D : f(x) > \lambda\})$, and its $D$-dimensional subsets are called high-density regions.
+Let $f$ be the density of the probability distribution~$P$ generating the observed sample $X=\{x_1, \dots, x_n\} \subset \mathbb{R}^d$.
+For a threshold value $\lambda >0$, the corresponding super level set of $f$ is $L_f (\lambda) := \text{cl} (\{x \in \mathbb{R}^s : f(x) > \lambda\})$, and its $d$-dimensional subsets are called high-density regions.
 The high-density clusters of $P$ are the maximal connected subsets of $L_f (\lambda)$.
@@
 We will refer to this construction as the $\lambda$-tree.
-Alternatively, \cite{kent2013debacl} introduced the \textit{$\kappa$-tree}, which facilitates the interpretation of the tree by precisely encoding the probability content of each tree branch rather than the density level.
+Alternatively, \cite{kent2013debacl} introduced the $\alpha$-tree and $\kappa$-tree, which facilitate the interpretation of the tree by precisely encoding the probability content of each tree branch rather than the density level.
 Cluster trees are particularly useful for high dimensional data, whose spatial organization is difficult to represent.
@@
 <<eq18, echo=TRUE, eval=TRUE>>=
-X1=cbind(rnorm(300,1,.8),rnorm(300,5,0.8))
-X2=cbind(rnorm(300,3.5,.8),rnorm(300,5,0.8))
-X3=cbind(rnorm(300,6,1),rnorm(300,1,1))
-XX=rbind(X1,X2,X3)
+X1 <- cbind(rnorm(300, 1, .8), rnorm(300, 5, 0.8))
+X2 <- cbind(rnorm(300, 3.5, .8), rnorm(300, 5, 0.8))
+X3 <- cbind(rnorm(300, 6, 1), rnorm(300, 1, 1))
+XX <- rbind(X1, X2, X3)
 @
@@
 <<eq18b, echo=TRUE, eval=TRUE>>=
-Tree=clusterTree(XX, k=100, density="knn", printStatus=F)
-TreeKDE=clusterTree(XX, k=100, h=0.3, density="kde", printStatus=F)
+Tree <- clusterTree(XX, k = 100, density = "knn",
+                    printProgress = FALSE)
+TreeKDE <- clusterTree(XX, k = 100, h = 0.3, density = "kde",
+                       printProgress = FALSE)
 @
@@
 
-The \code{"clusterTree"} objects \code{Tree} and \code{TreeKDE} contain information about both the $\lambda$-tree and the $\kappa$-tree.
+The \code{"clusterTree"} objects \code{Tree} and \code{TreeKDE} contain information about the $\lambda$-tree, $\alpha$-tree and $\kappa$-tree.
 The function \code{plot} for objects of the class \code{"clusterTree"} produces the plots in Figure \ref{fig:eq18d}.
@@
 <<eq18c, echo=TRUE, eval=FALSE>>=
-plot(Tree, type="lambda", main="lambda Tree (knn)")
-plot(Tree, type="kappa", main="kappa Tree (knn)")
-plot(TreeKDE, type="lambda", main="lambda Tree (kde)")
-plot(TreeKDE, type="kappa", main="kappa Tree (kde)")
+plot(Tree, type = "lambda", main = "lambda Tree (knn)")
+plot(Tree, type = "kappa", main = "kappa Tree (knn)")
+plot(TreeKDE, type = "lambda", main = "lambda Tree (kde)")
+plot(TreeKDE, type = "kappa", main = "kappa Tree (kde)")
 @
@@
 
-par(mfrow=c(2,3))
-par(mai=c(0.25,0.35,0.3,0.3))
+par(mfrow = c(2,3))
+par(mai = c(0.25,0.35,0.3,0.3))
 
-plot(XX, pch=16, cex=0.6, main="Data")
+plot(XX, pch = 16, cex = 0.6, main = "Data")
 
 # plot lambda trees
-plot(Tree, type="lambda", main="lambda Tree (knn)")
-plot(TreeKDE, type="lambda", main="lambda Tree (kde)")
+plot(Tree, type = "lambda", main = "lambda Tree (knn)")
+plot(TreeKDE, type = "lambda", main = "lambda Tree (kde)")
 
 # plot clusters
-plot(XX, pch=19, cex=0.6, main="cluster labels (knn)")
-for (i in Tree$id){
-	points(matrix(XX[Tree$DataPoints[[i]],],ncol=2), col=i, pch=19, cex=0.6)
+plot(XX, pch = 19, cex = 0.6, main = "cluster labels (knn)")
+for (i in Tree[["id"]]) {
+	points(matrix(XX[Tree[["DataPoints"]][[i]], ], ncol = 2), col = i,
+         pch = 19, cex = 0.6)
 }
@@
 #plot kappa trees
-plot(Tree, type="kappa", main="kappa Tree (knn)")
-plot(TreeKDE, type="kappa", main="kappa Tree (kde)")
+plot(Tree, type = "kappa", main = "kappa Tree (knn)")
+plot(TreeKDE, type = "kappa", main = "kappa Tree (kde)")
 @

### 2015-09-15 R/bootstrapDiagram.R

@@ -1,57 +1,154 @@
 bootstrapDiagram <- 
-function(X, FUN, lim, by, sublevel=TRUE, library="Dionysus", B=30, alpha=0.05, distance="bottleneck", dimension=1, p=1, printProgress=FALSE, ...){
-     
-	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be a matrix of coordinates")
-	if (class(FUN)!="function") stop("FUN should be function")
-	if (2*ncol(X)!=length(lim)) stop("dimension of X does not match with lim")
-	if (!is.vector(by) || length(by)!=1) stop("by should be a positive number")
-	if (!is.logical(sublevel)) stop("sublevel should be logical")
-	if (!is.numeric(B) || length(B)!=1 || B<1) stop("B should be a positive number")
-	if (!is.numeric(alpha)) stop("alpha should be a number")
-	if (!is.numeric(dimension) || length(dimension)!=1) stop("dimension should be a positive integer")
-	if (!is.vector(p) || length(p)!=1 || p < 1) stop("p should be a positive integer")	
-     # if (!is.logical(parallel)) stop("parallel should be logical")
-     if (!is.logical(printProgress)) stop("printProgress should be logical")
+function(X, FUN, lim, by, maxdimension = length(lim) / 2 - 1,
+         sublevel = TRUE, library = "Dionysus", B = 30, alpha = 0.05,
+         distance = "bottleneck", dimension = min(1, maxdimension),
+         p = 1, parallel = FALSE, printProgress = FALSE, weight = NULL, ...) {
 
+  if (!is.numeric(X) && !is.data.frame(X)) {
+    stop("X should be a matrix of coordinates")
+  }
+  if (class(FUN) != "function") {
+    stop("FUN should be function")
+  }
+  if (!is.numeric(lim) || length(lim) %% 2 != 0) {
+    stop("lim should be either a numeric matrix or a numeric vector of even elements")
+  }
+  if (!is.numeric(by) || any(by <= 0)) {
+    stop("by should be positive")
+  }
+  if (2 * NCOL(X) != length(lim)) {
+    stop("dimension of X does not match with lim")
+  }
+  if (length(by) != 1 && length(by) != NCOL(X)) {
+    stop("by should be either a number or a vector of length equals dimension of grid")
+  }
+  if (!is.numeric(maxdimension) ||
+      length(maxdimension) != 1 || maxdimension < 0) {
[…]

### 2015-09-15 R/kernelDist.R

@@ -1,15 +1,26 @@
-kernelDist<-
-function(X, Grid, h, printProgress=FALSE){
+kernelDist <-
+function(X, Grid, h, weight = 1, printProgress = FALSE) {
 
-	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be a matrix of coordinates")
-	if (!is.numeric(Grid) && !is.data.frame(Grid)) stop("Grid should be a matrix of coordinates")
-	if (!is.vector(h) || length(h)!=1) stop("h should be a positive number")
-	if (!is.logical(printProgress)) stop("printProgress should be a logical variable")
-	
-    X=as.matrix(X) 
-	Grid=as.matrix(Grid)
-	out=rep(0, nrow(Grid))
-	outR= .C("kdeDist", as.double(X), as.integer(nrow(X)), as.integer(ncol(X)), as.double(Grid), as.integer(nrow(Grid)), as.double(h), as.integer(printProgress),
-	"out"=as.double(out),dup=FALSE, package="TDA")$out
-	return(outR)
+  if (!is.numeric(X) && !is.data.frame(X)) {
+    stop("X should be a matrix of coordinates")
+  }
+  if (!is.numeric(Grid) && !is.data.frame(Grid)) {
+    stop("Grid should be a matrix of coordinates")
+  }
+  if (NCOL(X) != NCOL(Grid)) {
+    stop("dimensions of X and Grid do not match")
+  }
+  if (!is.numeric(h) || length(h) != 1 || h <= 0) {
+    stop("h should be a positive number")
+  }
+  if (!is.numeric(weight) ||
+      (length(weight) != 1 && length(weight) != NROW(X))) {
+    stop("weight should be either a number or a vector of length equals the number of sample")
+  }
+  if (!is.logical(printProgress)) {
+    stop("printProgress should be a logical variable")
+  }
+  
+  return (KdeDist(X = as.matrix(X), Grid = as.matrix(Grid), h = h,
+      weight = weight, printProgress = printProgress))
 }

### 2015-09-15 R/knnDE.R

@@ -1,14 +1,20 @@
-knnDE<-
+knnDE <-
 function(X, Grid, k){
 
-	if (!is.numeric(X) && !is.data.frame(X)) stop("X should be a matrix of coordinates")
-	if (!is.numeric(Grid) && !is.data.frame(Grid)) stop("Grid should be a matrix of coordinates")
-	if (!is.vector(k) || length(k)!=1) stop("k should be a positive integer")
-	
-	d=ncol(X)
-	n=nrow(X)
-	r.k= apply(knnx.dist(X, Grid, k=k, algorithm="kd_tree"), 1, max)
-	v.d=pi^(d/2) /gamma(d/2+1)
-	out=k/(n*v.d*r.k^d)	
-	return(out)
+  if (!is.numeric(X) && !is.data.frame(X)) {
+    stop("X should be a matrix of coordinates")
+  }
+  if (!is.numeric(Grid) && !is.data.frame(Grid)) {
+    stop("Grid should be a matrix of coordinates")
+  }
+  if (!is.numeric(k) || length(k) != 1 || k < 1) {
+    stop("k should be a positive integer")
+  }
+
+  d <- ncol(X)
+  n <- nrow(X)
+  r.k <- apply(FNN::knnx.dist(X, Grid, k = k, algorithm = "kd_tree"), 1, max)
+  v.d <- pi^(d/2) /gamma(d/2+1)
+  out <- k / (n * v.d * r.k ^ d)  
+  return(out)
 }

### 2015-09-15 R/plot.maxPersistence.R

@@ -1,59 +1,76 @@
-plot.maxPersistence<-
-function(x, features="dimension", colorBand="pink", colorBorder=NA , ...){
-	
-	parameter=x$parameters
-	Pers=x$Persistence
-	eps=x$bands
-	Kseq=length(parameter)
-	maxPers=0
-	for (k in 1:Kseq){
-		maxPers=max(maxPers, Pers[[k]][,2])	
-	}
+plot.maxPersistence <-
+function(x, features = "dimension", colorBand = "pink", colorBorder = NA, ...) {
+  
+  parameter <- x[["parameters"]]
+  Pers <- x[["Persistence"]]
+  eps <- x[["bands"]]
+  Kseq <- length(parameter)
+  maxPers <- 0
+  for (k in seq_len(Kseq)) {
+    maxPers <- max(maxPers, Pers[[k]][, 2]) 
+  }
 
-	if (x$bandFUN=="bootstrapBand" || features=="all")
-	{
-		dimension=NULL	
-	} else {
-		dimension=x$dimension	
-	}
+  if (x[["bandFUN"]] == "bootstrapBand" || features == "all")
+  {
+    dimension <- NULL
+  } else {
+    dimension <- x[["dimension"]]
+  }
 
-	plot(parameter, rep(maxPers, Kseq), type="n", ylim=c(0,1.18*maxPers), xlab="",ylab="", axes=F, ... )
-	title(xlab="parameter", ylab="Persistence", line=2.2)
-	axis(1)
-	axis(2)
-	
-	eps=pmin(eps, rep(1.1*maxPers/2, length(eps)) )
-	polygon(c(parameter, parameter[Kseq:1]), c(2*eps, rep(0,Kseq)), col=colorBand, border=colorBorder, lwd=1.5)
-	
-	for (i in 1:Kseq){	
-		
-		if (x$bandFUN=="bootstrapBand"|| features=="all")
-		{
-			selectDimension=1:nrow(Pers[[i]])
-		} else {
-			selectDimension=which(Pers[[i]][,1]==dimension)
-		}
+  graphics::plot(parameter, rep(maxPers, Kseq), type = "n",
+      ylim = c(0, 1.18 * maxPers), xlab = "", ylab = "", axes=FALSE, ...)
+  graphics::title(xlab = "parameter", ylab = "Persistence", line = 2.2)
+  graphics::axis(1)
+  graphics::axis(2)
+  
+  eps <- pmin(eps, rep(1.1 * maxPers / 2, length(eps)))
[…]

### 2015-09-15 inst/doc/article.R

@@ -3,255 +3,301 @@
 ###################################################
 ### code chunk number 1: initial_settings
 ###################################################
-options(stringsAsFactors=FALSE)
-options(width=65)
-options(prompt="> ")
+options(stringsAsFactors = FALSE)
+options(width = 65)
+options(prompt = "> ")
 
 
 ###################################################
 ### code chunk number 2: eq1
 ###################################################
 library("TDA")
-X = circleUnif(400)
+X <- circleUnif(400)
 
-Xlim=c(-1.6, 1.6);   Ylim=c(-1.7, 1.7);   by=0.065
+Xlim <- c(-1.6, 1.6);  Ylim <- c(-1.7, 1.7);  by <- 0.065
 
-Xseq=seq(Xlim[1], Xlim[2], by=by)
-Yseq=seq(Ylim[1], Ylim[2], by=by)
-Grid=expand.grid(Xseq,Yseq)
+Xseq <- seq(Xlim[1], Xlim[2], by = by)
+Yseq <- seq(Ylim[1], Ylim[2], by = by)
+Grid <- expand.grid(Xseq, Yseq)
 
 
 ###################################################
 ### code chunk number 3: eq2
 ###################################################
-distance= distFct(X=X, Grid=Grid)
+distance <- distFct(X = X, Grid = Grid)
 
 
 ###################################################
 ### code chunk number 4: eq3
 ###################################################
-m0=0.1
-DTM=dtm(X=X, Grid=Grid, m0=m0)
+m0 <- 0.1
+DTM <- dtm(X = X, Grid = Grid, m0 = m0)
 
 
 ###################################################
 ### code chunk number 5: eq4
 ###################################################
-k=60
-kNN=knnDE(X=X, Grid=Grid, k=k)
+k <- 60
+kNN <- knnDE(X = X, Grid = Grid, k = k)
 
 
 ###################################################
 ### code chunk number 6: eq5
 ###################################################
-h=0.3
-KDE= kde(X=X, Grid=Grid, h=h)
+h <- 0.3
+KDE <- kde(X = X, Grid = Grid, h = h)
 
 
 ###################################################
 ### code chunk number 7: eq6
[…]

### 2015-09-15 src/CGAL/Alpha_shape_3.h

@@ -1,0 +1,1955 @@
+// Copyright (c) 1997, 2012  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org).
+// You can redistribute it and/or modify it under the terms of the GNU
+// General Public License as published by the Free Software Foundation,
+// either version 3 of the License, or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the so
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Tran Kai Frank DA <Frank.Da@sophia.inria.fr>
+//                 Andreas Fabri <Andreas.Fabri@geometryfactory.com>
+//                 Mariette Yvinec <Mariette.Yvinec@sophia.inria.fr>
+
+#ifndef CGAL_ALPHA_SHAPE_3_H
+#define CGAL_ALPHA_SHAPE_3_H
+
+#include <CGAL/basic.h>
+
+#include <set>
+#include <map>
+#include <list>
+#include <vector>
+#include <algorithm>
+#include <utility>
+#include <iostream>
+
+#include <CGAL/Triangulation_utils_3.h>
+#include <CGAL/Object.h>
+#include <CGAL/Unique_hash_map.h>
+#include <CGAL/Compact_container.h>
+#include <CGAL/Alpha_shape_vertex_base_3.h>
+#include <CGAL/Alpha_shape_cell_base_3.h>
+#include <CGAL/internal/Lazy_alpha_nt_3.h>
+#include <CGAL/iterator.h>
+#ifdef CGAL_USE_GEOMVIEW
+#include <CGAL/IO/Geomview_stream.h>  // TBC
+#endif
+
+//-------------------------------------------------------------------
+namespace CGAL {
+//-------------------------------------------------------------------
+
+template < class Dt, class ExactAlphaComparisonTag = Tag_false >
+class Alpha_shape_3 : public Dt
+{
+  // DEFINITION The class Alpha_shape_3<Dt> represents the family
+  // of alpha-shapes for a set of points (or a set of weighted points)
+  // for all possible values of alpha. The alphashape is defined  through
+  // the Delaunay tetrahedralization of the points
[…]

### 2015-09-15 src/CGAL/Bbox_2.h

@@ -1,0 +1,233 @@
+// Copyright (c) 1999,2004  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+// Author(s)     : Andreas Fabri
+
+#ifndef CGAL_BBOX_2_H
+#define CGAL_BBOX_2_H
+
+#include <CGAL/basic.h>
+#include <CGAL/kernel_assertions.h>
+#include <CGAL/IO/io.h>
+#include <CGAL/Dimension.h>
+#include <CGAL/array.h>
+
+namespace CGAL {
+
+template < typename T >
+struct Simple_cartesian;
+
+class Bbox_2
+{
+  typedef cpp11::array<double, 4>            BBox_rep_2;
+
+  BBox_rep_2 rep;
+
+public:
+
+  typedef Dimension_tag<2>  Ambient_dimension;
+  typedef Dimension_tag<2>  Feature_dimension;
+
+  typedef Simple_cartesian<double>  R;
+
+             Bbox_2() {}
+
+             Bbox_2(double x_min, double y_min,
+                    double x_max, double y_max)
+		 : rep(CGAL::make_array(x_min, y_min, x_max, y_max)) {}
+
+  inline bool       operator==(const Bbox_2 &b) const;
+  inline bool       operator!=(const Bbox_2 &b) const;
+
+  inline int        dimension() const;
+  inline double     xmin() const;
+  inline double     ymin() const;
[…]

### 2015-09-15 src/CGAL/Bbox_3.h

@@ -1,0 +1,261 @@
+// Copyright (c) 1999,2004  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+// Author(s)     : Andreas Fabri
+
+#ifndef CGAL_BBOX_3_H
+#define CGAL_BBOX_3_H
+
+#include <CGAL/basic.h>
+#include <CGAL/IO/io.h>
+#include <CGAL/Dimension.h>
+#include <CGAL/array.h>
+
+namespace CGAL {
+
+template < typename T >
+struct Simple_cartesian;
+
+class Bbox_3
+{
+  cpp11::array<double, 6>   rep;
+
+public:
+
+  typedef Dimension_tag<3>  Ambient_dimension;
+  typedef Dimension_tag<3>  Feature_dimension;
+
+  typedef Simple_cartesian<double>  R;
+
+        Bbox_3() {}
+
+        Bbox_3(double x_min, double y_min, double z_min,
+               double x_max, double y_max, double z_max)
+	  : rep(CGAL::make_array(x_min, y_min, z_min, x_max, y_max, z_max)) {}
+
+  inline bool operator==(const Bbox_3 &b) const;
+  inline bool operator!=(const Bbox_3 &b) const;
+
+  inline int dimension() const;
+  double  xmin() const;
+  double  ymin() const;
+  double  zmin() const;
+  double  xmax() const;
+  double  ymax() const;
[…]

### 2015-09-15 src/CGAL/Delaunay_triangulation_3.h

@@ -1,0 +1,1927 @@
+// Copyright (c) 1999-2004   INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org).
+// You can redistribute it and/or modify it under the terms of the GNU
+// General Public License as published by the Free Software Foundation,
+// either version 3 of the License, or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+//
+// Author(s)     : Monique Teillaud <Monique.Teillaud@sophia.inria.fr>
+//                 Sylvain Pion
+//                 Andreas Fabri <Andreas.Fabri@sophia.inria.fr>
+//                 Clement Jamin
+
+#ifndef CGAL_DELAUNAY_TRIANGULATION_3_H
+#define CGAL_DELAUNAY_TRIANGULATION_3_H
+
+#include <CGAL/basic.h>
+
+#ifdef CGAL_CONCURRENT_TRIANGULATION_3_PROFILING
+# define CGAL_PROFILE
+# include <CGAL/Profile_counter.h>
+#endif
+
+#include <utility>
+#include <vector>
+
+#include <CGAL/Triangulation_3.h>
+#include <CGAL/iterator.h>
+#include <CGAL/Location_policy.h>
+#ifdef CGAL_TRIANGULATION_3_PROFILING
+# include <CGAL/Mesh_3/Profiling_tools.h>
+#endif
+
+#ifndef CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
+#include <CGAL/Spatial_sort_traits_adapter_3.h>
+#include <CGAL/internal/info_check.h>
+
+#include <boost/tuple/tuple.hpp>
+#include <boost/iterator/zip_iterator.hpp>
+#include <boost/mpl/and.hpp>
+#endif //CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
+
+#ifdef CGAL_LINKED_WITH_TBB
[…]

### 2015-09-15 src/CGAL/IO/Geomview_stream.h

@@ -1,0 +1,596 @@
+// Copyright (c) 1997,1998,1999,2000,2001  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Andreas Fabri, Sylvain Pion
+
+#ifndef CGAL_GEOMVIEW_STREAM_H
+#define CGAL_GEOMVIEW_STREAM_H
+
+#include <CGAL/basic.h>
+#include <CGAL/Bbox_2.h>
+#include <CGAL/Bbox_3.h>
+#include <CGAL/IO/Color.h>
+#include <CGAL/IO/Ostream_iterator.h>
+
+#include <CGAL/export/CGAL.h>
+
+#include <map>
+#include <vector>
+#include <utility>
+#include <string>
+#include <iterator>
+#include <algorithm>
+
+namespace CGAL {
+
+class CGAL_EXPORT Geomview_stream {
+public:
+    Geomview_stream(const Bbox_3 &bbox = Bbox_3(0,0,0, 1,1,1),
+		    const char *machine = NULL,
+		    const char *login = NULL);
+
+    ~Geomview_stream();
+
+    Geomview_stream &operator<<(const Color &c);
+    Geomview_stream &operator<<(const std::string & s);
+    Geomview_stream &operator<<(int i);
+    Geomview_stream &operator<<(unsigned int i);
+    Geomview_stream &operator<<(long i);
+    Geomview_stream &operator<<(unsigned long i);
+    Geomview_stream &operator<<(double d);
+
+    template < class InputIterator >
+    void
+    draw_triangles(InputIterator begin, InputIterator end);
+
[…]

### 2015-09-15 src/CGAL/IO/generic_copy_OFF.h

@@ -1,0 +1,102 @@
+// Copyright (c) 1997  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Lutz Kettner  <kettner@mpi-sb.mpg.de>
+
+#ifndef CGAL_IO_GENERIC_COPY_OFF_H
+#define CGAL_IO_GENERIC_COPY_OFF_H 1
+
+#include <CGAL/basic.h>
+#include <cstddef>
+#include <CGAL/IO/File_header_OFF.h>
+#include <CGAL/IO/File_scanner_OFF.h>
+#include <iostream>
+
+namespace CGAL {
+
+template <class Writer>
+void
+generic_copy_OFF( File_scanner_OFF& scanner,
+                  std::ostream& out,
+                  Writer& writer) {
+    std::istream& in = scanner.in();
+    // scans a polyhedral surface in OFF from `in' and writes it
+    // to `out' in the format provided by `writer'.
+    if ( ! in) {
+        if ( scanner.verbose()) {
+            //std::cerr << " " << std::endl;
+            //std::cerr << "generic_copy_OFF(): "
+            //             "input error: file format is not in OFF."
+            //          << std::endl;
+        }
+        return;
+    }
+
+    // Print header. Number of halfedges is only trusted if it is
+    // a polyhedral surface.
+    writer.write_header( out,
+                         scanner.size_of_vertices(),
+                         scanner.polyhedral_surface() ?
+                             scanner.size_of_halfedges() : 0,
+                         scanner.size_of_facets());
+
+    // read in all vertices
[…]

### 2015-09-15 src/CGAL/IO/io.h

@@ -1,0 +1,311 @@
+// Copyright (c) 1997  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+//
+// Author(s)     : Andreas Fabri
+
+
+#ifndef CGAL_IO_H
+#define CGAL_IO_H
+
+
+#include <cstdio>
+#include <string>
+#include <iostream>
+#include <CGAL/tags.h>
+#include <CGAL/IO/io_tags.h>
+#include <CGAL/IO/Color.h>
+
+
+namespace CGAL {
+
+class IO {
+public:
+    CGAL_EXPORT static int mode;
+    enum Mode {ASCII = 0, PRETTY, BINARY};
+};
+
+template <class T, class F = ::CGAL::Null_tag >
+class Output_rep {
+    const T& t;
+public:
+    //! initialize with a const reference to \a t.
+    Output_rep( const T& tt) : t(tt) {}
+    //! perform the output, calls \c operator\<\< by default.
+    std::ostream& operator()( std::ostream& out) const { return (out << t); }
+};
+
+/*! \relates Output_rep
+    \brief stream output of the \c Output_rep calls its \c operator().
+*/
+template <class T, class F>
+std::ostream&
+operator<<( std::ostream& out, Output_rep<T,F> rep) {
+    return rep( out);
+}
+
+//! generic IO output format manipulator.
+template <class T>
+Output_rep<T>
+oformat( const T& t) { return Output_rep<T>(t); }
+
+//! generic IO output format manipulator with formatting tag.
[…]

### 2015-09-15 src/CGAL/IO/read_off_points.h

@@ -1,0 +1,431 @@
+// Copyright (c) 2007-09  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org).
+// You can redistribute it and/or modify it under the terms of the GNU
+// General Public License as published by the Free Software Foundation,
+// either version 3 of the License, or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+// Author(s) : Pierre Alliez and Laurent Saboret
+
+#ifndef CGAL_READ_OFF_POINTS_H
+#define CGAL_READ_OFF_POINTS_H
+
+#include <CGAL/property_map.h>
+#include <CGAL/value_type_traits.h>
+#include <CGAL/point_set_processing_assertions.h>
+
+#include <boost/version.hpp>
+#if BOOST_VERSION >= 104000
+  #include <boost/property_map/property_map.hpp>
+#else
+  #include <boost/property_map.hpp>
+#endif
+
+#include <iostream>
+#include <sstream>
+#include <string>
+
+namespace CGAL {
+
+
+//===================================================================================
+/// \ingroup PkgPointSetProcessing
+/// Reads points (positions + normals, if available) from a .off ASCII stream.
+/// The function expects for each point a line with the x y z position,
+/// optionally followed by the nx ny nz normal.
+/// Faces are ignored.
+///
+/// @tparam OutputIteratorValueType type of objects that can be put in `OutputIterator`.
+///         It is default to `value_type_traits<OutputIterator>::%type` and can be omitted when the default is fine.
+/// @tparam OutputIterator iterator over output points.
+/// @tparam PointPMap is a model of `WritablePropertyMap` with a value_type = Point_3<Kernel>.
+///        It can be omitted if OutputIterator value_type is convertible to Point_3<Kernel>.
+/// @tparam NormalPMap is a model of `WritablePropertyMap` with a value_type = Vector_3<Kernel>.
+/// @tparam Kernel Geometric traits class.
+///        It can be omitted and deduced automatically from PointPMap value_type.
+///
+/// @return true on success.
+
+// This variant requires all parameters.
[…]

### 2015-09-15 src/CGAL/IO/read_xyz_points.h

@@ -1,0 +1,427 @@
+// Copyright (c) 2007-09  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org).
+// You can redistribute it and/or modify it under the terms of the GNU
+// General Public License as published by the Free Software Foundation,
+// either version 3 of the License, or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+// Author(s) : Pierre Alliez and Laurent Saboret
+
+#ifndef CGAL_READ_XYZ_POINTS_H
+#define CGAL_READ_XYZ_POINTS_H
+
+#include <CGAL/property_map.h>
+#include <CGAL/value_type_traits.h>
+#include <CGAL/point_set_processing_assertions.h>
+
+#include <boost/version.hpp>
+#if BOOST_VERSION >= 104000
+  #include <boost/property_map/property_map.hpp>
+#else
+  #include <boost/property_map.hpp>
+#endif
+
+#include <iostream>
+#include <sstream>
+#include <string>
+
+namespace CGAL {
+
+
+//===================================================================================
+/// \ingroup PkgPointSetProcessing
+/// Reads points (positions + normals, if available) from a .xyz ASCII stream.
+/// The function expects for each point a line with the x y z position,
+/// optionally followed by the nx ny nz normal.
+/// The first line may contain the number of points in the file.
+/// Empty lines and comments starting by # character are allowed.
+///
+/// @tparam OutputIteratorValueType type of objects that can be put in `OutputIterator`.
+///         It is default to `value_type_traits<OutputIterator>::%type` and can be omitted when the default is fine.
+/// @tparam OutputIterator iterator over output points.
+/// @tparam PointPMap is a model of `WritablePropertyMap` with a value_type = Point_3<Kernel>.
+///        It can be omitted if OutputIterator value_type is convertible to Point_3<Kernel>.
+/// @tparam NormalPMap is a model of `WritablePropertyMap` with a value_type = Vector_3<Kernel>.
+/// @tparam Kernel Geometric traits class.
+///        It can be omitted and deduced automatically from PointPMap value_type.
+///
[…]

### 2015-09-15 src/CGAL/Kernel_d/function_objectsCd.h

@@ -1,0 +1,812 @@
+// Copyright (c) 2000,2001  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Michael Seel, Kurt Mehlhorn
+
+#ifndef CGAL_FUNCTION_OBJECTSCD_H
+#define CGAL_FUNCTION_OBJECTSCD_H
+
+#include <CGAL/basic.h>
+#include <CGAL/enum.h>
+#include <CGAL/Referenced_argument.h>
+
+#undef CGAL_KD_TRACE
+#undef CGAL_KD_TRACEN
+#undef CGAL_KD_TRACEV
+#define CGAL_KD_TRACE(t)  std::cerr << t
+#define CGAL_KD_TRACEN(t) std::cerr << t << std::endl
+#define CGAL_KD_TRACEV(t) std::cerr << #t << " = " << (t) << std::endl
+ 
+namespace CGAL {
+
+template <typename K>
+class Compute_coordinateCd {
+  typedef typename K::FT             FT;
+  typedef typename K::Point_d        Point_d;
+  public:
+  typedef FT                         result_type;
+  result_type 
+    operator()(const Point_d& p, int i) const
+  {
+    return p.cartesian(i);
+  }
+};
+
+template <typename K>
+class Point_dimensionCd {
+  typedef typename K::FT             FT;
+  typedef typename K::Point_d        Point_d;
+  public:
+  typedef int                       result_type;
+  result_type 
+    operator()(const Point_d& p) const
+  {
+    return p.dimension();
+  }
+};
+
+template <typename K>
+class Less_coordinateCd {
[…]

### 2015-09-15 src/CGAL/Point_2.h

@@ -1,0 +1,270 @@
+// Copyright (c) 1999  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+//
+// Author(s)     : Andreas Fabri, Stefan Schirra
+
+#ifndef CGAL_POINT_2_H
+#define CGAL_POINT_2_H
+
+#include <CGAL/Origin.h>
+#include <CGAL/Bbox_2.h>
+#include <CGAL/assertions.h>
+#include <boost/type_traits/is_same.hpp>
+#include <CGAL/Kernel/Return_base_tag.h>
+#include <CGAL/representation_tags.h>
+#include <CGAL/Dimension.h>
+
+namespace CGAL {
+
+template <class R_>
+class Point_2 : public R_::Kernel_base::Point_2
+{
+  typedef typename R_::RT                    RT;
+  typedef typename R_::FT                    FT;
+  typedef typename R_::Aff_transformation_2  Aff_transformation_2;
+  typedef typename R_::Kernel_base::Point_2  RPoint_2;
+
+  typedef Point_2                            Self;
+  CGAL_static_assertion((boost::is_same<Self, typename R_::Point_2>::value));
+
+public:
+
+  typedef Dimension_tag<2>  Ambient_dimension;
+  typedef Dimension_tag<0>  Feature_dimension;
+
+  typedef RPoint_2 Rep;
+  typedef typename R_::Cartesian_const_iterator_2 Cartesian_const_iterator;
+
+  const Rep& rep() const
+  {
+    return *this;
+  }
+
+  Rep& rep()
+  {
[…]

### 2015-09-15 src/CGAL/Point_3.h

@@ -1,0 +1,302 @@
+// Copyright (c) 1999  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+//
+// Author(s)     : Andreas Fabri, Stefan Schirra
+
+#ifndef CGAL_POINT_3_H
+#define CGAL_POINT_3_H
+
+#include <CGAL/Origin.h>
+#include <CGAL/representation_tags.h>
+#include <CGAL/assertions.h>
+#include <boost/type_traits/is_same.hpp>
+#include <CGAL/Kernel/Return_base_tag.h>
+#include <CGAL/Bbox_3.h>
+#include <CGAL/Dimension.h>
+
+namespace CGAL {
+
+template <class R_>
+class Point_3 : public R_::Kernel_base::Point_3
+{
+  typedef typename R_::RT                    RT;
+  typedef typename R_::FT                    FT;
+  typedef typename R_::Vector_3              Vector_3;
+  typedef typename R_::Aff_transformation_3  Aff_transformation_3;
+
+  typedef Point_3                            Self;
+  CGAL_static_assertion((boost::is_same<Self, typename R_::Point_3>::value));
+
+public:
+
+  typedef Dimension_tag<3>  Ambient_dimension;
+  typedef Dimension_tag<0>  Feature_dimension;
+
+  typedef typename R_::Kernel_base::Point_3  Rep;
+  typedef typename R_::Cartesian_const_iterator_3 Cartesian_const_iterator;
+
+  const Rep& rep() const
+  {
+    return *this;
+  }
+
+  Rep& rep()
+  {
[…]

### 2015-09-15 src/CGAL/Regular_triangulation_3.h

@@ -1,0 +1,2152 @@
+// Copyright (c) 1999-2004  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org).
+// You can redistribute it and/or modify it under the terms of the GNU
+// General Public License as published by the Free Software Foundation,
+// either version 3 of the License, or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+//
+//
+// Author(s)     : Monique Teillaud <Monique.Teillaud@sophia.inria.fr>
+//                 Sylvain Pion
+//                 Christophe Delage <Christophe.Delage@sophia.inria.fr>
+//                 Clement Jamin
+
+#ifndef CGAL_REGULAR_TRIANGULATION_3_H
+#define CGAL_REGULAR_TRIANGULATION_3_H
+
+#include <CGAL/basic.h>
+
+#include <set>
+#ifdef CGAL_LINKED_WITH_TBB
+# include <tbb/enumerable_thread_specific.h>
+#endif
+
+#include <CGAL/Triangulation_3.h>
+#include <CGAL/Regular_triangulation_cell_base_3.h>
+#include <boost/bind.hpp>
+
+#ifndef CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
+#include <CGAL/Spatial_sort_traits_adapter_3.h>
+#include <CGAL/internal/info_check.h>
+
+#include <boost/iterator/zip_iterator.hpp>
+#include <boost/mpl/and.hpp>
+#endif //CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
+#ifdef CGAL_TRIANGULATION_3_PROFILING
+# include <CGAL/Mesh_3/Profiling_tools.h>
+#endif
+
+#if defined(BOOST_MSVC)
+#  pragma warning(push)
+#  pragma warning(disable:4355) // complaint about using 'this' to
+#endif                          // initialize a member
+
+namespace CGAL {
+
+  /************************************************
+   *
+   * Regular_triangulation_3 class
+   *
+   ************************************************/
+
[…]

### 2015-09-15 src/CGAL/aff_transformation_tags.h

@@ -1,0 +1,47 @@
+// Copyright (c) 1999  
+// Utrecht University (The Netherlands),
+// ETH Zurich (Switzerland),
+// INRIA Sophia-Antipolis (France),
+// Max-Planck-Institute Saarbruecken (Germany),
+// and Tel-Aviv University (Israel).  All rights reserved. 
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Andreas Fabri
+ 
+
+#ifndef CGAL_AFF_TRANSFORMATION_TAGS_H
+#define CGAL_AFF_TRANSFORMATION_TAGS_H
+
+#include <CGAL/basic.h>
+
+namespace CGAL {
+
+class Translation {};
+class Rotation {};
+class Scaling {};
+class Reflection {};
+class Identity_transformation {};
+
+CGAL_EXPORT extern const Translation             TRANSLATION;
+CGAL_EXPORT extern const Rotation                ROTATION;
+CGAL_EXPORT extern const Scaling                 SCALING;
+CGAL_EXPORT extern const Reflection              REFLECTION;
+CGAL_EXPORT extern const Identity_transformation IDENTITY;
+
+} //namespace CGAL
+
+#endif // CGAL_AFF_TRANSFORMATION_TAGS_H

### 2015-09-15 src/CGAL/internal/Projection_traits_3.h

@@ -1,0 +1,636 @@
+// Copyright (c) 1997-2010  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Mariette Yvinec, Sebastien Loriot
+
+#ifndef CGAL_INTERNAL_PROJECTION_TRAITS_3_H
+#define CGAL_INTERNAL_PROJECTION_TRAITS_3_H
+
+#include <CGAL/assertions.h>
+
+#include <CGAL/Point_3.h>
+#include <CGAL/Segment_3.h>
+#include <CGAL/Triangle_3.h>
+
+#include <CGAL/Kernel/global_functions_2.h>
+
+namespace CGAL { 
+
+namespace internal {
+
+//project Point_3 along coordinate dim
+template <class R,int dim>
+struct Projector;
+
+//project onto yz
+template <class R>
+struct Projector<R,0>
+{
+  typedef typename R::Less_y_3                Less_x_2;
+  typedef typename R::Less_z_3                Less_y_2;
+  typedef typename R::Compare_y_3             Compare_x_2;
+  typedef typename R::Compare_z_3             Compare_y_2;
+  typedef typename R::Equal_y_3               Equal_x_2;
+  typedef typename R::Equal_z_3               Equal_y_2;  
+  
+  static typename R::FT x(const typename R::Point_3& p) {return p.y();}
+  static typename R::FT y(const typename R::Point_3& p) {return p.z();}
+  static const int x_index=1;
+  static const int y_index=2;
+};
+//project onto xz
+template <class R>
+struct Projector<R,1>
+{
+  typedef typename R::Less_x_3                Less_x_2;
+  typedef typename R::Less_z_3                Less_y_2;
+  typedef typename R::Compare_x_3             Compare_x_2;
[…]

### 2015-09-15 src/CGAL/internal/Static_filters/Static_filter_error.h

@@ -1,0 +1,169 @@
+// Copyright (c) 1999,2000  INRIA Sophia-Antipolis (France).
+// All rights reserved.
+//
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
+//
+// Licensees holding a valid commercial license may use this file in
+// accordance with the commercial license agreement provided with the software.
+//
+// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
+// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
+//
+// $URL$
+// $Id$
+// 
+//
+// Author(s)     : Sylvain Pion
+
+#ifndef CGAL_INTERNAL_STATIC_FILTER_ERROR_H
+#define CGAL_INTERNAL_STATIC_FILTER_ERROR_H
+
+// This file contains the description of the class Static_filter_error.
+// The goal of this class is to be run by some overloaded predicates,
+// to compute error bound done in these functions.
+// 
+// The original idea is from Olivier Devillers.
+
+// It is still EXPERIMENTAL.
+
+// TODO:
+// - I need to add some missing operators and functions, min/max...
+// - Remove the degree stuff, it's only meant for debug (?).
+// - Add __attribute__((const)) for optimizing ?
+
+#include <CGAL/basic.h>
+#include <CGAL/FPU.h>
+
+namespace CGAL { namespace internal {
+
+struct Static_filter_error
+{
+  typedef Static_filter_error Sfe;
+
+  Static_filter_error () {}
+
+  Static_filter_error (const int &i, const double &e = 0, const int &d = 1)
+      : _b(i), _e(e), _d(d) {}
+
+  Static_filter_error (const double &b, const double &e = 0, const int &d = 1)
+      : _b(b), _e(e), _d(d) {}
+
+  static double ulp ()
+  {
+      FPU_CW_t backup = FPU_get_and_set_cw(CGAL_FE_UPWARD);
+      double e = ulp(1);
+      FPU_set_cw(backup);
+      return e;
+  }
+
+  static double ulp (double d)
+  {
+      // You are supposed to call this function with rounding towards
+      // +infinity, and on a positive number.
[…]

### 2015-09-15 src/gudhi/Simplex_tree.h

@@ -1,805 +1,841 @@
- /*    This file is part of the Gudhi Library. The Gudhi library 
-  *    (Geometric Understanding in Higher Dimensions) is a generic C++ 
-  *    library for computational topology.
-  *
-  *    Author(s):       Clément Maria
-  *
-  *    Copyright (C) 2014  INRIA Sophia Antipolis-Méditerranée (France)
-  *
-  *    This program is free software: you can redistribute it and/or modify
-  *    it under the terms of the GNU General Public License as published by
-  *    the Free Software Foundation, either version 3 of the License, or
-  *    (at your option) any later version.
-  *
-  *    This program is distributed in the hope that it will be useful,
-  *    but WITHOUT ANY WARRANTY; without even the implied warranty of
-  *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-  *    GNU General Public License for more details.
-  *
-  *    You should have received a copy of the GNU General Public License
-  *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
-  */
+/*    This file is part of the Gudhi Library. The Gudhi library
+ *    (Geometric Understanding in Higher Dimensions) is a generic C++
+ *    library for computational topology.
+ *
+ *    Author(s):       Clément Maria
+ *
+ *    Copyright (C) 2014  INRIA Sophia Antipolis-Méditerranée (France)
+ *
+ *    This program is free software: you can redistribute it and/or modify
+ *    it under the terms of the GNU General Public License as published by
+ *    the Free Software Foundation, either version 3 of the License, or
+ *    (at your option) any later version.
+ *
+ *    This program is distributed in the hope that it will be useful,
+ *    but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *    GNU General Public License for more details.
+ *
+ *    You should have received a copy of the GNU General Public License
+ *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
 
-#ifndef GUDHI_SIMPLEX_TREE_H
-#define GUDHI_SIMPLEX_TREE_H
+#ifndef SRC_SIMPLEX_TREE_INCLUDE_GUDHI_SIMPLEX_TREE_H_
+#define SRC_SIMPLEX_TREE_INCLUDE_GUDHI_SIMPLEX_TREE_H_
 
[…]

### 2015-09-15 src/gudhi/Simplex_tree/indexing_tag.h

@@ -1,30 +1,39 @@
- /*    This file is part of the Gudhi Library. The Gudhi library 
-  *    (Geometric Understanding in Higher Dimensions) is a generic C++ 
-  *    library for computational topology.
-  *
-  *    Author(s):       Clément Maria
-  *
-  *    Copyright (C) 2014  INRIA Sophia Antipolis-Méditerranée (France)
-  *
-  *    This program is free software: you can redistribute it and/or modify
-  *    it under the terms of the GNU General Public License as published by
-  *    the Free Software Foundation, either version 3 of the License, or
-  *    (at your option) any later version.
-  *
-  *    This program is distributed in the hope that it will be useful,
-  *    but WITHOUT ANY WARRANTY; without even the implied warranty of
-  *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-  *    GNU General Public License for more details.
-  *
-  *    You should have received a copy of the GNU General Public License
-  *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
-  */
+/*    This file is part of the Gudhi Library. The Gudhi library
+ *    (Geometric Understanding in Higher Dimensions) is a generic C++
+ *    library for computational topology.
+ *
+ *    Author(s):       Clément Maria
+ *
+ *    Copyright (C) 2014  INRIA Sophia Antipolis-Méditerranée (France)
+ *
+ *    This program is free software: you can redistribute it and/or modify
+ *    it under the terms of the GNU General Public License as published by
+ *    the Free Software Foundation, either version 3 of the License, or
+ *    (at your option) any later version.
+ *
+ *    This program is distributed in the hope that it will be useful,
+ *    but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *    GNU General Public License for more details.
+ *
+ *    You should have received a copy of the GNU General Public License
+ *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+
+#ifndef SRC_SIMPLEX_TREE_INCLUDE_GUDHI_SIMPLEX_TREE_INDEXING_TAG_H_
[…]

### 2015-09-15 src/tdautils/gridUtils.h

@@ -12,13 +12,12 @@
 #include <numeric>
 
 
-
 #if 1
-#include <fstream>
 #include <boost/archive/text_oarchive.hpp>
 #include <boost/archive/text_iarchive.hpp>
 #include <boost/serialization/vector.hpp>
 #endif
+
 
 
 typedef         unsigned                                            Vertex;
@@ -36,40 +35,45 @@ typedef         OffsetBeginMap<Fltr, Persistence,
                                Persistence::iterator>               FiltrationPersistenceMap;
 
 
+
 // add a single edge to the filtration
-void addEdge(Fltr& filtr, const double * const fcnvalues, 
-            int vert01, int vert02)
-{
+template <typename RealVector>
+void addEdge(Fltr& filtr, const RealVector& FUNvalues, 
+            int vert01, int vert02) {
      VertexVector vertices(3);
      vertices[0] = vert01;
      vertices[1] = vert02;
      VertexVector::const_iterator bg = vertices.begin();
 
-     double value = std::max(fcnvalues[vert01], fcnvalues[vert02]);
+     double value = std::max(FUNvalues[vert01], FUNvalues[vert02]);
      filtr.push_back(Smplx(bg, bg + 2, value)); 
-         // std::max(fcnvalues.at(vert03),fcnvalues.at(vert04))),
+         // std::max(FUNvalues.at(vert03),FUNvalues.at(vert04))),
 } // end function to add a single edge
 
+
+
 // add a single triangle to the filtration
-void addTri(Fltr& filtr, const double * const fcnvalues, 
-            int vert01, int vert02, int vert03)
-{
+template <typename RealVector>
+void addTri(Fltr& filtr, const RealVector& FUNvalues, 
+            int vert01, int vert02, int vert03) {
      VertexVector vertices(3);
      vertices[0] = vert01;
      vertices[1] = vert02;
      vertices[2] = vert03;
      VertexVector::const_iterator bg = vertices.begin();
 
-     double value = std::max(std::max(fcnvalues[vert01], fcnvalues[vert02]),
-                    fcnvalues[vert03]);
+     double value = std::max(std::max(FUNvalues[vert01], FUNvalues[vert02]),
+                    FUNvalues[vert03]);
[…]

### 2016-06-16 R/dtm.R
version 1.5

@@ -1,5 +1,5 @@
 dtm <-
-function(X, Grid, m0, weight = 1) {
+function(X, Grid, m0, r = 2, weight = 1) {
 
   if (!is.numeric(X) && !is.data.frame(X)) {
     stop("X should be a matrix of coordinates")
@@ -13,6 +13,9 @@ function(X, Grid, m0, weight = 1) {
   if (!is.numeric(m0) || length(m0) != 1 || m0 < 0 || m0 > 1) {
     stop("m0 should be a number between 0 and 1")
   }
+  if (!is.numeric(r) || length(r) != 1 || r < 1) {
+    stop("r should be a number greater than or equal to 1")
+  }
   if (!is.numeric(weight) || 
       (length(weight) != 1 && length(weight) != NROW(X))) {
     stop("weight should be either a number or a vector of length equals the number of sample")
@@ -20,11 +23,12 @@ function(X, Grid, m0, weight = 1) {
 
   # without weight
   if (length(weight) == 1) {
-    X <- as.matrix(X) 
-    k0 <- ceiling(m0 * NROW(X))
-    distances <- FNN::knnx.dist(X, as.matrix(Grid), k = k0,
-        algorithm = c("kd_tree"))
-    return (sqrt(apply(distances^2, 1, sum)/k0))
+    X <- as.matrix(X)
+    weightBound <- m0 * NROW(X)
+    knnDistance <- FNN::knnx.dist(
+		data = X, query = as.matrix(Grid), k = ceiling(weightBound),
+		algorithm = c("kd_tree"))
+    return (Dtm(knnDistance = knnDistance, weightBound = weightBound, r = r))
 
   # with weight
   } else {
@@ -39,10 +43,10 @@ function(X, Grid, m0, weight = 1) {
         break
       }
     }
-    indexDistance <- FNN::get.knnx(X0, as.matrix(Grid), k = k0,
-        algorithm = c("kd_tree"))
-    return (Dtm(knnIndex = indexDistance[["nn.index"]],
-        knnDistance = indexDistance[["nn.dist"]], weight = weight0,
-        weightBound = weightBound))
+    knnDistanceIndex <- FNN::get.knnx(
+	    data = X0, query = as.matrix(Grid), k = k0, algorithm = c("kd_tree"))
+    return (DtmWeight(
+	    knnDistance = knnDistanceIndex[["nn.dist"]], weightBound = weightBound,
+		r = r, knnIndex = knnDistanceIndex[["nn.index"]], weight = weight0))
   }
 }

### 2016-06-16 R/plotRule.R

@@ -19,7 +19,7 @@ function(Tree) {
       
       for (i in uniqueParentsNo0) {
         
-        bros <- Tree[["sons"]][[i]]
+        bros <- Tree[["children"]][[i]]
         newID <- bros
         orderID <- bros[order(Tops[bros], decreasing = TRUE)]
         newID[order(Tops[bros], decreasing=TRUE)] <- bros  
@@ -42,10 +42,10 @@ function(Tree) {
           }
             
           for (j in seq(along = newID)) {
-            if (!is.null(Tree[["sons"]][bros[j]][[1]])){
-              NewTree[["sons"]][[newID[j]]] <- Tree[["sons"]][bros[j]][[1]]
+            if (!is.null(Tree[["children"]][bros[j]][[1]])){
+              NewTree[["children"]][[newID[j]]] <- Tree[["children"]][bros[j]][[1]]
             } else {
-              NewTree[["sons"]][[newID[j]]] <- NA
+              NewTree[["children"]][[newID[j]]] <- NA
             }
           }
           

### 2016-06-16 src/CGAL/IO/io.h

@@ -28,7 +28,9 @@
 
 
 #include <cstdio>
+#include <cctype>
 #include <string>
+#include <locale>
 #include <iostream>
 #include <CGAL/tags.h>
 #include <CGAL/IO/io_tags.h>
@@ -39,7 +41,17 @@ namespace CGAL {
 
 class IO {
 public:
-    CGAL_EXPORT static int mode;
+#ifndef CGAL_HEADER_ONLY
+  CGAL_EXPORT static int mode;
+  static int& get_static_mode()
+  { return IO::mode; }
+#else // CGAL_HEADER_ONLY
+  static int& get_static_mode()
+  {
+    static int mode = std::ios::xalloc();
+    return mode;
+  }
+#endif // CGAL_HEADER_ONLY
     enum Mode {ASCII = 0, PRETTY, BINARY};
 };
 
@@ -90,6 +102,62 @@ public:
     //! perform the input, calls \c operator\>\> by default.
     std::istream& operator()( std::istream& in) const { return (in >> t); }
 };
+
+#if CGAL_FORCE_IFORMAT_DOUBLE || \
+  ( ( _MSC_VER > 1600 ) && (! defined( CGAL_NO_IFORMAT_DOUBLE )) )
+template <>
+class Input_rep<double> {
+    double& t;
+public:
+  //! initialize with a reference to \a t.
+  Input_rep( double& tt) : t(tt) {}
+
+  std::istream& operator()( std::istream& is) const 
+  {
+    typedef std::istream istream;
+    typedef istream::char_type char_type;
+    typedef istream::int_type int_type;
+    typedef istream::traits_type traits_type;
+
+    std::string buffer;
+    buffer.reserve(32);
+
+    char_type c;
+    do {
+      const int_type i = is.get();
+      if(i == traits_type::eof()) {
+	return is;
+      }
+      c = static_cast<char_type>(i);
+    }while (std::isspace(c));
+    if(c == '-'){
+      buffer += '-';
+    } else if(c != '+'){
+      is.unget();
+    }
+    do {
+      const int_type i = is.get();
+      if(i == traits_type::eof()) {
+	is.clear(is.rdstate() & ~std::ios_base::failbit);
+	break;
+      }
+      c = static_cast<char_type>(i);
+      if(std::isdigit(c) || (c =='.') || (c =='E') || (c =='e') || (c =='+') || (c =='-')){
+        buffer += c;
+      }else{
[…]

### 2016-06-16 src/CGAL/Regular_triangulation_3.h

@@ -27,8 +27,12 @@
 #include <CGAL/basic.h>
 
 #include <set>
+
 #ifdef CGAL_LINKED_WITH_TBB
+# include <CGAL/point_generators_3.h>
+# include <tbb/parallel_for.h>
 # include <tbb/enumerable_thread_specific.h>
+# include <tbb/concurrent_vector.h>
 #endif
 
 #include <CGAL/Triangulation_3.h>
@@ -44,6 +48,10 @@
 #endif //CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
 #ifdef CGAL_TRIANGULATION_3_PROFILING
 # include <CGAL/Mesh_3/Profiling_tools.h>
+#endif
+
+#ifdef CGAL_CONCURRENT_TRIANGULATION_3_ADD_TEMPORARY_POINTS_ON_FAR_SPHERE
+#include <CGAL/point_generators_3.h>
 #endif
 
 #if defined(BOOST_MSVC)
@@ -185,6 +193,79 @@ namespace CGAL {
       insert(first, last);
     }
 
+  private:
+
+#ifdef CGAL_CONCURRENT_TRIANGULATION_3_ADD_TEMPORARY_POINTS_ON_FAR_SPHERE
+    std::vector<Vertex_handle> 
+    add_temporary_points_on_far_sphere(const size_t num_points)
+    {
+      std::vector<Vertex_handle> far_sphere_vertices;
+
+      const size_t MIN_NUM_POINTS_FOR_FAR_SPHERE_POINTS = 1000000;
+      if (num_points >= MIN_NUM_POINTS_FOR_FAR_SPHERE_POINTS)
+      {
+        // Add temporary vertices on a "far sphere" to reduce contention on
+        // the infinite vertex
+
+        // Get bbox
+        const Bbox_3 &bbox = *this->get_bbox();
+        // Compute radius for far sphere
+        const double& xdelta = bbox.xmax() - bbox.xmin();
+        const double& ydelta = bbox.ymax() - bbox.ymin();
+        const double& zdelta = bbox.zmax() - bbox.zmin();
+        const double radius = 1.3 * 0.5 * std::sqrt(xdelta*xdelta +
+          ydelta*ydelta +
+          zdelta*zdelta);
+
+        // WARNING - TODO: this code has to be fixed because Vector_3 is not 
+        // required by the traits concept
+        const typename Gt::Vector_3 center(
[…]

### 2016-06-16 src/CGAL/Sqrt_extension/convert_to_bfi.h

@@ -27,6 +27,7 @@
 #include <CGAL/convert_to_bfi.h>
 #include <CGAL/Coercion_traits.h>
 #include <CGAL/Sqrt_extension/Sqrt_extension_type.h>
+#include <CGAL/assertions.h>
 
 
 // Disbale SQRT_EXTENSION_TO_BFI_CACHE by default

### 2016-06-16 src/CGAL/assertions.h

@@ -80,6 +80,7 @@ inline bool possibly(Uncertain<bool> c);
 
 #if defined(CGAL_NO_ASSERTIONS)
 #  define CGAL_assertion(EX) (static_cast<void>(0))
+#  define CGAL_destructor_assertion(EX) (static_cast<void>(0))
 #  define CGAL_assertion_msg(EX,MSG) (static_cast<void>(0))
 #  define CGAL_assertion_code(CODE)
 #  ifdef CGAL_ASSUME
@@ -92,6 +93,8 @@ inline bool possibly(Uncertain<bool> c);
 #else // no CGAL_NO_ASSERTIONS
 #  define CGAL_assertion(EX) \
    (CGAL::possibly(EX)?(static_cast<void>(0)): ::CGAL::assertion_fail( # EX , __FILE__, __LINE__))
+#  define CGAL_destructor_assertion(EX) \
+   (CGAL::possibly(EX)||std::uncaught_exception()?(static_cast<void>(0)): ::CGAL::assertion_fail( # EX , __FILE__, __LINE__))
 #  define CGAL_assertion_msg(EX,MSG) \
    (CGAL::possibly(EX)?(static_cast<void>(0)): ::CGAL::assertion_fail( # EX , __FILE__, __LINE__, MSG))
 #  define CGAL_assertion_code(CODE) CODE
@@ -342,5 +345,7 @@ inline bool possibly(Uncertain<bool> c);
 // This comes last as it is dependant on the macros to be defined.
 // But the macros need CGAL::possibly().
 #include <CGAL/Uncertain.h>
+
+#include "assertions.cpp"
 
 #endif // CGAL_ASSERTIONS_H

### 2016-06-16 src/CGAL/internal/Static_filters/Power_test_3.h

@@ -24,6 +24,7 @@
 
 #include <CGAL/Profile_counter.h>
 #include <CGAL/internal/Static_filters/Static_filter_error.h>
+#include <CGAL/internal/Static_filters/tools.h>
 #include <cmath>
 
 
@@ -132,6 +133,8 @@ namespace CGAL { namespace internal { namespace Static_filters_predicates {
       CGAL_BRANCH_PROFILER_3("semi-static failures/attempts/calls to   : Power_test_3 with 4+1 wpoints", tmp);
       
       double px, py, pz, pwt, qx, qy, qz, qwt, rx, ry, rz, rwt, sx, sy, sz, swt, tx, ty, tz, twt;
+      init_double(px, py, pz, pwt, qx, qy, qz, qwt, rx, ry, rz, rwt, (FT*)(0));
+      init_double(sx, sy, sz, swt, tx, ty, tz, twt, (FT*)(0));
       if( fit_in_double(p.x(), px) && fit_in_double(p.y(), py)       &&
           fit_in_double(p.z(), pz) && fit_in_double(p.weight(), pwt) && 
           fit_in_double(q.x(), qx) && fit_in_double(q.y(), qy)       &&
@@ -276,6 +279,8 @@ namespace CGAL { namespace internal { namespace Static_filters_predicates {
       CGAL_BRANCH_PROFILER_3("semi-static failures/attempts/calls to   : Power_test_3 with 3+1 wpoints", tmp);
 
       double px, py, pz, pwt, qx, qy, qz, qwt, rx, ry, rz, rwt, tx, ty, tz, twt;
+      init_double(px, py, pz, pwt, qx, qy, qz, qwt, (FT*)(0));
+      init_double( rx, ry, rz, rwt, tx, ty, tz, twt, (FT*)(0));
       if( fit_in_double(p.x(), px) && fit_in_double(p.y(), py)       &&
           fit_in_double(p.z(), pz) && fit_in_double(p.weight(), pwt) && 
           fit_in_double(q.x(), qx) && fit_in_double(q.y(), qy)       &&

### 2016-06-16 src/gudhi/Debug_utils.h

@@ -1,0 +1,55 @@
+/*    This file is part of the Gudhi Library. The Gudhi library
+ *    (Geometric Understanding in Higher Dimensions) is a generic C++
+ *    library for computational topology.
+ *
+ *    Author(s):       David Salinas
+ *
+ *    Copyright (C) 2014  INRIA Sophia Antipolis-Mediterranee (France)
+ *
+ *    This program is free software: you can redistribute it and/or modify
+ *    it under the terms of the GNU General Public License as published by
+ *    the Free Software Foundation, either version 3 of the License, or
+ *    (at your option) any later version.
+ *
+ *    This program is distributed in the hope that it will be useful,
+ *    but WITHOUT ANY WARRANTY; without even the implied warranty of
+ *    MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ *    GNU General Public License for more details.
+ *
+ *    You should have received a copy of the GNU General Public License
+ *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
+ */
+#ifndef DEBUG_UTILS_H_
+#define DEBUG_UTILS_H_
+
+#include <iostream>
+
+#ifndef NDEBUG
+  // GUDHI_DEBUG is the Gudhi official flag for debug mode.
+  #define GUDHI_DEBUG
+#endif
+
+// GUDHI_CHECK throw an exception if expression is false in debug mode, but does nothing in release mode
+// Could assert in release mode, but cmake sets NDEBUG (for "NO DEBUG") in this mode, means assert does nothing.
+#ifdef GUDHI_DEBUG
+  #define GUDHI_CHECK(expression, excpt) if ((expression) == 0) throw excpt
+#else
+  #define GUDHI_CHECK(expression, excpt) (void) 0
+#endif
+
+#define PRINT(a) std::cerr << #a << ": " << (a) << " (DISP)" << std::endl
+
+// #define DBG_VERBOSE
+#ifdef DBG_VERBOSE
+  #define DBG(a) std::cout << "DBG: " << (a) << std::endl
+  #define DBGMSG(a, b) std::cout << "DBG: " << a << b << std::endl
+  #define DBGVALUE(a) std::cout << "DBG: " <<  #a << ": " << a << std::endl
[…]

### 2016-06-16 src/gudhi/Simplex_tree/Simplex_tree_siblings.h

@@ -20,11 +20,12 @@
  *    along with this program.  If not, see <http://www.gnu.org/licenses/>.
  */
 
-#ifndef SRC_SIMPLEX_TREE_INCLUDE_GUDHI_SIMPLEX_TREE_SIMPLEX_TREE_SIBLINGS_H_
-#define SRC_SIMPLEX_TREE_INCLUDE_GUDHI_SIMPLEX_TREE_SIMPLEX_TREE_SIBLINGS_H_
+#ifndef SIMPLEX_TREE_SIMPLEX_TREE_SIBLINGS_H_
+#define SIMPLEX_TREE_SIMPLEX_TREE_SIBLINGS_H_
 
-#include "boost/container/flat_map.hpp"
-#include "Simplex_tree_node_explicit_storage.h"
+#include <gudhi/Simplex_tree/Simplex_tree_node_explicit_storage.h>
+
+#include <boost/container/flat_map.hpp>
 
 #include <utility>
 #include <vector>
@@ -56,7 +57,7 @@ class Simplex_tree_siblings {
 
   /* Default constructor.*/
   Simplex_tree_siblings()
-      : oncles_(NULL),
+      : oncles_(nullptr),
         parent_(-1),
         members_() {
   }
@@ -71,14 +72,14 @@ class Simplex_tree_siblings {
   /* \brief Constructor with initialized set of members.
    *
    * 'members' must be sorted and unique.*/
-  Simplex_tree_siblings(Simplex_tree_siblings * oncles, Vertex_handle parent,
-                        const std::vector<std::pair<Vertex_handle, Node> > & members)
+  template<typename RandomAccessVertexRange>
+  Simplex_tree_siblings(Simplex_tree_siblings * oncles, Vertex_handle parent, const RandomAccessVertexRange & members)
       : oncles_(oncles),
         parent_(parent),
         members_(boost::container::ordered_unique_range, members.begin(),
                  members.end()) {
-    for (auto map_it = members_.begin(); map_it != members_.end(); map_it++) {
-      map_it->second.assign_children(this);
+    for (auto& map_el : members_) {
+      map_el.second.assign_children(this);
     }
   }
 
@@ -90,19 +91,12 @@ class Simplex_tree_siblings {
    * present in the node.
    */
   void insert(Vertex_handle v, Filtration_value filtration_value) {
-    typename Dictionary::iterator sh = members_.find(v);
[…]

### 2016-06-16 src/tdautils/phatUtils.h

@@ -15,7 +15,7 @@
 #include <limits>
 #include <algorithm>
 
-#include <tdautils/dionysusUtils.h>
+#include <utilities/timer.h>
 
 
 template< typename Diagrams, typename PersistencePairs, typename Evaluator,
@@ -27,10 +27,12 @@ inline void initDiagramsPhat(Diagrams& diagrams, const PersistencePairs& pairs,
 	diagrams.resize(maxdimension + 1);
 	typename Diagrams::value_type::value_type dgmPoint(2);
 
-	// Manually add 0th homology for minimum to infinity
-	dgmPoint[0] = evaluator(simplex_map_inv[0]);
-	dgmPoint[1] = std::numeric_limits< double >::infinity();
-	diagrams[0].push_back(dgmPoint);
+	// If persistence is not empty, manually add 0th homology for minimum to infinity
+	if (simplex_map_inv.size() > 0) {
+  	dgmPoint[0] = evaluator(simplex_map_inv[0]);
+  	dgmPoint[1] = std::numeric_limits< double >::infinity();
+  	diagrams[0].push_back(dgmPoint);
+	}
 
 	for (phat::index idx = 0; idx < pairs.get_num_pairs(); ++idx) {
 		const typename SimplexMapInv::value_type& b =
@@ -60,11 +62,13 @@ inline void initLocationsPhat(Locations& locations,
 	typename Locations::value_type::value_type locPoint(2);
 	locations.resize(maxdimension + 1);
 
-	// Manually trace back birth & death point
-	locPoint[0] = getLocation(simplex_map_inv[0], FUNvalues);
-	locPoint[1] = (unsigned)(std::max_element(
-			FUNvalues.begin(), FUNvalues.end()) - FUNvalues.begin() + 1);
-	locations[0].push_back(locPoint);
+	// If persistence is not empty, manually trace back birth & death point
+	if (simplex_map_inv.size() > 0) {
+  	locPoint[0] = getLocation(simplex_map_inv[0], FUNvalues);
+  	locPoint[1] = (unsigned)(std::max_element(
+  			FUNvalues.begin(), FUNvalues.end()) - FUNvalues.begin() + 1);
+  	locations[0].push_back(locPoint);
+	}
 
 	for (phat::index idx = 0; idx < pairs.get_num_pairs(); ++idx) {
 		const typename SimplexMapInv::value_type& b =

### 2016-06-16 src/topology/static-persistence.h

@@ -1,6 +1,9 @@
 #ifndef __STATIC_PERSISTENCE_H__
 #define __STATIC_PERSISTENCE_H__
 
+// for Rcpp
+#include <Rcpp.h>
+
 #include "order.h"
 #include "cycles.h"
 #include "filtration.h"
@@ -137,7 +140,7 @@ class StaticPersistence
         // Acts as an archetype and if necessary a base class for visitors passed to <pair_simplices(bg, end, visitor)>.
         struct                          PairVisitor
         {
-                                        PairVisitor(unsigned size): show_progress(size)         {}
+                                        PairVisitor(unsigned size): show_progress(size, Rcpp::Rcout)         {}
             // Function: init(i)
             // Called after OrderElement pointed to by `i` has been initialized 
             // (its cycle is set to be its boundary, and pair is set to self, i.e. `i`)

### 2017-12-09 R/alphaComplexDiag.R
version 1.6

@@ -1,30 +1,66 @@
 alphaComplexDiag <-
-# function(X, maxalphasquare, library = "GUDHI", printProgress = FALSE) {
-function(X, library = "GUDHI", printProgress = FALSE) {
+# function(X, maxdimension = NCOL(X) - 1, maxalphasquare, library = "GUDHI",
+#     location = FALSE, printProgress = FALSE) {
+function(X, maxdimension = NCOL(X) - 1, library = "GUDHI", location = FALSE,
+    printProgress = FALSE) {
 
   if (!is.numeric(X) && !is.data.frame(X)) {
     stop("X should be a matrix of coordinates")
   }
+  if (!is.numeric(maxdimension) || length(maxdimension) != 1 ||
+      maxdimension < 0) {
+    stop("maxdimnsion should be a nonnegative integer")
+  }
 #  if (!is.numeric(maxalphasquare) || length(maxalphasquare) != 1 || maxalphasquare < 0) {
 #    stop("maxalphasquare should be a nonnegative number")
 #  }
-  if (library == "gudhi" || library == "Gudhi") {
-    library <- "GUDHI"
+  if (length(library) == 1) {
+    library <- rep(library, 2)
   }
-  if (library != "GUDHI") {
-    stop("library should be 'GUDHI'")
+  if (library[1] == "gudhi" || library[1] == "Gudhi") {
+    library[1] <- "GUDHI"
+  }
+  if (library[1] != "GUDHI") {
+    stop("library for building a filtration should be 'GUDHI'")
+  }
+  if (library[2] == "gudhi" || library[2] == "Gudhi") {
+    library[2] <- "GUDHI"
+  }
+  if (library[2] == "dionysus" || library[2] == "DIONYSUS") {
+    library[2] <- "Dionysus"
+  }
+  if (library[2] == "phat" || library[2] == "Phat") {
+    library[2] <- "PHAT"
+  }
+  if (library[2] != "GUDHI" && library[2] != "Dionysus" && library[2] != "PHAT") {
+    stop("library for computing persistence diagram should be a string: either 'GUDHI', 'Dionysus', or 'PHAT'")
+  }
+  if (!is.logical(location)) {
+    stop("location should be logical")
   }
   if (!is.logical(printProgress)) {
     stop("printProgress should be logical")
   }
 
   X <- as.matrix(X)
+  maxdimension <- min(maxdimension, NCOL(X) - 1)
[…]

### 2017-12-09 R/gridDiag.R

@@ -1,7 +1,7 @@
 gridDiag <-
 function(X = NULL, FUN = NULL, lim = NULL, by = NULL, FUNvalues = NULL,
          maxdimension = max(NCOL(X), length(dim(FUNvalues))) - 1,
-         sublevel = TRUE, library = "Dionysus", location = FALSE,
+         sublevel = TRUE, library = "GUDHI", location = FALSE,
          printProgress = FALSE, diagLimit = NULL, ...) {
 
   if (!xor(is.null(X) || is.null(FUN), is.null(FUNvalues))) {
@@ -46,14 +46,17 @@ function(X = NULL, FUN = NULL, lim = NULL, by = NULL, FUNvalues = NULL,
   if (!is.logical(sublevel)) {
     stop("sublevel should be logical")
   }
+  if (library == "gudhi" || library == "Gudhi") {
+    library <- "GUDHI"
+  }
   if (library == "dionysus" || library == "DIONYSUS") {
     library <- "Dionysus"
   }
   if (library == "phat" || library == "Phat") {
     library <- "PHAT"
   }
-  if (library != "Dionysus" && library != "PHAT") {
-    stop("library should be a string: either 'Dionysus' or 'PHAT'")
+  if (library != "GUDHI" && library != "Dionysus" && library != "PHAT") {
+    stop("library for computing persistence diagram should be a string: either 'GUDHI', 'Dionysus', or 'PHAT'")
   }
   if (!is.logical(location)) {
     stop("location should be logical")
@@ -99,13 +102,15 @@ function(X = NULL, FUN = NULL, lim = NULL, by = NULL, FUNvalues = NULL,
 
   # compute persistence diagram of function values over a grid
   if (length(gridDim) <= 3) {
-    gridOut <- GridDiag(FUNvalues = FUNvalues, gridDim = as.integer(gridDim),
+    gridOut <- GridDiag(
+	    FUNvalues = FUNvalues, gridDim = as.integer(gridDim),
         maxdimension = as.integer(maxdimension), decomposition = "5tetrahedra",
-        library = library, location = location, printProgress = printProgress)
+		library = library, location = location, printProgress = printProgress)
   } else {
-    gridOut <- GridDiag(FUNvalues = FUNvalues, gridDim = as.integer(gridDim),
+    gridOut <- GridDiag(
+	    FUNvalues = FUNvalues, gridDim = as.integer(gridDim),
[…]

### 2017-12-09 R/ripsFiltration.R

@@ -1,0 +1,59 @@
+ripsFiltration <- function(
+    X, maxdimension, maxscale, dist = "euclidean", library = "GUDHI",
+	printProgress = FALSE) {
+
+  if (!is.numeric(X) && !is.data.frame(X)) {
+    stop("X should be a matrix of coordinates")
+  }
+  if (!is.numeric(maxdimension) ||
+      length(maxdimension) != 1 || maxdimension < 0) {
+    stop("maxdimension should be a nonnegative integer")
+  }
+  if (!is.numeric(maxscale) || length(maxscale) != 1) {
+    stop("maxscale should be a number")
+  }
+  if (dist != "euclidean" && dist != "arbitrary") {
+    stop ("dist should be either 'euclidean' or 'arbitrary'")
+  }
+  if (library == "gudhi" || library == "Gudhi") {
+    library <- "GUDHI"
+  }
+  if (library == "dionysus" || library == "DIONYSUS") {
+    library <- "Dionysus"
+  }
+  if (library != "GUDHI" && library != "Dionysus") {
+    stop("library should be a string: either 'GUDHI' or 'Dionysus'")
+  }
+  if (!is.logical(printProgress)) {
+    stop("printProgress should be logical")
+  }
+
+  if (dist == "arbitrary" && library == "GUDHI") {
+    library <- "Dionysus"
+  }
+
+  X <- as.matrix(X)
+
+  max_num_pairs <- 5000  # to be added as an option
+
+  # in 32bit architectures Dionysus L2 doesn't work
+  #ripsOut <- RipsDiag(X = X, maxdimension = maxdimension,
+  #    maxscale = maxscale, dist = dist, library = library,
+  #    location = location, printProgress = printProgress)
+  if (dist == "euclidean" && library != "GUDHI" &&
+      .Machine[["sizeof.pointer"]] != 8) {
+    ripsOut <- RipsFiltration(
+	    X = as.matrix(dist(X)), maxdimension = maxdimension,
+		maxscale = maxscale, dist = "arbitrary", library = library,
+        printProgress = printProgress)
+  } else {
+    ripsOut <- RipsFiltration(
+	    X = X, maxdimension = maxdimension, maxscale = maxscale, dist = dist,
+        library = library, printProgress = printProgress)
+  }
+  
+  out <- list(
+      "cmplx" = ripsOut[[1]], "values" = ripsOut[[2]], "increasing" = TRUE)
+
+  return (out)
+}

### 2017-12-09 src/alphaShape.h

@@ -1,0 +1,73 @@
+#include <string>
+#include <vector>
+
+// for changing formats and typecasting
+#include <tdautils/typecastUtils.h>
+
+//for GUDHI
+#include <tdautils/gudhiUtils.h>
+
+// for Dionysus
+#include <tdautils/dionysusUtils.h>
+
+// for phat
+#include <tdautils/phatUtils.h>
+
+#include <tdautils/gridUtils.h>
+
+
+
+// AlphaShapeDiag
+/** \brief Interface for R code, construct the persistence diagram of the alpha
+ *        shape complex constructed on the input set of points.
+ *
+ * @param[out] Rcpp::List     A list
+ * @param[in]  X              An nx3 matrix of coordinates,
+ * @param[in]  printProgress  Is progress printed?
+ */
+template< typename RealMatrix, typename Print >
+void alphaShapeDiag(
+  const RealMatrix  & X,             //points to some memory space
+  const unsigned      nSample,
+  const unsigned      nDim,
+  const int           maxdimension,
+  const std::string & libraryDiag,
+  const bool          location,
+  const bool          printProgress,
+  const Print       & print,
+  std::vector< std::vector< std::vector< double > > > & persDgm,
+  std::vector< std::vector< std::vector< unsigned > > > & persLoc,
+  std::vector< std::vector< std::vector< std::vector< unsigned > > > > & persCycle
+) {
+
+  int coeff_field_characteristic = 2;
+
+  float min_persistence = 0.0;
+
+  Gudhi::Simplex_tree<> smplxTree =
+    AlphaShapeFiltrationGudhi< Gudhi::Simplex_tree<> >(
+      X, printProgress, print);
+
+  // Compute the persistence diagram of the complex
+  if (libraryDiag[0] == 'G') {
+    FiltrationDiagGudhi(
+        smplxTree, coeff_field_characteristic, min_persistence, 2,
+        printProgress, persDgm);
+  }
+  else if (libraryDiag[0] == 'D') {
+    Fltr filtration = filtrationGudhiToDionysus< Fltr >(smplxTree);
+    FiltrationDiagDionysus< Persistence >(
+        filtration, maxdimension, location, printProgress, persDgm, persLoc,
+        persCycle);
+  }
+  else {
+    std::vector< phat::column > cmplx;
+    std::vector< double > values;
+    phat::boundary_matrix< phat::vector_vector > boundary_matrix;
[…]

### 2017-12-09 src/diag.cpp

@@ -12,23 +12,75 @@
 // for grid
 #include <tdautils/gridUtils.h>
 
-// for kernel density
-#include <tdautils/kernelUtils.h>
-
 // for changing formats and typecasting
 #include <tdautils/typecastUtils.h>
 
 //for GUDHI
 #include <tdautils/gudhiUtils.h>
 
-//for CGAL
-#include <tdautils/cgalUtils.h>
-
 // for Dionysus
 #include <tdautils/dionysusUtils.h>
 
 // for phat
 #include <tdautils/phatUtils.h>
+
+#include <tdautils/filtrationUtils.h>
+#include <tdautils/filtrationDiag.h>
+
+#include <alphaComplex.h>
+#include <alphaShape.h>
+#include <dtm.h>
+#include <grid.h>
+#include <kde.h>
+#include <rips.h>
+
+
+
+// GridFiltration by Brittany T. Fasy
+// modified by Jisu Kim for
+// arbitrary dimension & using memory as an input & setting maximum dimension.
+/** \brief Interface for R code, construct the persistence diagram
+* of sublevel/superlevel sets of a function evaluated over a grid of points
+*
+* @param[out] Rcpp::List     A list
+* @param[in]  FUNvalues      A vector of length m1*...*md of function values over grid
+* @param[in]  gridDim        A vector (m1, ..., md),
+*                            where mi is number of grid points in ith dimension
+* @param[in]  maxdimension   Max dimension of the homological features to be computed.
+*                            This equals (maximal dimension of the Rips complex) - 1
+* @param[in]  decomposition  Either "5tetrahedra" or "barycenter"
+* @param[in]  library        Either "Dionysus" or "PHAT"
+* @param[in]  location       Are location of birth point, death point,
+*                            and representative cycles returned?
+* @param[in]  printProgress  Is progress printed?
+*/
+// [[Rcpp::export]]
+Rcpp::List GridFiltration(
+    const Rcpp::NumericVector & FUNvalues,
+    const Rcpp::IntegerVector & gridDim,
+    const int                   maxdimension,
+    const std::string         & decomposition,
+    const bool                  printProgress
+) {
+#ifdef LOGGING
+  //rlog::RLogInit(argc, argv);
+
+  stdoutLog.subscribeTo(RLOG_CHANNEL("topology/persistence"));
+  //stdoutLog.subscribeTo(RLOG_CHANNEL("topology/chain"));
[…]

### 2017-12-09 src/kde.h

@@ -1,0 +1,131 @@
+// for kernel density
+#include <tdautils/kernelUtils.h>
+#include <string>
+
+
+
+// KDE function on a Grid
+template< typename RealVector, typename RealMatrix, typename Print >
+inline RealVector kde(
+    const RealMatrix  & X,
+    const RealMatrix  & Grid,
+    const unsigned      nSample,
+    const unsigned      dim,
+    const unsigned      nGrid,
+    const double        h,
+    const std::string & kertype,
+    const RealVector  & weight,
+    const bool          printProgress,
+    const Print       & print
+) {
+  const double pi = 3.141592653589793;
+  const double h_to_dim = pow(h, (int)dim);
+//  const double den = pow(h, (int)dim) * pow(2 * pi, dim / 2.0);
+  RealVector kdeValue;
+  int counter = 0, percentageFloor = 0;
+  int totalCount = nGrid;
+
+  if (printProgress) {
+    printProgressFrame(print);
+  }
+
+  double(*kernel)(const double);
+
+  if (dim <= 1 || kertype.length() > 12) {
+    if (kertype[0] == 'G' || kertype[0] == 'g') {
+      kernel = gaussianSquare;
+    }
+    if (kertype[0] == 'E' || kertype[0] == 'e') {
+      kernel = epanechnikovSquare;
+    }
+    kdeValue = computeKernel< RealVector >(
+        X, Grid, nSample, dim, nGrid, h * h, kernel, weight, printProgress,
+        print, counter, totalCount, percentageFloor);
+  }
+  else {
+    if (kertype[0] == 'G' || kertype[0] == 'g') {
+      kernel = gaussian;
+    }
+    if (kertype[0] == 'E' || kertype[0] == 'e') {
+      kernel = epanechnikov;
+    }
+    kdeValue = computeGaussOuter< RealVector >(
+        X, Grid, nSample, dim, nGrid, h, kernel, weight, printProgress,
+        print, counter, totalCount, percentageFloor);
+  }
+
+  for (unsigned gridIdx = 0; gridIdx < nGrid; ++gridIdx) {
+    kdeValue[gridIdx] /= h_to_dim;
+  }
+
+  if (printProgress) {
+    print("\n");
+  }
+
+  return kdeValue;
+}
+
+
+
+// kernel Dist function on a Grid
+template< typename RealVector, typename RealMatrix, typename Print >
+inline RealVector kdeDist(
+    const RealMatrix & X,
[…]

### 2017-12-09 src/tdautils/filtrationUtils.h

@@ -1,0 +1,79 @@
+#ifndef __FILTRATIONUTILS_H__
+#define __FILTRATIONUTILS_H__
+
+#include <vector>
+#include <string>
+#include <algorithm>
+#include <functional>
+
+#include <utilities/timer.h>
+
+
+
+template< typename VertexVector, typename RealVector >
+void filtrationSort(
+    std::vector< VertexVector > & cmplx, RealVector & values) {
+
+  std::vector< std::pair< double, unsigned > > vipairs(cmplx.size());
+  typename std::vector< VertexVector >::iterator iCmplx;
+  typename RealVector::iterator iValue;
+  unsigned idx;
+  typename std::vector< std::pair< double, unsigned > >::iterator iPair;
+  std::vector< VertexVector > cmplxTemp(cmplx.begin(), cmplx.end());
+
+  iValue = values.begin();
+  idx = 0;
+  for (iPair = vipairs.begin(); iPair != vipairs.end();
+       ++iPair, ++iValue, ++iCmplx, ++idx) {
+    *iPair = std::make_pair(*iValue, idx);
+  }
+
+  std::sort(vipairs.begin(), vipairs.end());
+
+  iCmplx = cmplx.begin();
+  iValue = values.begin();
+  for (iPair = vipairs.begin(); iPair != vipairs.end();
+       ++iPair, ++iCmplx, ++iValue) {
+    *iCmplx = cmplxTemp[iPair->second];
+    *iValue = iPair->first;
+  }
+}
+
+
+
+// funFiltration
+/** \brief Interface for R code, construct the persistence diagram from the
+*         filtration of the function values.
+*
+* @param[out] Rcpp::List  A list
+* @param[in]  FUNvalues   The inlut function values
+* @param[in]  cmplx       The input simplicial complex
+*/template< typename RealVector, typename VertexVector >
+inline void funFiltration(
+    const RealVector            & FUNvalues,
+    std::vector< VertexVector > & cmplx,
+    std::vector< double >       & values
+) {
+
+  const unsigned nCmplx = cmplx.size();
+  values = std::vector< double >(nCmplx);
+
+  std::vector< double >::iterator iValue = values.begin();
+  for (typename std::vector< VertexVector >::iterator iCmplx = cmplx.begin();
+    iCmplx != cmplx.end(); ++iCmplx, ++iValue) {
[…]

### 2018-04-28 inst/doc/article.Rnw
version 1.6.2

@@
 %\VignetteIndexEntry{Introduction to TDA}
-%% need no \usepackage{Sweave.sty}
+%% need no \usepackage{Sweave.sty
 
@@
 \usepackage{verbatim,float}
-\usepackage{graphicx,subfigure,url}
-\usepackage{dsfont,bm,color,multirow}
+\usepackage{graphicx,subcaption,url}
+\usepackage{bm,color}
 
 %spacing for captions
-\usepackage{sidecap}
 \usepackage[font=small,skip=0pt]{caption}
@@
 
-\author{Brittany T. Fasy, Jisu Kim, Fabrizio Lecci \and Cl\'ement Maria \\ \small{In collaboration with the CMU TopStat Group}}
-\Plainauthor{Brittany Terese Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria}
+\author{Brittany T. Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, \and Vincent Rouvreau \\ \small{In collaboration with the CMU TopStat Group}}
+\Plainauthor{Brittany Terese Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, Vincent Rouvreau}
 
@@
 can be quantified with persistent homology. We provide an \proglang{R} interface
-for the efficient algorithms of the \proglang{C++} libraries \pkg{GUDHI}, \pkg{Dionysus} and \pkg{PHAT}, 
+for the efficient algorithms of the \proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, and \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}},
 including a function for the persistent homology of the Rips filtration, 
@@
   Computer Science Department\\
-  Tulane University\\
+  Montana State University\\
+  Website: \url{ http://www.cs.montana.edu/brittany/ } \\
   Email: \email{brittany.fasy@alumni.duke.edu}\\
@@
   Carnegie Mellon University\\
+  Website: \url{ http://www.stat.cmu.edu/~jisuk/ } \\
   Email: \email{jisuk1@andrew.cmu.edu}\\
 
+  Fabrizio Lecci\\
+  Email: \email{fabrizio.lecci@gmail.com}\\
+
   Cl\'ement Maria\\
-  Geometrica Group\\
-  INRIA Sophia Antipolis-M\'editerran\'ee\\
+  School of Mathematics and Physics\\
+  The University of Queensland\\
+  Website: \url{ http://www-sop.inria.fr/members/Clement.Maria/ } \\
   Email: \email{clement.maria@inria.fr}\\
 
-  Fabrizio Lecci\\
-  Department of Statistics\\
-  Carnegie Mellon University\\
-  Email: \email{lecci@cmu.edu}\\
+  Vincent Rouvreau\\
+  GUDHI Group\\
+  INRIA Saclay\\
+  Email: \email{vincent.rouvreau@inria.fr}\\
 
@@
   Carnegie Mellon University\\
-  Website: \url{http://www.stat.cmu.edu/topstat/} \\
+  Website: \url{ http://www.stat.cmu.edu/topstat/ } \\
   Email: \email{topstat@stat.cmu.edu}
@@
 %by Andrew Tausz;
-%\textsf{\textbf{PHAT}}\footnote{https://code.google.com/p/phat/} is a C++
+%\textsf{\textbf{PHAT}}\footnote{https://bitbucket.org/phat-code/phat/} is a C++
 %library developed at the Institute of Science and Technology Austria;
@@
 which provides a user-friendly interface for the efficient algorithms of the 
-\proglang{C++} libraries \pkg{GUDHI} \citep{gudhi_stpcoh}, \pkg{Dionysus} \citep{dionysus_C}, and \pkg{PHAT} \citep{phat_C}.
+\proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} \citep{gudhi_stpcoh}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}} \citep{dionysus_C}, and \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}} \citep{phat_C}.
 
@@
 \item Given a probability measure $P$, the distance to measure (DTM) is defined for each $y \in \mathbb{R}^d$ as
-$$
-d_{m_0}(y) =\sqrt{ \frac{1}{m_0}\int_0^{m_0} ( G_y^{-1}(u))^2 du},
-$$
+\[
+d_{m0}(y) = \left(\frac{1}{m0}\int_0^{m0} ( G_y^{-1}(u))^{r} du\right)^{1/r},
+\]
 where
-$G_y(t) = P( \Vert X-y \Vert \leq t)$ and $0<m_0<1$ is a smoothing parameter. 
+$G_y(t) = P( \Vert X-y \Vert \leq t)$, and $m0 \in (0,1)$ and $r \in [1,\infty)$ are tuning parameters.
+As \code{m0} increases, DTM function becomes smoother, so \code{m0} can be understood as a smoothing parameter.
+\code{r} affects less but also changes DTM function as well.
+The default value of \code{r} is \code{2}.
 The DTM can be seen as a smoothed version of the distance function.
-For more details
-see \cite{chazal2011geometric}.\\
+See \citep[Definition 3.2]{chazal2011geometric} and \citep[Equation (2)]{ChazalMM2015} for a formal definition of the "distance to measure" function.
+
 Given $X=\{x_1, \dots, x_n\}$, the empirical version of the DTM is
-$$
-\hat d_{m_0}(y) = \sqrt{ \frac{1}{k} \sum_{x_i \in N_k(y)} \Vert x_i-y \Vert^2 },
-$$
-where $k= \lceil m_0 n \rceil$ and $N_k(y)$ is the 
+\[
+\hat d_{m0}(y) = \left(\frac{1}{k} \sum_{x_i \in N_k(y)} \Vert x_i-y \Vert^{r}\right)^{1/r},
+\]
+where $k= \lceil m0 * n \rceil$ and $N_k(y)$ is the 
 set containing the $k$ nearest neighbors 
 of $y$ among $x_1, \ldots, x_n$.
+
+For more details, see \citep{chazal2011geometric} and \citep{ChazalMM2015}.
+
 The DTM is computed for each point of the \code{Grid} with the following code:
@@
 \cite{fasy2014statistical} and \cite{chazal2014robust} prove the validity of the bootstrap algorithm for kernel density estimators, distance to measure, and kernel distance, and use it in the framework of persistent homology.
-The bootstrap algorithm is implemented in the function \code{bootstrapBand}, which provides the option of parallelizing the algorithm (\code{parallel=TRUE}) using the package \pkg{parallel}.
+The bootstrap algorithm is implemented in the function \code{bootstrapBand}, which provides the option of parallelizing the algorithm (\code{parallel = TRUE}) using the package \pkg{parallel}.
 The following code computes a 90\% confidence band for $\mathbb{E}[\hat p_h]$, showed in Figure \ref{fig:eq10}.
@@
       zlim = c(0, max(band[["band"]][posYgrid, 2])),
-      ylim = range(Yseq), xlab = "", ylab = "", zlab="",theta = 0,
+      ylim = range(Yseq), xlab = "", ylab = "", zlab = "", theta = 0,
       phi = 25, ltheta = 50, col = "pink", border = NA, d = 0.5,
@@
 From version 1.2, \code{gridDiag} works in arbitrary dimension.
-The core of the function is written in \proglang{C++} and the user can choose to compute persistence diagrams using either the \pkg{Dionysus} library or the \pkg{PHAT} library.
-
-The following code computes the persistent homology of the superlevel sets \linebreak (\code{sublevel=FALSE}) of the kernel density estimator (\code{FUN=kde}, \code{h=0.3}) using the point cloud stored in the matrix~\code{X} from the previous example.
+The core of the function is written in \proglang{C++} and the user can choose to compute persistence diagrams using either  the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}}.
+
+The following code computes the persistent homology of the superlevel sets \linebreak (\code{sublevel = FALSE}) of the kernel density estimator (\code{FUN = kde}, \code{h = 0.3}) using the point cloud stored in the matrix~\code{X} from the previous example.
 The same code would work for the other functions defined in Section \ref{sec:distances} (it is sufficient to replace \code{kde} and its smoothing parameter \code{h} with another function and the corresponding parameter).
@@
 
-<<eq11, echo=TRUE, eval=TRUE>>=
-Diag <- gridDiag(X = X, FUN = kde, h = 0.3, lim = cbind(Xlim, Ylim),
-                 by = by, sublevel = FALSE, library = "Dionysus",
+<<eqGrid1, echo=TRUE, eval=TRUE>>=
+DiagGrid <- gridDiag(
+    X = X, FUN = kde, h = 0.3, lim = cbind(Xlim, Ylim), by = by,
+    sublevel = FALSE, library = "Dionysus", location = TRUE,
     printProgress = FALSE)
@@
 We plot the data and the diagram, using the function \code{plot}, implemented as a standard \code{S3} method for objects of the class \code{"diagram"}.
-The following command produces the third plot in Figure \ref{fig:eq11b}.
-
-<<eq11b, echo=TRUE, eval=FALSE>>=
-plot(Diag[["diagram"]], band = 2 * band[["width"]],
+The following command produces the third plot in Figure \ref{fig:eqGrid3}.
+
+<<eqGrid2, echo=TRUE, eval=FALSE>>=
+plot(DiagGrid[["diagram"]], band = 2 * band[["width"]],
      main = "KDE Diagram")
@@
 \begin{figure}[h!tb]
-\setkeys{Gin}{width=5.5in, height=2in}
+\setkeys{Gin}{width=5.5in, height=5.5in}
 \begin{center}
-<<eq11b, echo=FALSE, eval=TRUE, fig=TRUE, height=2, width=5.5>>=
-par(mfrow = c(1, 3))
-par(mai = c(0.3, 0.25, 0.3, 0.1))
+<<eqGrid3, echo=FALSE, eval=TRUE, fig=TRUE, height=7, width=6>>=
+par(mfrow = c(2, 2))
 plot(X, pch = 16, cex = 0.6, xlab = "", ylab = "", main = "Sample X")
@@
       scale = FALSE, expand = 3, shade = 0.9)
-par(mai = c(0.55, 0.6, 0.3, 0.1))
-plot(Diag[["diagram"]], band = 2 * band[["width"]],
+plot(DiagGrid[["diagram"]], band = 2 * band[["width"]],
      main = "KDE Diagram")
+one <- which(DiagGrid[["diagram"]][, 1] == 1)
+plot(X, pch = 16, cex = 0.6, xlab = "", ylab = "", main = "Representative loop")
+for (i in seq(along = one)) {
+  points(DiagGrid[["birthLocation"]][one[i], , drop = FALSE], pch = 15, cex = 3,
+      col = i)
+  points(DiagGrid[["deathLocation"]][one[i], , drop = FALSE], pch = 17, cex = 3,
+      col = i)
+  for (j in seq_len(dim(DiagGrid[["cycleLocation"]][[one[i]]])[1])) {
+    lines(DiagGrid[["cycleLocation"]][[one[i]]][j, , ], pch = 19, cex = 1, col = i)
+  }
+}
+
 @
@@
 The features are born at high levels of the density and die at lower levels. The pink 90\% confidence band separates significant features from noise.}
-\label{fig:eq11b}
+\label{fig:eqGrid3}
 \end{figure}
@@
 
-The option \code{band=2*band$width} produces a pink confidence band for the persistence diagram, using the confidence band constructed for the corresponding kernel density estimator in the previous section.
+The option (\code{band = 2 * band[["width"]]}) produces a pink confidence band for the persistence diagram, using the confidence band constructed for the corresponding kernel density estimator in the previous section.
 The features above the band can be interpreted as representing significant homological features, while points in the band are not significantly different from noise.
@@
 
-The function \code{plot} for the class \code{"diagram"} provide the options of rotating the diagram (\code{rotated=TRUE}), drawing the barcode in place of the diagram (\code{barcode=TRUE}), as well as other standard graphical options.
+The function \code{plot} for the class \code{"diagram"} provide the options of rotating the diagram (\code{rotated = TRUE}), drawing the barcode in place of the diagram (\code{barcode = TRUE}), as well as other standard graphical options.
 See Figure \ref{fig:eq11c}.
@@
 par(mfrow = c(1, 2), mai = c(0.8, 0.8, 0.3, 0.1))
-plot(Diag[["diagram"]], rotated = TRUE, band = band[["width"]],
+plot(DiagGrid[["diagram"]], rotated = TRUE, band = band[["width"]],
      main = "Rotated Diagram")
-plot(Diag[["diagram"]], barcode = TRUE,  main = "Barcode")
+plot(DiagGrid[["diagram"]], barcode = TRUE, main = "Barcode")
 @
@@
 The \code{ripsDiag} function computes the persistence diagram of the Rips filtration built on top of a point cloud.
-The user can choose to compute the Rips persistence diagram using either the \proglang{C++}  library \pkg{GUDHI}, or \pkg{Dionysus}.
+The user can choose to compute the Rips filtration using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} or \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
+Then for computing the persistence diagram from the Rips filtration, the user can use either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
 \\The following code generates 60 points from two circles:
 
-<<eq12, echo=TRUE, eval=TRUE>>=
+<<eqRips1, echo=TRUE, eval=TRUE>>=
 Circle1 <- circleUnif(60)
@@
 
-<<eq12b, echo=TRUE, eval=TRUE>>=
+<<eqRips2, echo=TRUE, eval=TRUE>>=
 maxscale <- 5        # limit of the filtration
@@
 
-<<eq12c, echo=TRUE, eval=TRUE>>=
-Diag <- ripsDiag(X = Circles, maxdimension, maxscale,
-                 library = "GUDHI", printProgress = FALSE)
-@
-Alternatively, using the option \code{dist="arbitrary"} in \code{ripsDiag()}, the input \code{X} can be an $n \times n$ matrix of distances.
-This option is useful when the user wants to consider a Rips filtration constructed using an arbitrary distance and is currently only available for the option \code{library="Dionysus"}.\\
-
-Finally we plot the data and the the diagram:\\
+<<eqRips3, echo=TRUE, eval=TRUE>>=
+DiagRips <- ripsDiag(X = Circles, maxdimension, maxscale,
+    library = c("GUDHI", "Dionysus"), location = TRUE, printProgress = FALSE)
+@
+Alternatively, using the option (\code{dist = "arbitrary"}) in \code{ripsDiag()}, the input \code{X} can be an $n \times n$ matrix of distances.
+This option is useful when the user wants to consider a Rips filtration constructed using an arbitrary distance and is currently only available for the option (\code{library = "Dionysus"}).\\
+
+Finally we plot the data and the diagram, as in Figure \ref{fig:eqRips4}.:\\
 \phantom{o}
+% PLOT START
+\begin{figure}[h!tb]
+\setkeys{Gin}{width=6in, height=2in}
+\begin{center}
+<<eqRips4, echo=FALSE, eval=TRUE, fig=TRUE, height=3, width=9>>=
+par(mfrow = c(1, 3), mai=c(0.8, 0.8, 0.3, 0.3))
+plot(Circles, pch = 16, xlab = "",ylab = "", main = "Two Circles")
+plot(DiagRips[["diagram"]], main = "Rips persistence diagram")
+one <- which(DiagRips[["diagram"]][, 1] == 1 &
+    DiagRips[["diagram"]][, 3] - DiagRips[["diagram"]][, 2] > 0.5)
+plot(Circles, col = 2, main = "Representative loop")
+for (i in seq(along = one)) {
+  for (j in seq_len(dim(DiagRips[["cycleLocation"]][[one[i]]])[1])) {
+    lines(DiagRips[["cycleLocation"]][[one[i]]][j, , ], pch = 19, cex = 1, col = i)
+  }
+}
+@
+\end{center}
+\caption{Rips persistence diagram. Black points represent connected components and red triangles represent loops.}
+\label{fig:eqRips4}
+\end{figure}
+% PLOT END
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Alpha Complex Persistence Diagram}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+For a finite set of points $X \subset \mathbb{R}^{d}$, the {\em Alpha complex} $Alpha(X,s)$ is a simplicial subcomplex of the Delaunay complex of $X$ consisting of simplices of circumradius less than or equal to $\sqrt{s}$.
+For each $u \in X$, let $V_{u}$ be its Voronoi cell, i.e.
+$V_{u}=\{x\in \mathbb{R}^{d}:\,d(x,u)\leq d(x,v)\text{ for all }v\in X\}$, and $B_{u}(r)$ be the closed ball with center $u$ and radius $r$.
+Let $R_{u}(r)$ consists of be the intersection of earh ball of radius $r$ with the voronoi cell of $u$, i.e.
+$R_{u}(r)=B_{u}(r)\cap V_{u}$.
+Then $Alpha(X,s)$ is defined as
+\[
+Alpha(X,r)=\left\{ \sigma\subset X:\,\underset{u\in\sigma}{\bigcap}R_{u}(\sqrt{s})\neq\emptyset\right\} .
+\]
+See \citep[Section 3.4]{edelsbrunner2010computational} and \citep{gudhi:AlphaComplex}.
+The sequence of Alpha complexes obtained by gradually increasing the parameter $s$ creates an Alpha complex filtration.
+
+The \code{alphaComplexDiag} function computes the Alpha complex filtration built on top of a point cloud, using the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}.
+Then for computing the persistence diagram from the Alpha complex filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
+
+We first generate 30 points from a circle:
+
+<<eqAlphaComplex1, echo=TRUE, eval=TRUE>>=
+X <- circleUnif(n = 30)
+@
+
+and the following code compute the persistence diagram of the alpha complex filtration using the point cloud \code{X}, with printing its progress (\code{printProgress = FALSE}).
+The function \code{alphaComplexDiag} returns an object of the class \code{"diagram"}.
+
+<<eqAlphaComplex2, echo=TRUE, eval=TRUE>>=
+# persistence diagram of alpha complex
+DiagAlphaCmplx <- alphaComplexDiag(
+    X = X, library = c("GUDHI", "Dionysus"), location = TRUE,
+    printProgress = TRUE)
+@
+
+And we plot the diagram in Figure \ref{fig:eqAlphaComplex3}.
+
+% PLOT START
+\begin{figure}[h!tb]
+\setkeys{Gin}{width=4in, height=2.3in}
+\begin{center}
+<<eqAlphaComplex3, echo=TRUE, eval=TRUE, fig=TRUE, height=3, width=8>>=
+# plot
+par(mfrow = c(1, 2))
+plot(DiagAlphaCmplx[["diagram"]], main = "Alpha complex persistence diagram")
+one <- which(DiagAlphaCmplx[["diagram"]][, 1] == 1)
+one <- one[which.max(
+    DiagAlphaCmplx[["diagram"]][one, 3] - DiagAlphaCmplx[["diagram"]][one, 2])]
+plot(X, col = 1, main = "Representative loop")
+for (i in seq(along = one)) {
+  for (j in seq_len(dim(DiagAlphaCmplx[["cycleLocation"]][[one[i]]])[1])) {
+    lines(DiagAlphaCmplx[["cycleLocation"]][[one[i]]][j, , ], pch = 19,
+	    cex = 1, col = i + 1)
+  }
+}
+par(mfrow = c(1, 1))
+@
+\end{center}
+\caption{Persistence diagram of Alpha complex. Black points represent connected components and red triangles represent loops.}
+\label{fig:eqAlphaComplex3}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{Persistence Diagram of Alpha Shape}
+%%%%%%%%%%%%%%%%%%%%%%%%%%
+The {\em Alpha shape complex} $S(X,\alpha)$ is the polytope with its boundary consisting of $\alpha$-exposed simplices, where a simplex $\sigma$ is $\alpha$-exposed if there is an open ball $b$ of radius $\alpha$ such that $b \cap X = \emptyset$ and $\partial b \cap X = \sigma$.
+Suppose $\mathbb{R}^{d}$ is filled with ice cream, then consider scooping out the ice cream with sphere-shaped spoon of radius $\alpha$ without touching the points $X$.
+$S(X,\alpha)$ is the remaining polytope with straightening round surfaces.
+See \citep{fischer2005} and \citep{edelsbrunner1994}.
+The sequence of Alpha shape complexes obtained by gradually increasing the parameter $\alpha$ creates an Alpha shape complex filtration.
+
+The \code{alphaShapeDiag} function computes the persistence diagram of the Alpha shape filtration built on top of a point cloud in 3 dimension, using the C++ library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}.
+Then for computing the persistence diagram from the Alpha shape filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
+Currently the point data cloud should lie in 3 dimension.
+
+We first generate 30 points from a cylinder:
+
+<<eqAlphaShape1, echo=TRUE, eval=TRUE>>=
+n <- 30
+X <- cbind(circleUnif(n = n), runif(n = n, min = -0.1, max = 0.1))
+@
+
+and the following code compute the persistence diagram of the alpha shape filtration using the point cloud \code{X}, with printing its progress (\code{printProgress = TRUE}).
+The function \code{alphaShapeDiag} returns an object of the class \code{"diagram"}.
+
+<<eqAlphaShape2, echo=TRUE, eval=TRUE>>=
+DiagAlphaShape <- alphaShapeDiag(
+    X = X, maxdimension = 1, library = c("GUDHI", "Dionysus"), location = TRUE,
+    printProgress = TRUE)
+@
+
+And we plot the diagram and first two dimension of data in Figure \ref{fig:eqAlphaShape3}.
+
+% PLOT START
+\begin{figure}[h!tb]
+\setkeys{Gin}{width=4in, height=2.3in}
+\begin{center}
+<<eqAlphaShape3, echo=TRUE, eval=TRUE, fig=TRUE, height=3, width=8>>=
+par(mfrow = c(1, 2))
+plot(DiagAlphaShape[["diagram"]])
+plot(X[, 1:2], col = 2, main = "Representative loop of alpha shape filtration")
+one <- which(DiagAlphaShape[["diagram"]][, 1] == 1)
+one <- one[which.max(
+    DiagAlphaShape[["diagram"]][one, 3] - DiagAlphaShape[["diagram"]][one, 2])]
+for (i in seq(along = one)) {
+  for (j in seq_len(dim(DiagAlphaShape[["cycleLocation"]][[one[i]]])[1])) {
+    lines(
+        DiagAlphaShape[["cycleLocation"]][[one[i]]][j, , 1:2], pch = 19,
+        cex = 1, col = i)
+  }
+}
+par(mfrow = c(1, 1))
+@
+\end{center}
+\caption{Persistence diagram of Alpha shape. Black points represent connected components and red triangles represent loops.}
+\label{fig:eqAlphaShape3}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%
+\subsection{Persistence Diagrams from Filtration}
+%%%%%%%%%%%%%%%%%%
+Rather than computing persistence diagrams from built-in function, it is also possible to compute persistence diagrams from a user-defined filtration.
+A filtration consists of simplicial complex and the filtration values on each simplex.
+The functions \code{ripsDiag}, \code{alphaComplexDiag}, \code{alphaShapeDiag} have their counterparts for computing corresponding filtrations instead of persistence diagrams: namely, \code{ripsFiltration} corresponds to the Rips filtration built on top of a point cloud, \code{alphaComplexFiltration} to the alpha complex filtration, and \code{alphaShapeFiltration} to the alpha shape filtration.
+
+We first generate 100 points from a circle:
+
+<<eqFiltration1, echo=TRUE, eval=TRUE>>=
+X <- circleUnif(n = 100)
+@
+
+\noindent Then, after specifying the limit of the Rips filtration and the max dimension of the homological features, the following code compute the Rips filtration using the point cloud \code{X}.
+
+<<eqFiltration2, echo=TRUE, eval=TRUE>>=
+maxscale <- 0.4      # limit of the filtration
+maxdimension <- 1    # components and loops
+FltRips <- ripsFiltration(X = X, maxdimension = maxdimension,
+    maxscale = maxscale, dist = "euclidean", library = "GUDHI",
+	printProgress = TRUE)
+@
+
+\noindent One way of defining a user-defined filtration is to build a filtration from a simplicial complex and function values on the vertices.
+The function \code{funFiltration} takes function values (\code{FUNvalues}) and simplicial complex (\code{cmplx}) as input, and build a filtration, where a filtration value on a simplex is defined as the maximum of function values on the vertices of the simplex.
+
+\noindent In the following example, the function \code{funFiltration} construct a filtration from a Rips complex and the DTM function values on data points.
+
+<<eqFiltration3, echo=TRUE, eval=TRUE>>=
+m0 <- 0.1
+dtmValues <- dtm(X = X, Grid = X, m0 = m0)
+FltFun <- funFiltration(FUNvalues = dtmValues, cmplx = FltRips[["cmplx"]])
+@
+
+\noindent Once the filtration is computed, the function \code{filtrationDiag} computes the persistence diagram from the filtration.
+The user can choose to compute the persistence diagram using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} or \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
+
+<<eqFiltration4, echo=TRUE, eval=TRUE>>=
+DiagFltFun <- filtrationDiag(filtration = FltFun, maxdimension = maxdimension,
+    library = "Dionysus", location = TRUE, printProgress = TRUE)
+@
+
+Then we plot the data and the diagram in Figure \ref{fig:eqFiltration5}.
 % PLOT START
@@
 \begin{center}
-<<eq12d, echo=TRUE, eval=TRUE,  fig=TRUE, height=2.8, width=5.6>>=
+<<eqFiltration5, echo=TRUE, eval=TRUE, fig=TRUE, height=3, width=6>>=
 par(mfrow = c(1, 2), mai=c(0.8, 0.8, 0.3, 0.3))
-plot(Circles, pch = 16, xlab = "",ylab = "")
-plot(Diag[["diagram"]])
+plot(X, pch = 16, xlab = "",ylab = "")
+plot(DiagFltFun[["diagram"]], diagLim = c(0, 1))
 @
 \end{center}
-\caption{Rips persistence diagram. Black points represent connected components and red triangles represent loops.}
-\label{fig:eq12d}
+\caption{Persistence diagram from Rips filtration and DTM function values. Black points represent connected components and red triangles represent loops.}
+\label{fig:eqFiltration5}
 \end{figure}
@@
 Standard metrics for measuring the distance between two persistence diagrams are the bottleneck distance and the $p$th Wasserstein distance \citep{edelsbrunner2010computational}. 
-The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \proglang{C++}  library \pkg{Dionysus}.
+The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \proglang{C++} library \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
 
@@
 and we compute the bottleneck distance and the $2$nd Wasserstein distance between the two diagrams.
-In the following code, the option \code{dimension=1} specifies that the distances between diagrams are computed using only one dimensional features (loops).
+In the following code, the option \code{dimension = 1} specifies that the distances between diagrams are computed using only one dimensional features (loops).
 
@@
 The landscape and silhouette functions can be evaluated over a one-dimensional grid of points \code{tseq} using the functions \code{landscape} and \code{silhouette}.
-In the following code, we use the persistence diagram from Figure \ref{fig:eq12d} to construct the corresponding landscape and silhouette for
-one-dimensional features (\code{dimension=1}).
-The option \code{KK=1} specifies that we are interested in the 1st landscape function, and \code{p=1} is the power of the weights in the definition of the silhouette function.
+In the following code, we use the persistence diagram from Figure \ref{fig:eqRips4} to construct the corresponding landscape and silhouette for
+one-dimensional features (\code{dimension = 1}).
+The option (\code{KK = 1}) specifies that we are interested in the 1st landscape function, and (\code{p = 1}) is the power of the weights in the definition of the silhouette function.
 
 <<eq14, echo=TRUE, eval=TRUE>>=
+maxscale <- 5
 tseq <- seq(0, maxscale, length = 1000)   #domain
-Land <- landscape(Diag[["diagram"]], dimension = 1, KK = 1, tseq)
-Sil <- silhouette(Diag[["diagram"]], p = 1, dimension = 1, tseq)
-@
-
-The functions \code{landscape} and \code{silhouette} return real valued vectors, which can be simply plotted with \code{plot(tseq, Land, type="l"); plot(tseq, Sil, type="l")}.
+Land <- landscape(DiagRips[["diagram"]], dimension = 1, KK = 1, tseq)
+Sil <- silhouette(DiagRips[["diagram"]], p = 1, dimension = 1, tseq)
+@
+
+The functions \code{landscape} and \code{silhouette} return real valued vectors, which can be simply plotted with \code{plot(tseq, Land, type = "l"); plot(tseq, Sil, type = "l")}.
 See Figure \ref{fig:eq14b}.
@@
 <<eq14b, echo=FALSE, eval=TRUE, fig=TRUE, height=2.4, width=8>>=
+ylimit <- c(0, max(c(Land, Sil)) * 1.2)
 par(mfrow = c(1, 2), mai = c(0.5, 0.45, 0.3, 0.3))
-plot(tseq, Land, type = "l", lwd = 3, ylab = "",
+plot(tseq, Land, type = "l", lwd = 3, ylim = ylimit, ylab = "",
      main = "1st Landscape, dim = 1", asp = 1, col = 2)
-plot(tseq, Sil, type = "l", lwd = 3, ylab = "",
+plot(tseq, Sil, type = "l", lwd = 3, ylim = ylimit, ylab = "",
      main = "Silhouette(p = 1), dim = 1", asp = 1, col = 2)
@@
 \end{center}
-\caption{Landscape and Silhouette of the one-dimensional features of the diagram of Figure \ref{fig:eq12d}.}
+\caption{Landscape and Silhouette of the one-dimensional features of the diagram of Figure \ref{fig:eqRips4}.}
 \label{fig:eq14b}
@@
 For $n$ times, we subsample from the large point cloud, compute $n$ Rips diagrams and the corresponding
-1st landscape functions (\code{KK=1}), using 1 dimensional features (\code{dimension=1}):
+1st landscape functions (\code{KK = 1}), using 1 dimensional features (\code{dimension = 1}):
 
@@
 An unsolved problem in topological inference is how to
-choose the smoothing parameters, for example $h$ for KDE and 
-$m_0$ for DTM.
+choose the smoothing parameters, for example \code{h} for KDE and 
+\code{m0} for DTM.
 
@@
 
-The function \code{maxPersistence} can be parallelized (\code{parallel=TRUE}) and a progress bar can be printed (\code{printProgress=TRUE}):
+The function \code{maxPersistence} can be parallelized (\code{parallel = TRUE}) and a progress bar can be printed (\code{printProgress = TRUE}):
 
@@
 
-Then we use the function \code{clusterTree} to compute cluster trees using the k Nearest Neighbors density estimator (\code{k=100} nearest neighbors) and the Gaussian kernel density estimator, with smoothing parameter \code{h}.
+Then we use the function \code{clusterTree} to compute cluster trees using the k Nearest Neighbors density estimator (\code{k = 100} nearest neighbors) and the Gaussian kernel density estimator, with smoothing parameter \code{h}.
 
@@
 
-Note that, even when \code{kde} is used to estimate the density, we have to provide the option \code{k=100}, so that the algorithm can compute the connected components at each level of the density using a k Nearest Neighbors graph.
+Note that, even when \code{kde} is used to estimate the density, we have to provide the option (\code{k = 100}), so that the algorithm can compute the connected components at each level of the density using a k Nearest Neighbors graph.
 

### 2018-04-28 R/gridFiltration.R

@@ -89,13 +89,19 @@ function(X = NULL, FUN = NULL, lim = NULL, by = NULL, FUNvalues = NULL,
         maxdimension = as.integer(maxdimension), decomposition = "barycenter",
 		printProgress = printProgress)
   }
+
+  if (sublevel == FALSE) {
+    gridOut[[2]] <- -gridOut[[2]]
+  }
   
-  if (sublevel == TRUE) {
+  if (!is.null(lim) && !is.null(by)) {
     out <- list(
-      "cmplx" = gridOut[[1]], "values" = gridOut[[2]], "increasing" = sublevel)
+	    "cmplx" = gridOut[[1]], "values" = gridOut[[2]],
+        "increasing" = sublevel, "coordinates" = Grid[["grid"]])
   } else {
     out <- list(
-      "cmplx" = gridOut[[1]], "values" = -gridOut[[2]], "increasing" = sublevel)
+        "cmplx" = gridOut[[1]], "values" = -gridOut[[2]],
+        "increasing" = sublevel)
   }
 
   return (out)

### 2018-04-28 src/diag.cpp

@@ -400,10 +400,15 @@ Rcpp::List AlphaShapeFiltration(
   const bool                  printProgress
 ) {
 
+  Rcpp::NumericMatrix coordinates;
+
   Gudhi::Simplex_tree<> smplxTree =
       AlphaShapeFiltrationGudhi< Gudhi::Simplex_tree<> >(
-          X, printProgress, Rprintf);
-  return filtrationGudhiToRcpp< Rcpp::List, Rcpp::NumericVector >(smplxTree);
+          X, printProgress, Rprintf, coordinates);
+  Rcpp::List filtration =
+      filtrationGudhiToRcpp< Rcpp::List, Rcpp::NumericVector >(smplxTree);
+  filtration.push_back(coordinates);
+  return filtration;
 }
 
 
@@ -429,16 +434,18 @@ Rcpp::List AlphaShapeDiag(
   std::vector< std::vector< std::vector< unsigned > > > persLoc;
   std::vector< std::vector< std::vector< std::vector< unsigned > > > >
       persCycle;
+  Rcpp::NumericMatrix coordinates;
 
   alphaShapeDiag(
       X, X.nrow(), X.ncol(), maxdimension, libraryDiag, location,
-      printProgress, Rprintf, persDgm, persLoc, persCycle);
+      printProgress, Rprintf, persDgm, persLoc, persCycle, coordinates);
 
   // Output persistent diagram
   return Rcpp::List::create(
       concatStlToRcpp< Rcpp::NumericMatrix >(persDgm, true, 3),
       concatStlToRcpp< Rcpp::NumericMatrix >(persLoc, false, 2),
-      StlToRcppMatrixList< Rcpp::List, Rcpp::NumericMatrix >(persCycle));
+      StlToRcppMatrixList< Rcpp::List, Rcpp::NumericMatrix >(persCycle),
+      coordinates);
 }
 
 
@@ -461,7 +468,11 @@ Rcpp::List AlphaComplexFiltration(
   Gudhi::Simplex_tree<> smplxTree =
       AlphaComplexFiltrationGudhi< Gudhi::Simplex_tree<> >(
           X, printProgress, Rprintf);
-  return filtrationGudhiToRcpp< Rcpp::List, Rcpp::NumericVector >(smplxTree);
+  // 2018-04-24
+  // temporary fix for bug in alphaComplex
[…]

### 2018-04-28 src/topology/rips.hpp

@@ -31,7 +31,11 @@ generate(Dimension k, DistanceType max, const Functor& f, Iterator bg, Iterator 
     // candidates   = everything
     VertexContainer current;
     VertexContainer candidates(bg, end);
-    bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    // 2018-04-27
+    // change 'excluded' to 'excludedNext' (where excludedNext = excluded + 1) to prevent overflow
+    // and change logic accordingly
+    // bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    bron_kerbosch(current, candidates, candidates.begin(), k, neighbor, f);
 }
 
 template<class D, class S>
@@ -49,7 +53,11 @@ vertex_cofaces(IndexType v, Dimension k, DistanceType max, const Functor& f, Ite
     for (Iterator cur = bg; cur != end; ++cur)
         if (*cur != v && neighbor(v, *cur))
             candidates.push_back(*cur);
-    bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    // 2018-04-27
+    // change 'excluded' to 'excludedNext' (where excludedNext = excluded + 1) to prevent overflow
+    // and change logic accordingly
+    // bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    bron_kerbosch(current, candidates, candidates.begin(), k, neighbor, f);
 }
 
 template<class D, class S>
@@ -75,7 +83,11 @@ edge_cofaces(IndexType u, IndexType v, Dimension k, DistanceType max, const Func
             rLog(rlRipsDebug,   "  added candidate: %d", *cur);
         }
 
-    bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    // 2018-04-27
+    // change 'excluded' to 'excludedNext' (where excludedNext = excluded + 1) to prevent overflow
+    // and change logic accordingly
+    // bron_kerbosch(current, candidates, boost::prior(candidates.begin()), k, neighbor, f);
+    bron_kerbosch(current, candidates, candidates.begin(), k, neighbor, f);
 }
 
 template<class D, class S>
@@ -111,7 +123,11 @@ cofaces(const Simplex& s, Dimension k, DistanceType max, const Functor& f, Itera
         }
     }
 
[…]

### 2018-08-06 src/tdautils/typecastUtils.h
version 1.6.4

@@ -283,10 +283,9 @@ inline RcppList StlToRcppMatrixList(
 template< typename SimplexHandle, typename SimplexTree, typename RealVector >
 void filtrationGudhiOne(
     const SimplexHandle & sh, SimplexTree & smplxTree, const int idxShift,
-    RealVector & cmplxVec, double & value, RealVector & boundaryVec,
-    // 2018-04-24
-    // temporary fix for bug in alphaComplex
-    bool isBdy = true) {
+    // 2018-08-04
+    // switching back to original code
+    RealVector & cmplxVec, double & value, RealVector & boundaryVec) {
 
   const unsigned nVtx = smplxTree.dimension(sh) + 1;
 
@@ -302,23 +301,21 @@ void filtrationGudhiOne(
 
   value = SimplexTree::filtration(sh);
 
-  // 2018-04-24
-  // temporary fix for bug in alphaComplex
-  if (isBdy) {
+  // 2018-08-04
+  // switching back to original code
 
-    // might need to change for cubical complex
-    if (nVtx > 1) {
-      boundaryVec = RealVector(nVtx);
-    }
-    const typename SimplexTree::Boundary_simplex_range & smplxRange =
+  // might need to change for cubical complex
+  if (nVtx > 1) {
+    boundaryVec = RealVector(nVtx);
+  }
+  const typename SimplexTree::Boundary_simplex_range & smplxRange =
       smplxTree.boundary_simplex_range(sh);
 
-    typename RealVector::iterator iBdyVec = boundaryVec.begin();
-    for (typename SimplexTree::Boundary_simplex_iterator iBdySpx =
+  typename RealVector::iterator iBdyVec = boundaryVec.begin();
+  for (typename SimplexTree::Boundary_simplex_iterator iBdySpx =
       smplxRange.begin(); iBdySpx != smplxRange.end(); ++iBdySpx, ++iBdyVec) {
-      // R is 1-base, while C++ is 0-base
-      *iBdyVec = SimplexTree::key(*iBdySpx) + idxShift;
-    }
+    // R is 1-base, while C++ is 0-base
+    *iBdyVec = SimplexTree::key(*iBdySpx) + idxShift;
   }
 }
 
@@ -329,10 +326,9 @@ template< typename IntegerVector, typename SimplexTree, typename VectorList,
           typename RealVector >
 inline void filtrationGudhiToTda(
[…]

### 2019-02-07 R/findKtree.R
version 1.6.5

@@ -3,14 +3,18 @@ function(bb, parent, children, compBranch, n) {
   kappaTop <- numeric(bb)
   kappaBottom <- numeric(bb)
   for (i in seq_len(bb)) {
-    if (is.na(parent[i])) {
+    # 2019-01-18
+	# is.na() returns a vector, wrap with all() or any() if needed
+    if (any(is.na(parent[i]))) {
       kappaBottom[i] <- NA
       kappaTop[i] <- NA
     } else if (parent[i] == 0) { 
         kappaBottom[i] <- 0 
         if (i <= length(children)) {
           Kchildren <- children[[i]]
-          if (!is.null(Kchildren) && !is.na(Kchildren)) {
+		  # 2019-01-18
+	      # is.na() returns a vector, wrap with all() or any() if needed
+          if (!is.null(Kchildren) && all(!is.na(Kchildren))) {
           kappaTop[i] <-
               (length(compBranch[[i]]) - length(unlist(compBranch[Kchildren]))) / n
           } else {
@@ -23,7 +27,10 @@ function(bb, parent, children, compBranch, n) {
         kappaBottom[i] <- kappaTop[parent[i]]
         if (i <= length(children)) {
           Kchildren <- children[[i]]
-          if (!is.null(Kchildren) && length(Kchildren) != 0 && !is.na(Kchildren)) {
+		  # 2019-01-18
+		  # is.na() returns a vector, wrap with all() or any() if needed
+          if (!is.null(Kchildren) && length(Kchildren) != 0
+		      && all(!is.na(Kchildren))) {
             kappaTop[i] <- kappaBottom[i] + (length(compBranch[[i]]) -
                 length(unlist(compBranch[Kchildren]))) / n
           } else {

### 2019-02-07 inst/doc/article.R

@@ -1,5 +1,5 @@
 ### R code from vignette source 'article.Rnw'
-### Encoding: CP949
+### Encoding: UTF-8
 
 ###################################################
 ### code chunk number 1: initial_settings

### 2019-02-07 src/CGAL/Cartesian/Aff_transformation_rep_2.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Andreas Fabri, Lutz Kettner
@@ -144,19 +145,19 @@ friend class Scaling_repC2<R>;
             {
               case 0: return t11;
               case 1: return t12;
-              case 2: return t13;
+              default: return t13;
             }
     case 1: switch (j)
             {
               case 0: return t21;
               case 1: return t22;
-              case 2: return t23;
+              default: return t23;
             }
     case 2: switch (j)
             {
               case 0: return FT(0);
               case 1: return FT(0);
-              case 2: return FT(1);
+              default: return FT(1);
             }
     }
     return FT(0);

### 2019-02-07 src/CGAL/Cartesian_converter.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Sylvain Pion
@@ -31,6 +32,7 @@
 // provided you give a NT converter from A to B.
 // There's a Homogeneous counterpart.
 
+#include <CGAL/Cartesian_converter_fwd.h>
 #include <CGAL/basic.h>
 #include <CGAL/NT_converter.h>
 #include <CGAL/Enum_converter.h>
@@ -90,7 +92,7 @@ struct Converting_visitor : boost::static_visitor<> {
 
 template < class K1, class K2,
 //          class Converter = NT_converter<typename K1::FT, typename K2::FT> >
-           class Converter = typename internal::Default_converter<K1, K2>::Type >
+           class Converter>
 class Cartesian_converter : public Enum_converter
 {
     typedef Enum_converter   Base;
@@ -218,6 +220,13 @@ public:
 	return Point_2(c(a.x()), c(a.y()));
     }
 
+    typename K2::Weighted_point_2
+    operator()(const typename K1::Weighted_point_2 &a) const
+    {
+        typedef typename K2::Weighted_point_2 Weighted_point_2;
+	return Weighted_point_2(operator()(a.point()), operator()(a.weight()));
+    }
+
     typename K2::Vector_2
     operator()(const typename K1::Vector_2 &a) const
     {
@@ -284,6 +293,13 @@ public:
     {
         typedef typename K2::Point_3 Point_3;
 	return Point_3(c(a.x()), c(a.y()), c(a.z()));
+    }
+
+    typename K2::Weighted_point_3
+    operator()(const typename K1::Weighted_point_3 &a) const
+    {
+        typedef typename K2::Weighted_point_3 Weighted_point_3;
+	return Weighted_point_3((*this)(a.point()), c(a.weight()));
     }
 
     typename K2::Vector_3

### 2019-02-07 src/CGAL/Compact_container.h

@@ -15,15 +15,19 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 // Author(s)     : Sylvain Pion
 
 #ifndef CGAL_COMPACT_CONTAINER_H
 #define CGAL_COMPACT_CONTAINER_H
 
-#include <CGAL/basic.h>
+#include <CGAL/disable_warnings.h>
+
+#include <CGAL/config.h>
 #include <CGAL/Default.h>
 
+#include <cmath>
 #include <iterator>
 #include <algorithm>
 #include <vector>
@@ -234,9 +238,9 @@ class Compact_container
   typedef Allocator_                                Al;
   typedef typename Default::Get< Al, CGAL_ALLOCATOR(T) >::type Allocator;
   typedef Increment_policy_                         Ip;
-  typedef typename Default::Get< Ip, 
+  typedef typename Default::Get< Ip,
             Addition_size_policy<CGAL_INIT_COMPACT_CONTAINER_BLOCK_SIZE,
-                             CGAL_INCREMENT_COMPACT_CONTAINER_BLOCK_SIZE> 
+                             CGAL_INCREMENT_COMPACT_CONTAINER_BLOCK_SIZE>
           >::type                                   Increment_policy;
   typedef TimeStamper_                              Ts;
   typedef Compact_container <T, Al, Ip, Ts>         Self;
@@ -248,14 +252,24 @@ public:
 
   typedef T                                         value_type;
   typedef Allocator                                 allocator_type;
-  typedef typename Allocator::reference             reference;
-  typedef typename Allocator::const_reference       const_reference;
+
+  typedef value_type&                               reference;
+  typedef const value_type&                         const_reference;
+
+#ifdef CGAL_CXX11
+  typedef typename std::allocator_traits<Allocator>::pointer               pointer;
+  typedef typename std::allocator_traits<Allocator>::const_pointer         const_pointer;
+  typedef typename std::allocator_traits<Allocator>::size_type             size_type;
+  typedef typename std::allocator_traits<Allocator>::difference_type       difference_type;
+#else
   typedef typename Allocator::pointer               pointer;
[…]

### 2019-02-07 src/CGAL/Conic_2.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Bernd Gaertner, Sven Schoenherr <sven@inf.ethz.ch>
@@ -27,7 +28,7 @@
 #define CGAL_CONIC_2_H
 
 #include <CGAL/Point_2.h>
-#include <CGAL/Conic_misc.h>
+#include <CGAL/Kernel/Conic_misc.h>
 #include <CGAL/Dimension.h>
 
 namespace CGAL {

### 2019-02-07 src/CGAL/Gmpzf.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Michael Hemmer   <hemmer@mpi-inf.mpg.de>
@@ -36,7 +37,7 @@ public:
     typedef Tag_true            Is_exact;
 
     struct Is_zero
-        : public std::unary_function< Type, bool > {
+        : public CGAL::cpp98::unary_function< Type, bool > {
     public:
         bool operator()( const Type& x ) const {
             return x.is_zero();
@@ -44,7 +45,7 @@ public:
     };
 
     struct Integral_division
-        : public std::binary_function< Type,
+        : public CGAL::cpp98::binary_function< Type,
                                 Type,
                                 Type > {
     public:
@@ -56,7 +57,7 @@ public:
     };
 
     struct Gcd
-        : public std::binary_function< Type,
+        : public CGAL::cpp98::binary_function< Type,
                                 Type,
                                 Type > {
     public:
@@ -69,7 +70,7 @@ public:
     };
 
     class Div
-        : public std::binary_function< Type, Type, Type > {
+        : public CGAL::cpp98::binary_function< Type, Type, Type > {
     public:
         Type operator()( const Type& x, const Type& y ) const {
             return Type(x).div( y );
@@ -79,7 +80,7 @@ public:
     typedef INTERN_AST::Mod_per_operator< Type > Mod;
   
   class Is_square
-    : public std::binary_function< Type, Type&, bool > {
+    : public CGAL::cpp98::binary_function< Type, Type&, bool > {
   public:      
     bool operator()( const Type& x, Type& y ) const {
       y = CGAL::approximate_sqrt(x);
@@ -103,7 +104,7 @@ public:
   typedef AST::Is_zero Is_zero;
   
     struct Sgn
-        : public std::unary_function< Type, ::CGAL::Sign > {
+        : public CGAL::cpp98::unary_function< Type, ::CGAL::Sign > {
     public:
         ::CGAL::Sign operator()( const Type& x ) const {
             return x.sign();
@@ -111,7 +112,7 @@ public:
     };
 
     struct Compare
-        : public std::binary_function< Type,
+        : public CGAL::cpp98::binary_function< Type,
                                   Type,
                                   Comparison_result > {
     public:
@@ -123,7 +124,7 @@ public:
     };
 
     struct To_double
[…]

### 2019-02-07 src/CGAL/IO/File_binary_mesh_3.h

@@ -1,74 +1,0 @@
-// Copyright (c) 2012  GeometryFactory Sarl (France).
-// All rights reserved.
-//
-// This file is part of CGAL (www.cgal.org).
-// You can redistribute it and/or modify it under the terms of the GNU
-// General Public License as published by the Free Software Foundation,
-// either version 3 of the License, or (at your option) any later version.
-//
-// Licensees holding a valid commercial license may use this file in
-// accordance with the commercial license agreement provided with the software.
-//
-// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
-// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-//
-// $URL$
-// $Id$
-//
-//
-// Author(s)     : Laurent Rineau
-
-#ifndef CGAL_IO_FILE_BINARY_MESH_3_H
-#define CGAL_IO_FILE_BINARY_MESH_3_H
-
-#include <iostream>
-#include <string>
-#include <CGAL/Mesh_3/io_signature.h>
-
-namespace CGAL {
-
-namespace Mesh_3 {
-
-template <class C3T3>
-bool
-save_binary_file(std::ostream& os,
-                 const C3T3& c3t3)
-{
-  os << "binary CGAL c3t3 " << CGAL::Get_io_signature<C3T3>()() << "\n";
-  CGAL::set_binary_mode(os);
-  return !!(os << c3t3);
-  // call operator!() twice, because operator bool() is C++11
-}
-
-template <class C3T3>
-bool load_binary_file(std::istream& is, C3T3& c3t3)
-{
-  std::string s;
-  is >> s;
-  if (s != "binary" ||
-      !(is >> s) ||
-      s != "CGAL" ||
-      !(is >> s) ||
-      s != "c3t3") 
-  {
-    return false;
-  }
-  std::getline(is, s);
-  if(s != "") {
-    if(s != std::string(" ") + CGAL::Get_io_signature<C3T3>()()) {
-      //std::cerr << "load_binary_file:"
-      //          << "\n  expected format: " << CGAL::Get_io_signature<C3T3>()()
-      //          << "\n       got format:" << s << std::endl;
-      return false;
-    }
-  }
-  CGAL::set_binary_mode(is);
-  is >> c3t3;
-  return !!is;
-  // call operator!() twice, because operator bool() is C++11
-}
-
[…]

### 2019-02-07 src/CGAL/IO/Gps_iostream.h

@@ -1,52 +1,0 @@
-// Copyright (c) 2009  Tel-Aviv University (Israel).
-// All rights reserved.
-//
-// This file is part of CGAL (www.cgal.org).
-// You can redistribute it and/or modify it under the terms of the GNU
-// General Public License as published by the Free Software Foundation,
-// either version 3 of the License, or (at your option) any later version.
-//
-// Licensees holding a valid commercial license may use this file in
-// accordance with the commercial license agreement provided with the software.
-//
-// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
-// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-//
-// $URL$
-// $Id$
-// 
-//
-// Author(s)     : Efi Fogel <efif@post.tau.ac.il>
-
-#ifndef CGAL_GPS_IOSTREAM_H
-#define CGAL_GPS_IOSTREAM_H
-
-#include <iostream>
-#include <list>
-
-#include <CGAL/basic.h>
-#include <CGAL/General_polygon_set_2.h>
-
-namespace CGAL {
-
-template <typename Traits>
-std::ostream & operator<< (std::ostream& os,
-                           const CGAL::General_polygon_set_2<Traits> & pgn_set)
-{
-  typedef typename CGAL::General_polygon_set_2<Traits>::Polygon_with_holes_2
-                                                        Polygon_with_holes_2;
-  typedef std::list<Polygon_with_holes_2>               Pgn_with_holes_container;
-
-  Pgn_with_holes_container res;
-  pgn_set.polygons_with_holes (std::back_inserter (res));
-
-  //std::cout << pgn_set.number_of_polygons_with_holes() << std::endl;
-  std::copy(res.begin(), res.end(),
-            std::ostream_iterator<Polygon_with_holes_2>(std::cout, "\n"));
-  
-  return os;
-}
-
-} //namespace CGAL
-
-#endif

### 2019-02-07 src/CGAL/IO/read_xyz_points.h

@@ -1,428 +1,0 @@
-// Copyright (c) 2007-09  INRIA Sophia-Antipolis (France).
-// All rights reserved.
-//
-// This file is part of CGAL (www.cgal.org).
-// You can redistribute it and/or modify it under the terms of the GNU
-// General Public License as published by the Free Software Foundation,
-// either version 3 of the License, or (at your option) any later version.
-//
-// Licensees holding a valid commercial license may use this file in
-// accordance with the commercial license agreement provided with the software.
-//
-// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
-// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-//
-// $URL$
-// $Id$
-//
-// Author(s) : Pierre Alliez and Laurent Saboret
-
-#ifndef CGAL_READ_XYZ_POINTS_H
-#define CGAL_READ_XYZ_POINTS_H
-
-#include <CGAL/property_map.h>
-#include <CGAL/value_type_traits.h>
-#include <CGAL/point_set_processing_assertions.h>
-
-#include <boost/version.hpp>
-#if BOOST_VERSION >= 104000
-  #include <boost/property_map/property_map.hpp>
-#else
-  #include <boost/property_map.hpp>
-#endif
-
-#include <iostream>
-#include <sstream>
-#include <string>
-
-namespace CGAL {
-
-
-//===================================================================================
-/// \ingroup PkgPointSetProcessing
-/// Reads points (positions + normals, if available) from a .xyz ASCII stream.
-/// The function expects for each point a line with the x y z position,
-/// optionally followed by the nx ny nz normal.
-/// The first line may contain the number of points in the file.
-/// Empty lines and comments starting by # character are allowed.
-///
-/// @tparam OutputIteratorValueType type of objects that can be put in `OutputIterator`.
-///         It is default to `value_type_traits<OutputIterator>::%type` and can be omitted when the default is fine.
-/// @tparam OutputIterator iterator over output points.
-/// @tparam PointPMap is a model of `WritablePropertyMap` with  value type `Point_3<Kernel>`.
-///        It can be omitted if the value type of OutputIterator value_type is convertible to `Point_3<Kernel>`.
-/// @tparam NormalPMap is a model of `WritablePropertyMap` with value type  `Vector_3<Kernel>`.
-/// @tparam Kernel Geometric traits class.
[…]

### 2019-02-07 src/CGAL/Intersection_traits_3.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Philipp Möller
@@ -22,6 +23,7 @@
 #define CGAL_INTERSECTION_TRAITS_3_H
 
 #include <CGAL/Intersection_traits.h>
+#include <vector>
 
 #if !(CGAL_INTERSECTION_VERSION < 2)
 

### 2019-02-07 src/CGAL/Kernel_checker.h

@@ -1,324 +1,0 @@
-// Copyright (c) 2001  
-// Utrecht University (The Netherlands),
-// ETH Zurich (Switzerland),
-// INRIA Sophia-Antipolis (France),
-// Max-Planck-Institute Saarbruecken (Germany),
-// and Tel-Aviv University (Israel).  All rights reserved. 
-//
-// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
-// modify it under the terms of the GNU Lesser General Public License as
-// published by the Free Software Foundation; either version 3 of the License,
-// or (at your option) any later version.
-//
-// Licensees holding a valid commercial license may use this file in
-// accordance with the commercial license agreement provided with the software.
-//
-// This file is provided AS IS with NO WARRANTY OF ANY KIND, INCLUDING THE
-// WARRANTY OF DESIGN, MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE.
-//
-// $URL$
-// $Id$
-// 
-//
-// Author(s)     : Sylvain Pion
-
-#ifndef CGAL_KERNEL_CHECKER_H
-#define CGAL_KERNEL_CHECKER_H
-
-// This file contains the definition of a kernel traits checker.
-//
-// TODO:
-// - At the moment, only predicates are checked.  To handle constructions as
-//   well, the best approach is probably to have objects be pairs, and do
-//   everything in parallel (cf Curved_kernel's traits checker for an example).
-//   So the template parameter will be a comparator, not a converter.
-
-#include <CGAL/basic.h>
-#include <CGAL/use.h>
-#include <utility>
-#include <typeinfo>
-
-namespace CGAL {
-
-// Small utility to manipulate pairs for kernel objects, and
-// simple things for bool, Sign...  Object is yet another case...
-template < typename T1, typename T2 >
-struct Pairify {
-  typedef std::pair<T1, T2>  result_type;
-  result_type operator()(const T1 &t1, const T2 &t2) const
-  { return std::make_pair(t1, t2); }
-};
-
-template <>
-struct Pairify <bool, bool> {
-  typedef bool   result_type;
-  result_type operator()(const bool &t1, const bool &t2) const
[…]

### 2019-02-07 src/CGAL/MP_Float_impl.h

@@ -15,6 +15,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Sylvain Pion
@@ -46,7 +47,7 @@ inline
 int my_nearbyint(const T& d)
 {
   int z = int(d);
-  T frac = d - z;
+  T frac = d - T(z);
 
   CGAL_assertion(CGAL::abs(frac) < T(1.0));
 
@@ -95,9 +96,9 @@ void MP_Float::construct_from_builtin_fp_type(T d)
     T orig = d, sum = 0;
     while (true) {
       int r = my_nearbyint(d);
-      if (d-r >= T( INTERN_MP_FLOAT::base/2-1)/( INTERN_MP_FLOAT::base-1))
+      if (d-double(r) >= T( INTERN_MP_FLOAT::base/2-1)/( INTERN_MP_FLOAT::base-1))
         ++r;
-      v.push_back(r);
+      v.push_back(limb(r));
       // We used to do simply "d -= v.back();", but when the most significant
       // limb is 1 and the second is -32768, then it can happen that
       // |d - v.back()| > |d|, hence a bit of precision can be lost.

### 2019-02-07 src/CGAL/NewKernel_d/LA_eigen/LA.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 // Author(s)     : Marc Glisse
 
@@ -135,7 +136,12 @@ template<class NT_,class Dim_,class Max_dim_=Dim_> struct LA_eigen {
 
 	// m*a==b
 	template<class DV, class DM, class V>
-	static bool solve(DV&a, DM const&m, V const& b){
+	static void solve(DV&a, DM const&m, V const& b){
+		//a = m.colPivHouseholderQr().solve(b);
+		a = m.fullPivLu().solve(b);
+	}
+	template<class DV, class DM, class V>
+	static bool solve_and_check(DV&a, DM const&m, V const& b){
 		//a = m.colPivHouseholderQr().solve(b);
 		a = m.fullPivLu().solve(b);
 		return b.isApprox(m*a);

### 2019-02-07 src/CGAL/Point_2.h

@@ -1,9 +1,9 @@
-// Copyright (c) 1999  
+// Copyright (c) 1999
 // Utrecht University (The Netherlands),
 // ETH Zurich (Switzerland),
 // INRIA Sophia-Antipolis (France),
 // Max-Planck-Institute Saarbruecken (Germany),
-// and Tel-Aviv University (Israel).  All rights reserved. 
+// and Tel-Aviv University (Israel).  All rights reserved.
 //
 // This file is part of CGAL (www.cgal.org); you can redistribute it and/or
 // modify it under the terms of the GNU Lesser General Public License as
@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Andreas Fabri, Stefan Schirra
@@ -32,6 +33,7 @@
 #include <CGAL/Kernel/Return_base_tag.h>
 #include <CGAL/representation_tags.h>
 #include <CGAL/Dimension.h>
+#include <CGAL/result_of.h>
 
 namespace CGAL {
 
@@ -51,6 +53,7 @@ public:
   typedef Dimension_tag<2>  Ambient_dimension;
   typedef Dimension_tag<0>  Feature_dimension;
 
+  typedef typename R_::Weighted_point_2 Weighted_point_2;
   typedef RPoint_2 Rep;
   typedef typename R_::Cartesian_const_iterator_2 Cartesian_const_iterator;
 
@@ -74,6 +77,11 @@ public:
 
   Point_2(const RPoint_2& p)
     : RPoint_2(p)
+  {}
+
+  explicit
+  Point_2(const Weighted_point_2& wp)
+    : Rep(wp.point())
   {}
 
   template < typename T1, typename T2 >
@@ -121,7 +129,6 @@ public:
   }
 
 
-
   typename cpp11::result_of<typename R::Compute_hx_2(Point_2)>::type
   hx() const
   {
@@ -160,6 +167,21 @@ public:
   Point_2 transform(const Aff_transformation_2 &t) const
   {
     return t.transform(*this);
+  }
+
+  Point_2<R_>&
+  operator+=(const typename R::Vector_2 &v)
+  {
+    *this = R().construct_translated_point_2_object()(*this, v);
+    return *this;
+  }
+
+  Point_2<R_>&
+  operator-=(const typename R::Vector_2 &v)
+  {
+    *this = R().construct_translated_point_2_object()(*this,
+                  R().construct_opposite_vector_2_object()(v));
[…]

### 2019-02-07 src/CGAL/Point_3.h

@@ -1,9 +1,9 @@
-// Copyright (c) 1999  
+// Copyright (c) 1999
 // Utrecht University (The Netherlands),
 // ETH Zurich (Switzerland),
 // INRIA Sophia-Antipolis (France),
 // Max-Planck-Institute Saarbruecken (Germany),
-// and Tel-Aviv University (Israel).  All rights reserved. 
+// and Tel-Aviv University (Israel).  All rights reserved.
 //
 // This file is part of CGAL (www.cgal.org); you can redistribute it and/or
 // modify it under the terms of the GNU Lesser General Public License as
@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Andreas Fabri, Stefan Schirra
@@ -50,6 +51,7 @@ public:
   typedef Dimension_tag<3>  Ambient_dimension;
   typedef Dimension_tag<0>  Feature_dimension;
 
+  typedef typename R_::Weighted_point_3 Weighted_point_3;
   typedef typename R_::Kernel_base::Point_3  Rep;
   typedef typename R_::Cartesian_const_iterator_3 Cartesian_const_iterator;
 
@@ -73,6 +75,11 @@ public:
 
   Point_3(const Rep& p)
       : Rep(p) {}
+
+  explicit
+  Point_3(const Weighted_point_3& wp)
+    : Rep(wp.point())
+  {}
 
   template < typename T1, typename T2, typename T3 >
   Point_3(const T1& x, const T2& y, const T3& z)
@@ -175,6 +182,21 @@ public:
     return t.transform(*this);
   }
 
+  Point_3<R_>&
+  operator+=(const Vector_3 &v)
+  {
+    *this = R().construct_translated_point_3_object()(*this, v);
+    return *this;
+  }
+
+  Point_3<R_>&
+  operator-=(const Vector_3 &v)
+  {
+    *this = R().construct_translated_point_3_object()(*this,
+                  R().construct_opposite_vector_3_object()(v));
+    return *this;
+  }
+
 };
 
 template <class R>
@@ -242,7 +264,7 @@ template <class R >
 std::istream&
 extract(std::istream& is, Point_3<R>& p, const Cartesian_tag&)
 {
-    typename R::FT x, y, z;
+  typename R::FT x(0), y(0), z(0);
     switch(get_mode(is)) {
     case IO::ASCII :
         is >> iformat(x) >> iformat(y) >> iformat(z);
[…]

### 2019-02-07 src/CGAL/Spatial_sort_traits_adapter_d.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 // Author(s)     : Sebastien Loriot
 
@@ -22,13 +23,8 @@
 #define CGAL_SPATIAL_SORT_TRAITS_ADAPTER_D_H
 
 #include <boost/call_traits.hpp>
-#include <boost/version.hpp>
-#if BOOST_VERSION >= 104000
-  #include <boost/property_map/property_map.hpp>
-#else
-  #include <boost/property_map.hpp>
-#endif
 
+#include <CGAL/property_map.h>
 
 namespace CGAL{
 

### 2019-02-07 src/CGAL/TDS_full_cell_default_storage_policy.h

@@ -14,11 +14,15 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: GPL-3.0+
 //
 // Author(s)    : Samuel Hornus
 
 #ifndef CGAL_TDS_FULL_CELL_DEFAULT_STORAGE_POLICY_H
 #define CGAL_TDS_FULL_CELL_DEFAULT_STORAGE_POLICY_H
+
+#include <CGAL/license/Triangulation.h>
+
 
 #include <CGAL/Dimension.h>
 #include <CGAL/Compact_container.h>
@@ -61,10 +65,10 @@ struct TFC_data< Vertex_handle, Full_cell_handle, Dimen, TDS_full_cell_default_s
     {
         Xor_type result(0);
         for( int i = 0; i <= cur_dim; ++i )
-			// modified by Jisu KIM, 2017-04-24
-			// '*vertices_[i]' can fetch NULL memory if 'vertices_[i]' corresponds to NULL pointer
+            // modified by Jisu KIM, 2017-04-24
+            // '*vertices_[i]' can fetch NULL memory if 'vertices_[i]' corresponds to NULL pointer
             // result ^= reinterpret_cast<Xor_type>(&(*vertices_[i]));
-			result ^= reinterpret_cast<Xor_type>((vertices_[i]).operator->());
+            result ^= reinterpret_cast<Xor_type>((vertices_[i]).operator->());
         return result;
     }
     // ASSUMES |*this| is indeed a neighbor of neighbor(i):
@@ -82,10 +86,10 @@ struct TFC_data< Vertex_handle, Full_cell_handle, Dimen, TDS_full_cell_default_s
     {
         Xor_type opp_vertex = xor_of_vertices(cur_dim)
             ^ neighbors_[i]->xor_of_vertices(cur_dim)
-			// modified by Jisu KIM, 2017-04-24
-			// '*vertices_[i]' can fetch NULL memory if 'vertices_[i]' corresponds to NULL pointer
+            // modified by Jisu KIM, 2017-04-24
+            // '*vertices_[i]' can fetch NULL memory if 'vertices_[i]' corresponds to NULL pointer
             //^ reinterpret_cast<Xor_type>(&(*vertices_[i]));
-			^ reinterpret_cast<Xor_type>((vertices_[i]).operator->());
+            ^ reinterpret_cast<Xor_type>((vertices_[i]).operator->());
         Vertex_handle mirror;
[…]

### 2019-02-07 src/CGAL/Triangulation_hierarchy_3.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: GPL-3.0+
 //
 // Author(s)     : Olivier Devillers <Olivier.Devillers@sophia.inria.fr>
 //                 Sylvain Pion
@@ -21,25 +22,11 @@
 #ifndef CGAL_TRIANGULATION_HIERARCHY_3_H
 #define CGAL_TRIANGULATION_HIERARCHY_3_H
 
-#include <CGAL/basic.h>
-#include <CGAL/triangulation_assertions.h>
-#include <CGAL/Triangulation_hierarchy_vertex_base_3.h>
-#include <CGAL/Location_policy.h>
+#include <CGAL/license/Triangulation_3.h>
 
-#include <boost/random/linear_congruential.hpp>
-#include <boost/random/geometric_distribution.hpp>
-#include <boost/random/variate_generator.hpp>
-
-#ifndef CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
-#include <CGAL/Spatial_sort_traits_adapter_3.h>
-#include <CGAL/internal/info_check.h>
-
-#include <boost/tuple/tuple.hpp>
-#include <boost/iterator/zip_iterator.hpp>
-#include <boost/mpl/and.hpp>
-#endif //CGAL_TRIANGULATION_3_DONT_INSERT_RANGE_OF_POINTS_WITH_INFO
-
-namespace CGAL {
+// Commented because the class is actually used by Delaunay_triangulation_hierarchy_3.h
+// #define CGAL_DEPRECATED_HEADER "<CGAL/Triangulation_hierarchy_3.h>"
+// #include <CGAL/internal/deprecation_warning.h>
 
 // This class is deprecated, but must be kept for backward compatibility.
 //
@@ -48,6 +35,34 @@ namespace CGAL {
 // empty nutshell instead.
 //
 // Then, later, maybe merge the Compact/Fast codes in a cleaner factorized way.
+
+#include <CGAL/basic.h>
+#include <CGAL/internal/Has_nested_type_Bare_point.h>
+#include <CGAL/triangulation_assertions.h>
+#include <CGAL/Triangulation_hierarchy_vertex_base_3.h>
+#include <CGAL/Location_policy.h>
+
+#include <CGAL/internal/boost/function_property_map.hpp>
+
+#include <boost/random/linear_congruential.hpp>
[…]

### 2019-02-07 src/CGAL/Vector_2.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Andreas Fabri, Stefan Schirra
@@ -30,8 +31,11 @@
 #include <CGAL/assertions.h>
 #include <boost/type_traits/is_same.hpp>
 #include <CGAL/Kernel/Return_base_tag.h>
+#include <CGAL/kernel_assertions.h>
 #include <CGAL/representation_tags.h>
 #include <CGAL/Dimension.h>
+#include <CGAL/result_of.h>
+#include <CGAL/IO/io.h>
 
 namespace CGAL {
 
@@ -175,9 +179,21 @@ public:
     return R().construct_difference_of_vectors_2_object()(*this,v);
   }
 
+  Vector_2& operator-=(const Vector_2& v)
+  {
+    *this = R().construct_difference_of_vectors_2_object()(*this,v);
+    return *this;
+  }
+
   Vector_2 operator+(const Vector_2& v) const
   {
     return R().construct_sum_of_vectors_2_object()(*this,v);
+  }
+
+  Vector_2& operator+=(const Vector_2& v)
+  {
+    *this = R().construct_sum_of_vectors_2_object()(*this,v);
+    return *this;
   }
 
   Vector_2 operator/(const RT& c) const
@@ -185,9 +201,33 @@ public:
    return R().construct_divided_vector_2_object()(*this,c);
   }
 
+  Vector_2& operator/=(const RT& c)
+  {
+    *this = R().construct_divided_vector_2_object()(*this,c);
+    return *this;
+  }
+
   Vector_2 operator/(const typename First_if_different<FT,RT>::Type & c) const
   {
    return R().construct_divided_vector_2_object()(*this,c);
+  }
+
+  Vector_2& operator/=(const typename First_if_different<FT,RT>::Type & c)
+  {
+    *this = R().construct_divided_vector_2_object()(*this,c);
+    return *this;
+  }
+
+  Vector_2& operator*=(const RT& c)
+  {
+    *this = R().construct_scaled_vector_2_object()(*this,c);
+    return *this;
+  }
+
+  Vector_2& operator*=(const typename First_if_different<FT,RT>::Type & c)
+  {
+    *this = R().construct_scaled_vector_2_object()(*this,c);
+    return *this;
   }
 
   FT squared_length() const
@@ -296,7 +336,7 @@ template <class R >
 std::istream&
[…]

### 2019-02-07 src/CGAL/Vector_3.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Andreas Fabri, Stefan Schirra
@@ -28,10 +29,13 @@
 #include <CGAL/Origin.h>
 #include <CGAL/Kernel/mpl.h>
 #include <CGAL/representation_tags.h>
+#include <CGAL/kernel_assertions.h>
 #include <CGAL/assertions.h>
 #include <boost/type_traits/is_same.hpp>
 #include <CGAL/Kernel/Return_base_tag.h>
 #include <CGAL/Dimension.h>
+#include <CGAL/result_of.h>
+#include <CGAL/IO/io.h>
 
 namespace CGAL {
 
@@ -39,7 +43,7 @@ template <class R_>
 class Vector_3 : public R_::Kernel_base::Vector_3
 {
   typedef typename R_::RT                    RT;
-// http://doc.cgal.org/latest/Manual/devman_code_format.html#secprogramming_conventions
+// https://doc.cgal.org/latest/Manual/devman_code_format.html#secprogramming_conventions
   typedef typename R_::FT                    FT_;
   typedef typename R_::Segment_3             Segment_3;
   typedef typename R_::Ray_3                 Ray_3;
@@ -118,9 +122,21 @@ public:
     return R().construct_difference_of_vectors_3_object()(*this,v);
   }
 
+  Vector_3& operator-=(const Vector_3& v)
+  {
+    *this = R().construct_difference_of_vectors_3_object()(*this,v);
+    return *this;
+  }
+
   Vector_3 operator+(const Vector_3& v) const
   {
     return R().construct_sum_of_vectors_3_object()(*this,v);
+  }
+
+  Vector_3& operator+=(const Vector_3& v)
+  {
+    *this = R().construct_sum_of_vectors_3_object()(*this,v);
+    return *this;
   }
 
   Vector_3 operator/(const RT& c) const
@@ -128,9 +144,33 @@ public:
    return R().construct_divided_vector_3_object()(*this,c);
   }
 
+  Vector_3& operator/=(const RT& c)
+  {
+    *this = R().construct_divided_vector_3_object()(*this,c);
+    return *this;
+  }
+
   Vector_3 operator/(const typename First_if_different<FT_,RT>::Type & c) const
   {
    return R().construct_divided_vector_3_object()(*this,c);
+  }
+
[…]

### 2019-02-07 src/CGAL/algorithm.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Michael Hoffmann <hoffmann@inf.ethz.ch>
@@ -27,10 +28,16 @@
 #ifndef CGAL_ALGORITHM_H
 #define CGAL_ALGORITHM_H
 
-#include <CGAL/basic.h>
 #include <CGAL/config.h>
+#include <CGAL/utils.h>
+#include <CGAL/enum.h>
 #include <algorithm>
 #include <iosfwd>
+#include <iostream>
+
+#include <boost/random/random_number_generator.hpp>
+#include <boost/random.hpp>
+#include <boost/random/linear_congruential.hpp>
 
 #ifdef CGAL_CFG_NO_CPP0X_NEXT_PREV
 #  include <boost/next_prior.hpp>
@@ -362,6 +369,9 @@ output_range(std::ostream& os,
     return os;
 }
 
+
+namespace cpp98 {
+
 // Reimplementation of std::random_shuffle, for the use of Spatial_sorting.
 // We want an implementation of random_shuffle that produces the same
 // result on all platforms, for a given seeded random generator.
@@ -379,6 +389,20 @@ random_shuffle(RandomAccessIterator begin, RandomAccessIterator end,
     // interval [0, N[
   }
 }
+
+
+template <class RandomAccessIterator>
+void
+random_shuffle(RandomAccessIterator begin, RandomAccessIterator end)
+{
+  typedef std::iterator_traits<RandomAccessIterator> Iterator_traits;
+  typedef typename Iterator_traits::difference_type Diff_t;
+  boost::rand48 random;
+  boost::random_number_generator<boost::rand48, Diff_t> rng(random);
+  CGAL::cpp98::random_shuffle(begin,end, rng);
+}
+
+} // namespace cpp98
 
 namespace internal {
 namespace algorithm {

### 2019-02-07 src/CGAL/config.h

@@ -3,7 +3,7 @@
 // ETH Zurich (Switzerland),
 // INRIA Sophia-Antipolis (France),
 // Max-Planck-Institute Saarbruecken (Germany),
-// and Tel-Aviv University (Israel).  All rights reserved. 
+// and Tel-Aviv University (Israel).  All rights reserved.
 //
 // This file is part of CGAL (www.cgal.org); you can redistribute it and/or
 // modify it under the terms of the GNU Lesser General Public License as
@@ -18,15 +18,21 @@
 //
 // $URL$
 // $Id$
-// 
+// SPDX-License-Identifier: LGPL-3.0+
 //
-// Author(s)     : Wieger Wesselink 
+//
+//
+// Author(s)     : Wieger Wesselink
 //                 Michael Hoffmann <hoffmann@inf.ethz.ch>
 //                 Sylvain Pion
 //                 Laurent Rineau
 
 #ifndef CGAL_CONFIG_H
 #define CGAL_CONFIG_H
+
+#ifdef CGAL_HEADER_ONLY
+#  define CGAL_NO_AUTOLINK 1
+#endif
 
 // Workaround for a bug in Boost, that checks WIN64 instead of _WIN64
 //   https://svn.boost.org/trac/boost/ticket/5519
@@ -54,8 +60,9 @@
 // https://bugreports.qt-project.org/browse/QTBUG-22829
 #ifdef Q_MOC_RUN
 // When Qt moc runs on CGAL files, do not process
-// <boost/type_traits/has_operator.hpp>
+// <boost/type_traits/detail/has_binary_operator.hpp>
 #  define BOOST_TT_HAS_OPERATOR_HPP_INCLUDED
+#  define BOOST_TT_HAS_BIT_AND_HPP_INCLUDED
 #  define BOOST_TT_HAS_BIT_AND_ASSIGN_HPP_INCLUDED
 #  define BOOST_TT_HAS_BIT_OR_HPP_INCLUDED
 #  define BOOST_TT_HAS_BIT_OR_ASSIGN_HPP_INCLUDED
@@ -86,9 +93,32 @@
 // do not include <boost/random.hpp> either
 // it includes <boost/type_traits/has_binary_operator.hpp>
 #  define BOOST_RANDOM_HPP
+// <boost/type_traits/detail/has_prefix_operator.hpp> fails as well
+#  define BOOST_TT_HAS_COMPLEMENT_HPP_INCLUDED
+#  define BOOST_TT_HAS_DEREFERENCE_HPP_INCLUDED
[…]

### 2019-02-07 src/CGAL/float.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Geert-Jan Giezeman, Michael Hemmer
@@ -62,7 +63,7 @@ is_nan_by_mask_float(unsigned int u)
 
 template<>
 class Is_valid< float >
-  : public std::unary_function< float, bool > {
+  : public CGAL::cpp98::unary_function< float, bool > {
   public :
     bool operator()( const float& x ) const {
       float f = x;
@@ -75,7 +76,7 @@ class Is_valid< float >
 
 template<>
 class Is_valid< float >
-  : public std::unary_function< float, bool > {
+  : public CGAL::cpp98::unary_function< float, bool > {
   public :
     bool operator()( const float& x ) const {
       return (x == x);
@@ -92,7 +93,7 @@ template <> class Algebraic_structure_traits< float >
     typedef Tag_true             Is_numerical_sensitive;
 
     class Sqrt
-      : public std::unary_function< Type, Type > {
+      : public CGAL::cpp98::unary_function< Type, Type > {
       public:
         Type operator()( const Type& x ) const {
           return std::sqrt( x );
@@ -100,7 +101,7 @@ template <> class Algebraic_structure_traits< float >
     };
 
     class Kth_root
-      : public std::binary_function<int, Type, Type> {
+      : public CGAL::cpp98::binary_function<int, Type, Type> {
       public:
         Type operator()( int k, const Type& x) const {
           CGAL_precondition_msg( k > 0, "'k' must be positive for k-th roots");
@@ -115,13 +116,16 @@ template <> class Real_embeddable_traits< float >
 public:
 // Is_finite depends on platform
     class Is_finite
-      : public std::unary_function< Type, bool > {
+      : public CGAL::cpp98::unary_function< Type, bool > {
       public:
         bool operator()( const Type& x ) const {
-#ifdef CGAL_CFG_IEEE_754_BUG
+
+#if defined CGAL_CFG_IEEE_754_BUG
           Type f = x;
           IEEE_754_float* p = reinterpret_cast<IEEE_754_float*>(&f);
           return is_finite_by_mask_float( p->c );
+#elif !defined CGAL_CFG_NO_CPP0X_ISFINITE
+          return std::isfinite(x);
 #else
           return (x == x) && (is_valid(x-x));
 #endif

### 2019-02-07 src/CGAL/internal/Lazy_alpha_nt_3.h

@@ -14,22 +14,76 @@
 //
 // $URL$
 // $Id$
-// 
+// SPDX-License-Identifier: GPL-3.0+
 //
 // Author(s)     : Sébastien Loriot <sebastien.loriot@geometryfactory.com>
+//                 Mael Rouxel-Labbé
 
 #ifndef CGAL_INTERNAL_LAZY_ALPHA_NT_3_H
 #define CGAL_INTERNAL_LAZY_ALPHA_NT_3_H
 
+#include <CGAL/license/Alpha_shapes_3.h>
+
 #include <CGAL/assertions.h>
-#include <CGAL/Regular_triangulation_euclidean_traits_3.h>
+#include <CGAL/Cartesian_converter.h>
+#include <CGAL/internal/Exact_type_selector.h>
+#include <CGAL/Has_conversion.h>
+
 #include <boost/shared_ptr.hpp>
 #include <boost/type_traits.hpp>
+#include <boost/optional.hpp>
+
 #include <iostream>
 
 namespace CGAL {
 
 namespace internal{
+
+// check whether Cartesian_converter can do the following conversions
+//  -- Input_traits::(Weighted_)Point_3 to K2::(Weighted_)Point_3
+//  -- Input_traits::(Weighted_)Point_3 to K3::(Weighted_)Point_3
+//
+template < class Input_traits, class Kernel_approx, class Kernel_exact,
+           class Weighted_tag >
+class Is_traits_point_convertible_3
+{
+  typedef typename Kernel_traits<typename Input_traits::Point_3>::Kernel   Kernel_input;
+
+  typedef typename Input_traits::Point_3                                   K1P;
+  typedef typename Kernel_approx::Point_3                                  K2P;
+  typedef typename Kernel_exact::Point_3                                   K3P;
+
+public:
+  static const bool value
+    = (Has_conversion<Kernel_input, Kernel_approx, K1P, K2P>::value &&
+       Has_conversion<Kernel_input, Kernel_exact, K1P, K3P>::value);
+};
+
+template < class Input_traits, class Kernel_approx, class Kernel_exact >
+class Is_traits_point_convertible_3<Input_traits, Kernel_approx, Kernel_exact,
+                                  ::CGAL::Tag_true /* Weighted_tag */>
+{
+  typedef typename Kernel_traits<typename Input_traits::Point_3>::Kernel   Kernel_input;
+
+  typedef typename Input_traits::Weighted_point_3                          K1WP;
+  typedef typename Kernel_approx::Weighted_point_3                         K2WP;
[…]

### 2019-02-07 src/CGAL/internal/Static_filters/Angle_3.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Laurent Rineau
@@ -153,11 +154,11 @@ public:
     //std::cerr << "\n"
     //          << "Now for underflow/overflows...\n"
     //          << "       min_double/eps  = " 
-    //          << std::numeric_limits<double>::min() / err << std::endl
+    //          << (std::numeric_limits<double>::min)() / err << std::endl
     //          << "  sqrt(min_double/eps) = "
-    //          << CGAL::sqrt(std::numeric_limits<double>::min() / err) << std::endl
+    //          << CGAL::sqrt((std::numeric_limits<double>::min)() / err) << std::endl
     //          << "    sqrt(max_double/3) = "
-    //          << CGAL::sqrt(std::numeric_limits<double>::max() / 3) << std::endl;
+    //          << CGAL::sqrt((std::numeric_limits<double>::max)() / 3) << std::endl;
     return err;
   }
 

### 2019-02-07 src/CGAL/internal/Static_filters/Compare_weighted_squared_radius_3.h

@@ -1,10 +1,10 @@
 // Copyright (c) 2009  INRIA Sophia-Antipolis (France).
 // All rights reserved.
 //
-// This file is part of CGAL (www.cgal.org).
-// You can redistribute it and/or modify it under the terms of the GNU
-// General Public License as published by the Free Software Foundation,
-// either version 3 of the License, or (at your option) any later version.
+// This file is part of CGAL (www.cgal.org); you can redistribute it and/or
+// modify it under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation; either version 3 of the License,
+// or (at your option) any later version.
 //
 // Licensees holding a valid commercial license may use this file in
 // accordance with the commercial license agreement provided with the software.
@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 // Author(s)     : Sebastien Loriot
 // This predicates was generated using the fpg tool written by Andreas Meyer.

### 2019-02-07 src/CGAL/internal/Static_filters/Static_filters.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Sylvain Pion
@@ -29,7 +30,7 @@
 // logic and is not plug'n play (requires users providing bounds).
 // If it should be provided again, it should probably be separate.
 
-#include <CGAL/basic.h>
+#include <CGAL/config.h>
 
 #include <CGAL/Kernel/function_objects.h>
 #include <CGAL/Cartesian/function_objects.h>
@@ -37,19 +38,28 @@
 #include <CGAL/internal/Static_filters/tools.h>
 #include <CGAL/internal/Static_filters/Orientation_2.h>
 #include <CGAL/internal/Static_filters/Orientation_3.h>
+#include <CGAL/internal/Static_filters/Collinear_3.h>
 
 // for static filters added nov./dec. 2011
 #ifdef CGAL_DISABLE_STATIC_FILTERS_ADDED_2011
 #  define CGAL_NO_EQUAL_3_STATIC_FILTERS 1
+#  define CGAL_NO_COMPARE_X_2_STATIC_FILTERS 1
 #  define CGAL_NO_IS_DEGENERATE_3_STATIC_FILTERS 1
 #  define CGAL_NO_ANGLE_3_STATIC_FILTERS 1
-#  define CGAL_NO_DO_INTERSECT_3_STATIC_FILTERS 1
+#  define CGAL_NO_DO_INTERSECT_STATIC_FILTERS 1
+
 #endif // CGAL_DISABLE_STATIC_FILTERS_ADDED_2011
 
 
 #ifndef CGAL_NO_EQUAL_3_STATIC_FILTERS
 #  include <CGAL/internal/Static_filters/Equal_3.h>
+#  include <CGAL/internal/Static_filters/Equal_2.h>
 #endif // NOT CGAL_NO_EQUAL_3_STATIC_FILTERS
+
+#ifndef CGAL_NO_COMPARE_X_2_STATIC_FILTERS
+#  include <CGAL/internal/Static_filters/Compare_x_2.h>
+#  include <CGAL/internal/Static_filters/Compare_y_2.h>
+#endif // NOT CGAL_NO_COMPARE_X_2_STATIC_FILTERS
 
 #ifndef CGAL_NO_IS_DEGENERATE_3_STATIC_FILTERS
[…]

### 2019-02-07 src/CGAL/long_double.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 //
 //
 // Author(s)     : Sylvain Pion, Michael Hemmer
@@ -63,7 +64,7 @@ is_nan_by_mask_long_double(unsigned int h, unsigned int l)
 
 template<>
 class Is_valid< long double >
-  : public std::unary_function< long double, bool > {
+  : public CGAL::cpp98::unary_function< long double, bool > {
   public :
     bool operator()( const long double& x ) const {
       double d = x;
@@ -76,7 +77,7 @@ class Is_valid< long double >
 
 template<>
 class Is_valid< long double >
-  : public std::unary_function< long double, bool > {
+  : public CGAL::cpp98::unary_function< long double, bool > {
   public :
     bool operator()( const long double& x ) const {
       return (x == x);
@@ -96,7 +97,7 @@ template <> class Algebraic_structure_traits< long double >
     typedef Tag_true             Is_numerical_sensitive;
 
     class Sqrt
-      : public std::unary_function< Type, Type > {
+      : public CGAL::cpp98::unary_function< Type, Type > {
       public:
         Type operator()( const Type& x ) const {
           return std::sqrt( x );
@@ -104,7 +105,7 @@ template <> class Algebraic_structure_traits< long double >
     };
 
     class Kth_root
-      :public std::binary_function<int, Type, Type > {
+      :public CGAL::cpp98::binary_function<int, Type, Type > {
       public:
         Type operator()( int k,
                                         const Type& x) const {
@@ -121,7 +122,7 @@ template <> class Real_embeddable_traits< long double >
   public:
 
     class To_interval
-      : public std::unary_function< Type, std::pair< double, double > > {
+      : public CGAL::cpp98::unary_function< Type, std::pair< double, double > > {
       public:
         std::pair<double, double> operator()( const Type& x ) const {
           // The conversion long double to double does not always follow the
@@ -146,7 +147,7 @@ template <> class Real_embeddable_traits< long double >
 
 // Is_finite depends on platform
     class Is_finite
-      : public std::unary_function< Type, bool > {
+      : public CGAL::cpp98::unary_function< Type, bool > {
       public:
         bool operator()( const Type& x ) const {
 #ifdef CGAL_CFG_IEEE_754_BUG

### 2019-02-07 src/CGAL/squared_distance_3_1.h

@@ -18,6 +18,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 // 
 //
 // Author(s)     : Geert-Jan Giezeman, Andreas Fabri
@@ -175,6 +176,94 @@ squared_distance(
 }
 
 
+
+template <class K>
+typename K::Comparison_result
+compare_distance_pssC3(
+    const typename K::Point_3 &pt,
+    const typename K::Segment_3 &seg1,
+    const typename K::Segment_3 &seg2,
+    const K& k)
+{
+    typename K::Construct_vector_3 construct_vector;
+    typedef typename K::Vector_3 Vector_3;
+    typedef typename K::RT RT;
+    typedef typename K::FT FT;
+    FT d1=FT(0), d2=FT(0);
+    RT e1 = RT(1), e2 = RT(1);
+    // assert that the segment is valid (non zero length).
+    {
+      Vector_3 diff = construct_vector(seg1.source(), pt);
+      Vector_3 segvec = construct_vector(seg1.source(), seg1.target());
+      RT d = wdot(diff,segvec, k);
+      if (d <= (RT)0){
+        d1 = (FT(diff*diff));
+      }else{
+        RT e = wdot(segvec,segvec, k);
+        if (d > e){
+          d1 = squared_distance(pt, seg1.target(), k);
+        } else{
+          Vector_3 wcr = wcross(segvec, diff, k);
+          d1 = FT(wcr*wcr);
+          e1 = e;
+        }
+      }
+    }
+    {
+      Vector_3 diff = construct_vector(seg2.source(), pt);
+      Vector_3 segvec = construct_vector(seg2.source(), seg2.target());
+      RT d = wdot(diff,segvec, k);
+      if (d <= (RT)0){
+        d2 = (FT(diff*diff));
+      }else{
+        RT e = wdot(segvec,segvec, k);
+        if (d > e){
+          d2 = squared_distance(pt, seg2.target(), k);
+        } else{
+          Vector_3 wcr = wcross(segvec, diff, k);
+          d2 = FT(wcr*wcr);
+          e2 = e;
+        }
+      }
+    }
+    return compare(d1*e2, d2*e1);
+}
+
+template <class K>
+typename K::Comparison_result
+compare_distance_ppsC3(
+    const typename K::Point_3 &pt,
+    const typename K::Point_3 &pt2,
+    const typename K::Segment_3 &seg,
+    const K& k)
+{
[…]

### 2019-02-07 src/CGAL/tss.h

@@ -14,6 +14,7 @@
 //
 // $URL$
 // $Id$
+// SPDX-License-Identifier: LGPL-3.0+
 
 #ifndef CGAL_TSS_H
 #define CGAL_TSS_H
@@ -31,23 +32,65 @@
 
 
 #  ifdef CGAL_USE_BOOST_THREAD
+#    define CGAL_STATIC_THREAD_LOCAL_USE_BOOST 1
+
+#    define CGAL_STATIC_THREAD_LOCAL_VARIABLE_0(TYPE, VAR)               \
+       static boost::thread_specific_ptr<TYPE> VAR##_ptr;                   \
+       if(VAR##_ptr.get() == NULL) {VAR##_ptr.reset(new TYPE());} \
+       TYPE& VAR =  * VAR##_ptr.get()
 
 #    define CGAL_STATIC_THREAD_LOCAL_VARIABLE(TYPE, VAR, ARG1)               \
        static boost::thread_specific_ptr<TYPE> VAR##_ptr;                   \
        if(VAR##_ptr.get() == NULL) {VAR##_ptr.reset(new TYPE(ARG1));} \
        TYPE& VAR =  * VAR##_ptr.get()
 
-#  else
+#    define CGAL_STATIC_THREAD_LOCAL_VARIABLE_2(TYPE, VAR, ARG1, ARG2)       \
+       static boost::thread_specific_ptr<TYPE> VAR##_ptr;                   \
+       if(VAR##_ptr.get() == NULL) {VAR##_ptr.reset(new TYPE(ARG1,ARG2));}  \
+       TYPE& VAR =  * VAR##_ptr.get()
+
+#    define CGAL_STATIC_THREAD_LOCAL_VARIABLE_3(TYPE, VAR, ARG1, ARG2, ARG3)    \
+       static boost::thread_specific_ptr<TYPE> VAR##_ptr;                       \
+       if(VAR##_ptr.get() == NULL) {VAR##_ptr.reset(new TYPE(ARG1,ARG2,ARG3));} \
+       TYPE& VAR =  * VAR##_ptr.get()
+
+#    define CGAL_STATIC_THREAD_LOCAL_VARIABLE_4(TYPE, VAR, ARG1, ARG2, ARG3, ARG4)   \
[…]

### 2019-02-07 src/gudhi/Simplex_tree/Simplex_tree_node_explicit_storage.h

@@ -50,6 +50,16 @@ struct Simplex_tree_node_explicit_storage : SimplexTree::Filtration_simplex_base
     this->assign_filtration(filtration);
   }
 
+  // 2019-02-06, Jisu KIM
+  // Adding copy constructor
+  // for temporarily fixing [-Wclass-memaccess] warning in debian
+  Simplex_tree_node_explicit_storage(
+      const Simplex_tree_node_explicit_storage & stnes)
+      : children_(stnes.children()) {
+    this->assign_filtration(stnes.filtration());
+    this->assign_key(stnes.key());
+  }
+
   /*
    * Assign children to the node
    */
@@ -58,7 +68,10 @@ struct Simplex_tree_node_explicit_storage : SimplexTree::Filtration_simplex_base
   }
 
   /* Careful -> children_ can be NULL*/
-  Siblings * children() {
+  // 2019-02-06, Jisu KIM
+  // Changing to const function for adding copy constructor
+  //Siblings * children() {
+  Siblings * children() const {
     return children_;
   }
 

### 2021-03-11 inst/doc/article.Rnw
version 1.7.5

@@
 
-\author{Brittany T. Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, \and Vincent Rouvreau \\ \small{In collaboration with the CMU TopStat Group}}
-\Plainauthor{Brittany Terese Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, Vincent Rouvreau}
+\author{\parbox{0.7\linewidth}{Brittany T. Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, David L. Millman, \and Vincent Rouvreau} \\
+  \small{In collaboration with the CMU TopStat Group}}
+\Plainauthor{Brittany Terese Fasy, Jisu Kim, Fabrizio Lecci, Cl\'ement Maria, David L. Millman, Vincent Rouvreau}
 
@@
 can be quantified with persistent homology. We provide an \proglang{R} interface
-for the efficient algorithms of the \proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, and \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}},
+for the efficient algorithms of the \proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, and \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}},
 including a function for the persistent homology of the Rips filtration, 
@@
   Jisu Kim\\
-  Department of Statistics\\
-  Carnegie Mellon University\\
+  DataShape Group\\
+  INRIA Saclay\\
   Website: \url{ http://www.stat.cmu.edu/~jisuk/ } \\
-  Email: \email{jisuk1@andrew.cmu.edu}\\
+  Email: \email{jisu.kim@inria.fr}\\
 
@@
   Email: \email{clement.maria@inria.fr}\\
+
+  David L. Millman\\
+  Computer Science Department\\
+  Montana State University\\
+  Website: \url{ http://millman.us } \\
+  Email: \email{david.millman@montana.edu}\\
 
@@
 %library developed at the Institute of Science and Technology Austria;
-%\textsf{\textbf{Dionysus}}\footnote{http://www.mrzv.org/software/dionysus/} is
+%\textsf{\textbf{Dionysus}}\footnote{https://www.mrzv.org/software/dionysus/} is
 %a
@@
 which provides a user-friendly interface for the efficient algorithms of the 
-\proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} \citep{gudhi_stpcoh}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}} \citep{dionysus_C}, and \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}} \citep{phat_C}.
+\proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}} \citep{gudhi_stpcoh}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}} \citep{dionysus_C}, and \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}} \citep{phat_C}.
 
@@
 From version 1.2, \code{gridDiag} works in arbitrary dimension.
-The core of the function is written in \proglang{C++} and the user can choose to compute persistence diagrams using either  the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat }{\pkg{PHAT}}.
+The core of the function is written in \proglang{C++} and the user can choose to compute persistence diagrams using either  the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
 
@@
 The \code{ripsDiag} function computes the persistence diagram of the Rips filtration built on top of a point cloud.
-The user can choose to compute the Rips filtration using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} or \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
-Then for computing the persistence diagram from the Rips filtration, the user can use either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
+The user can choose to compute the Rips filtration using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}} or \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}.
+Then for computing the persistence diagram from the Rips filtration, the user can use either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
 \\The following code generates 60 points from two circles:
@@
 
-The \code{alphaComplexDiag} function computes the Alpha complex filtration built on top of a point cloud, using the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}.
-Then for computing the persistence diagram from the Alpha complex filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
+The \code{alphaComplexDiag} function computes the Alpha complex filtration built on top of a point cloud, using the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}.
+Then for computing the persistence diagram from the Alpha complex filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
 
@@
 
-The \code{alphaShapeDiag} function computes the persistence diagram of the Alpha shape filtration built on top of a point cloud in 3 dimension, using the C++ library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}}.
-Then for computing the persistence diagram from the Alpha shape filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ http://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
+The \code{alphaShapeDiag} function computes the persistence diagram of the Alpha shape filtration built on top of a point cloud in 3 dimension, using the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}.
+Then for computing the persistence diagram from the Alpha shape filtration, the user can use either the C++ library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, or \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}.
 Currently the point data cloud should lie in 3 dimension.
@@
 \noindent Once the filtration is computed, the function \code{filtrationDiag} computes the persistence diagram from the filtration.
-The user can choose to compute the persistence diagram using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software }{\pkg{GUDHI}} or \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
+The user can choose to compute the persistence diagram using either the \proglang{C++} library \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}} or \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}.
 
@@
 Standard metrics for measuring the distance between two persistence diagrams are the bottleneck distance and the $p$th Wasserstein distance \citep{edelsbrunner2010computational}. 
-The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \proglang{C++} library \href{ http://www.mrzv.org/software/dionysus }{\pkg{Dionysus}}.
+The \pkg{TDA} package includes the functions \code{bottleneck} and \code{wasserstein}, which are R wrappers of the functions ``bottleneck\_distance" and ``wasserstein\_distance" of the \proglang{C++} library \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}.
 

### 2021-03-11 src/gudhi/Simplex_tree.h

@@ -281,13 +281,22 @@ class Simplex_tree {
     return filtration_vect_;
   }
 
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing for taking const argument
+  Filtration_simplex_range const& filtration_simplex_range(Indexing_tag = Indexing_tag()) const {
+    return filtration_vect_;
+  }
+
   /** \brief Returns a range over the vertices of a simplex.
    *
    * The order in which the vertices are visited is the decreasing order for < on Vertex_handles,
    * which is consequenlty
    * equal to \f$(-1)^{\text{dim} \sigma}\f$ the canonical orientation on the simplex.
    */
-  Simplex_vertex_range simplex_vertex_range(Simplex_handle sh) {
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing for taking const argument
+  // Simplex_vertex_range simplex_vertex_range(Simplex_handle sh) {
+  Simplex_vertex_range simplex_vertex_range(Simplex_handle sh) const {
     assert(sh != null_simplex());  // Empty simplex
     return Simplex_vertex_range(Simplex_vertex_iterator(this, sh),
                                 Simplex_vertex_iterator(this));
@@ -308,7 +317,10 @@ class Simplex_tree {
    *
    * @param[in] sh Simplex for which the boundary is computed. */
   template<class SimplexHandle>
-  Boundary_simplex_range boundary_simplex_range(SimplexHandle sh) {
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing for taking const argument
+  //Boundary_simplex_range boundary_simplex_range(SimplexHandle sh) {
+  Boundary_simplex_range boundary_simplex_range(SimplexHandle sh) const {
     return Boundary_simplex_range(Boundary_simplex_iterator(this, sh),
                                   Boundary_simplex_iterator(this));
   }
@@ -337,7 +349,10 @@ class Simplex_tree {
   }
 
   /** \brief depth first search, inserts simplices when reaching a leaf. */
-  void rec_copy(Siblings *sib, Siblings *sib_source) {
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing for taking const argument
+  //void rec_copy(Siblings *sib, Siblings * sib_source) {
+  void rec_copy(Siblings *sib, const Siblings * const sib_source) {
     for (auto sh = sib->members().begin(), sh_source = sib_source->members().begin();
          sh != sib->members().end(); ++sh, ++sh_source) {
[…]

### 2021-03-11 src/gudhi/Simplex_tree/Simplex_tree_siblings.h

@@ -69,6 +69,17 @@ class Simplex_tree_siblings {
         members_() {
   }
 
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing [-Wclass-memaccess] warning
+  Simplex_tree_siblings(Simplex_tree_siblings * oncles, Vertex_handle parent, const Dictionary & members)
+      : oncles_(oncles),
+        parent_(parent),
+        members_(members) {
+    for (auto& map_el : members_) {
+      map_el.second.assign_children(this);
+    }
+  }
+
   /* \brief Constructor with initialized set of members.
    *
    * 'members' must be sorted and unique.*/
@@ -76,8 +87,15 @@ class Simplex_tree_siblings {
   Simplex_tree_siblings(Simplex_tree_siblings * oncles, Vertex_handle parent, const RandomAccessVertexRange & members)
       : oncles_(oncles),
         parent_(parent),
-        members_(boost::container::ordered_unique_range, members.begin(),
-                 members.end()) {
+    // 2021-01-29, Jisu KIM
+    // temporarily fixing [-Wclass-memaccess]
+        //members_(boost::container::ordered_unique_range, members.begin(),
+        //         members.end()) {
+        members_() {
+    for (auto& map_el : members) {
+      members_.emplace(map_el.first, map_el.second);
+    }
+
     for (auto& map_el : members_) {
       map_el.second.assign_children(this);
     }
@@ -109,6 +127,12 @@ class Simplex_tree_siblings {
   }
 
   Dictionary & members() {
+    return members_;
+  }
+
+  // 2021-02-08, Jisu KIM
+  // temporarily fixing for taking const argument
+  const Dictionary & members() const {
     return members_;
   }
 

### 2021-03-11 src/rips.h

@@ -54,10 +54,16 @@ inline void ripsFiltration(
     VectorList        & boundary
 ) {
   if (library[0] == 'G') {
-    Gudhi::Simplex_tree<> smplxTree =
+    // 2021-02-08, Jisu KIM
+    // fixing [-Wclass-memaccess] warning
+    //Gudhi::Simplex_tree<> smplxTree =
+    //    RipsFiltrationGudhi< Gudhi::Simplex_tree<> >(X, nSample, nDim,
+    //        maxdimension, maxscale, printProgress, print);
+    //filtrationGudhiToTda< IntVector >(smplxTree, cmplx, values, boundary);
+    filtrationGudhiToTda< IntVector >(
         RipsFiltrationGudhi< Gudhi::Simplex_tree<> >(X, nSample, nDim,
-            maxdimension, maxscale, printProgress, print);
-    filtrationGudhiToTda< IntVector >(smplxTree, cmplx, values, boundary);
+            maxdimension, maxscale, printProgress, print),
+        cmplx, values, boundary);
   }
   else {
 

### 2021-03-11 src/tdautils/gudhiUtils.h

@@ -42,7 +42,7 @@ typedef double     Filtration_value;
  */
 template< typename SimplexTree >
 inline void FiltrationDiagGudhi(
-    SimplexTree    & smplxTree,
+    SimplexTree &    smplxTree,
     const unsigned   coeffFieldCharacteristic,
     const double     minPersistence,
     const unsigned   maxdimension,

### 2025-02-02 inst/doc/article.Rnw
version 1.9.4

@@
 \Abstract{
-We present a short tutorial and introduction to using the \proglang{R} package 
-\pkg{TDA}, which provides some tools for Topological Data Analysis.
-In particular, it includes implementations of functions that, given some data, provide 
-topological information about the underlying space, such as the distance function, the distance to a measure,
-the kNN density estimator, the kernel density estimator, and the kernel distance.
-The salient topological features of the sublevel sets (or superlevel sets) of these functions
-can be quantified with persistent homology. We provide an \proglang{R} interface
-for the efficient algorithms of the \proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, and \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}},
-including a function for the persistent homology of the Rips filtration, 
-and one for the persistent homology of sublevel
-sets (or superlevel sets) of arbitrary functions evaluated over a grid of points.
-The significance of the features in the resulting persistence diagrams can be 
-analyzed with functions that implement the methods discussed in \cite{fasy2014statistical},
-\cite{chazal2014stochastic} and \cite{chazal2014robust}.
-The \proglang{R} package \pkg{TDA} also includes the implementation of an algorithm for density 
-clustering, which allows us to identify the spatial organization of the probability mass associated to 
-a density function and visualize it by means of a dendrogram, the cluster tree. 
+We present a short tutorial and introduction to the \proglang{R} package \pkg{TDA}, which provides tools for Topological Data Analysis.
+The package focuses on statistical analysis of persistent homology and density clustering.
+The package includes implementations of functions that extract topological information about the underlying space from data, such as the distance function, the distance to a measure, the kNN density estimator, the kernel density estimator, and the kernel distance.
+The salient topological features of the sublevel sets (or superlevel sets) of these functions can be quantified with persistent homology.
+This package provides an \proglang{R} interface for the efficient algorithms of the \proglang{C++} libraries \href{ https://project.inria.fr/gudhi/software/ }{\pkg{GUDHI}}, \href{ https://www.mrzv.org/software/dionysus/ }{\pkg{Dionysus}}, and \href{ https://bitbucket.org/phat-code/phat/ }{\pkg{PHAT}}, including a function for the persistent homology of the Rips filtration, and one for the persistent homology of sublevel sets (or superlevel sets) of arbitrary functions evaluated over a grid of points.
+The statistical significance of persistent homology features can be analyzed with functions that implement methods from \cite{FasyLRWBS2014}, \cite{ChazalFLRW2015} and \cite{ChazalFLMRW2017}.
+The \proglang{R} package \pkg{TDA} also includes the implementation of an algorithm for density clustering, which allows us to identify the spatial structure of a probability density function and visualize it by means of a dendrogram, the cluster tree.
 }
@@
   Jisu Kim\\
-  DataShape Group\\
-  INRIA Saclay\\
-  Website: \url{ http://www.stat.cmu.edu/~jisuk/ } \\
-  Email: \email{jisu.kim@inria.fr}\\
+  Department of Statistics\\
+  Seoul National University\\
+  Website: \url{ https://jkim82133.github.io/ } \\
+  Email: \email{jkim82133@snu.ac.kr}\\
 
@@
   Cl\'ement Maria\\
-  School of Mathematics and Physics\\
-  The University of Queensland\\
+  DataShape Group\\
+  INRIA Sophia Antipolis\\
   Website: \url{ http://www-sop.inria.fr/members/Clement.Maria/ } \\
@@
 enough for practical purposes. 
-\cite{fasy2014statistical}, \cite{chazal2014stochastic}, and \cite{chazal2014robust} propose several statistical methods to construct confidence sets for persistence diagrams and other summary functions that allow us to separate topological signal from topological noise.
+\cite{FasyLRWBS2014}, \cite{ChazalFLRW2015}, and \cite{ChazalFLMRW2017} propose several statistical methods to construct confidence sets for persistence diagrams and other summary functions that allow us to separate topological signal from topological noise.
 The methods are implemented in the \pkg{TDA} package and described in Section \ref{sec:persistent}.
@@
 
-\cite{fasy2014statistical} and \cite{chazal2014robust} prove the validity of the bootstrap algorithm for kernel density estimators, distance to measure, and kernel distance, and use it in the framework of persistent homology.
+\cite{FasyLRWBS2014} and \cite{ChazalFLMRW2017} prove the validity of the bootstrap algorithm for kernel density estimators, distance to measure, and kernel distance, and use it in the framework of persistent homology.
 The bootstrap algorithm is implemented in the function \code{bootstrapBand}, which provides the option of parallelizing the algorithm (\code{parallel = TRUE}) using the package \pkg{parallel}.
@@
 The features above the band can be interpreted as representing significant homological features, while points in the band are not significantly different from noise.
-The validity of the bootstrap confidence band for persistence diagrams of KDE, DTM, and Kernel Distance derive from the \textit{Stability Theorem} \citep{chazal2012structure} and is discussed in detail in \cite{fasy2014statistical} and \cite{chazal2014robust}.
+The validity of the bootstrap confidence band for persistence diagrams of KDE, DTM, and Kernel Distance derive from the \textit{Stability Theorem} \citep{chazal2012structure} and is discussed in detail in \cite{FasyLRWBS2014} and \cite{ChazalFLMRW2017}.
 
@@
 Persistence landscapes and silhouettes are real-valued functions that further summarize the information contained in a persistence diagram.
-They have been introduced and studied in \cite{bubenik2012statistical}, \cite{chazal2014stochastic}, and \cite{chazal2014subsampling}.
+They have been introduced and studied in \cite{bubenik2012statistical}, \cite{ChazalFLRW2015}, and \cite{ChazalFLMRW2015}.
 We briefly introduce the two functions.
@@
 
-Recent results in \cite{chazal2014stochastic} and \cite{chazal2014subsampling} show how to construct confidence bands for landscapes 
+Recent results in \cite{ChazalFLRW2015} and \cite{ChazalFLMRW2015} show how to construct confidence bands for landscapes 
 and silhouettes, using a bootstrap algorithm (multiplier bootstrap).
@@
 
-\cite{chazal2014robust} suggest the following method, that we describe here for the kernel density estimator,
+\cite{ChazalFLMRW2017} suggest the following method, that we describe here for the kernel density estimator,
 but works also for the kernel distance and the distance to measure.
